---
title: "trend_analysis"
author: "Dana Lanceman"
date: "10/1/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Prepare data

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = 'C:/Users/danal/OneDrive - UNSW/Desktop/Honours_UNSW/Assessment/cumbung_assessment')
```


```{r load packages}
library(readxl)
library(tidyverse)
library(Kendall)
library(ggeffects)
library(zoo)
library(ggpubr)
library(lubridate)
library(mgcv)
library(tidymv)

options(java.parameters = "-Xmx8000m") 
# the above increases the number of MB of ram that java can heap in order for xlsx package to run properly and save large files into an excel file. This line needs to be run before the xlsx package is loaded in. The number 8000 (MB) can be changed to whatever number you require.
library(xlsx)
```


```{r get daily flow/flood data}
# guess_max lets excel use more rows to decide the data type of the data (missing data makes it think the data type is logical otherwise)
# add a 'wateryear' and 'month' columns

add.date <- function(data){
  data %>% mutate(wateryear = year(Date),
       month = month(Date),
       ym = make_date(wateryear, month)) 
}

cflow <- read_excel("raw/correlate_flow_flood.xlsx",sheet = "cumbung", guess_max = 30000) %>% mutate(wateryear = year(`Date and time`),
       month = month(`Date and time`),
       ym = make_date(wateryear, month))
# check all data types are numerical
str(cflow)

hflow <- read_excel("raw/correlate_flow_flood.xlsx",sheet = "hattah", guess_max = 30000) %>% add.date()
str(hflow)

lflow <- read_excel("raw/correlate_flow_flood.xlsx",sheet = "lowbidgee", guess_max = 30000) %>% add.date()
str(lflow)

yflow <- read_excel("raw/correlate_flow_flood.xlsx",sheet = "yantabulla", guess_max = 30000) %>% add.date() 
str(yflow)

rflow <- read_excel("raw/correlate_flow_flood.xlsx",sheet = "riverland", guess_max = 30000) %>% add.date()
str(rflow)

pflow <- read_excel("raw/correlate_flow_flood.xlsx",sheet = "paroo", guess_max = 30000) %>% add.date()
str(pflow)

gflow <- read_excel("raw/correlate_flow_flood.xlsx",sheet = "gwydir", guess_max = 30000) %>% add.date()
str(gflow)

mflow <- read_excel("raw/correlate_flow_flood.xlsx",sheet = "macquarie", guess_max = 30000) %>% add.date()
str(mflow)


# also get monthly discharge for riverland, which are available long before daily data
rmflow <- read_excel("raw/correlate_flow_flood.xlsx",sheet = "river_monthly", guess_max = 30000) %>% add.date()
str(rmflow)
```



```{r fill gaps and log transform data}
cflow <- cflow %>% mutate(Bool_dis = na.approx(Bool_dis, maxgap = 10, na.rm = TRUE),
                          Bool_lev = na.approx(Bool_lev, maxgap = 10, na.rm = TRUE),
                          Corr_dis = na.approx(Corr_dis, maxgap = 10, na.rm = TRUE),
                          Corr_lev = na.approx(Corr_lev, maxgap = 10, na.rm = TRUE),
                          Four_dis = na.approx(Four_dis, maxgap = 10, na.rm = TRUE),
                          Four_lev = na.approx(Four_lev, maxgap = 10, na.rm = TRUE),
                          l.Four_dis = log(Four_dis + 1),
                          l.Four_lev = log(Four_lev + 1),
                          l.Corr_dis = log(Corr_dis + 1),
                          l.Corr_lev = log(Corr_lev + 1),
                          l.Bool_dis = log(Bool_dis + 1),
                          l.Bool_lev = log(Bool_lev + 1))

hflow <- hflow %>% mutate(Bar_lev = na.approx(Bar_lev, maxgap = 10, na.rm = TRUE),
                          Bar_dis = na.approx(Bar_dis, maxgap = 10, na.rm = TRUE),
                          Col_lev = na.approx(Col_lev, maxgap = 10, na.rm = TRUE),
                          Col_dis = na.approx(Col_dis, maxgap = 10, na.rm = TRUE),
                          l.Bar_lev = log(Bar_lev + 1),
                          l.Bar_dis = log(Bar_dis + 1),
                          l.Col_lev = log(Col_lev + 1),
                          l.Col_dis = log(Col_dis + 1))

lflow <- lflow %>% mutate(Hay_dis = na.approx(Hay_dis, maxgap = 10, na.rm = TRUE),
                          Red_lev = na.approx(Red_lev, maxgap = 10, na.rm = TRUE),
                          Red_dis = na.approx(Red_dis, maxgap = 10, na.rm = TRUE),
                          Hay_lev = na.approx(Hay_lev, maxgap = 10, na.rm = TRUE),
                          l.Hay_dis = log(Hay_dis + 1),
                          l.Red_lev = log(Red_lev + 1),
                          l.Red_dis = log(Red_dis + 1),
                          l.Hay_lev = log(Hay_lev + 1))

gflow <- gflow %>% mutate(Yarr_lev = na.approx(Yarr_lev, maxgap = 10, na.rm = TRUE),
                          Yarr_dis = na.approx(Yarr_dis, maxgap = 10, na.rm = TRUE),
                          l.Yarr_dis = log(Yarr_dis + 1),
                          l.Yarr_lev = log(Yarr_lev + 1))

mflow <- mflow %>% mutate(ox_lev = na.approx(ox_lev, maxgap = 10, na.rm = TRUE),
                          ox_dis = na.approx(ox_dis, maxgap = 10, na.rm = TRUE),
                          l.ox_dis = log(ox_dis + 1),
                          l.ox_lev = log(ox_lev + 1))

pflow <- pflow %>% mutate(will_lev = na.approx(will_lev, maxgap = 10, na.rm = TRUE),
                          will_dis = na.approx(will_dis, maxgap = 10, na.rm = TRUE),
                          l.will_dis = log(will_dis + 1),
                          l.will_lev = log(will_lev + 1))

rflow <- rflow %>% mutate(Flows_dis = na.approx(Flows_dis, maxgap = 10, na.rm = TRUE),
                          RenPS_lev = na.approx(RenPS_lev, maxgap = 10, na.rm = TRUE),
                          l.Flows_dis = log(Flows_dis + 1),
                          l.RenPS_lev = log(RenPS_lev + 1))

yflow <- yflow %>% mutate(Wyan_lev = na.approx(Wyan_lev, maxgap = 10, na.rm = TRUE),
                          Wyan_dis = na.approx(Wyan_dis, maxgap = 10, na.rm = TRUE),
                          Tur_lev = na.approx(Tur_lev, maxgap = 10, na.rm = TRUE),
                          Tur_dis = na.approx(Tur_dis, maxgap = 10, na.rm = TRUE),
                          l.Tur_lev = log(Tur_lev + 1),
                          l.Tur_dis = log(Tur_dis + 1),
                          l.Wyan_dis = log(Wyan_dis + 1),
                          l.Wyan_lev = log(Wyan_lev + 1)) 
```



```{r summarise water level data by month and log transform}
# just cumbung, hattah and lowbidgee

cflow.m <- cflow %>% group_by(ym) %>% summarise(nB = sum(!is.na(Bool_lev)), nF = sum(!is.na(Four_lev)), Bool_levA = mean(Bool_lev, na.rm = TRUE), Four_levA = mean(Four_lev, na.rm = TRUE), Bool_levM = max(Bool_lev, na.rm = TRUE), Four_levM = max(Four_lev, na.rm = TRUE)) %>% 
  mutate(l.Four_lev = log(Four_levA + 1),
        l.Bool_lev = log(Bool_levA + 1),
        l.Four_levM = log(Four_levM + 1),
        l.Bool_levM = log(Bool_levM + 1))

hflow.m <- hflow %>% group_by(ym) %>% summarise(nB = sum(!is.na(Bar_lev)), nC = sum(!is.na(Col_lev)), Bar_levA = mean(Bar_lev, na.rm = TRUE), Bar_levM = max(Bar_lev, na.rm = TRUE), Col_levA = mean(Col_lev, na.rm = TRUE), Col_levM = max(Col_lev, na.rm = TRUE)) %>% 
  mutate(l.Bar_lev = log(Bar_levA + 1),
         l.Col_lev = log(Col_levA + 1),
         l.Bar_levM = log(Bar_levM + 1),
         l.Col_levM = log(Col_levM + 1))

lflow.m <- lflow %>% group_by(ym) %>% summarise(nH = sum(!is.na(Hay_lev)), nR = sum(!is.na(Red_lev)), Hay_levA = mean(Hay_lev, na.rm = TRUE), Hay_levM = max(Hay_lev, na.rm = TRUE), Red_levA = mean(Red_lev, na.rm = TRUE), Red_levM = max(Red_lev, na.rm = TRUE)) %>% 
  mutate(l.Hay_lev = log(Hay_levA + 1),
         l.Red_lev = log(Red_levA + 1),
         l.Hay_levM = log(Hay_levM + 1),
         l.Red_levM = log(Red_levM + 1))

```


```{r summarise data per year and log transform annual data}
cflow.WY <- cflow %>% group_by(wateryear) %>% summarise(nBl = sum(!is.na(Bool_lev)), nFl = sum(!is.na(Four_lev)), nBd = sum(!is.na(Bool_dis)), nFd = sum(!is.na(Four_dis)), Bool_dis = mean(Bool_dis, na.rm = TRUE), Four_dis = mean(Four_dis, na.rm = TRUE), Bool_lev = mean(Bool_lev, na.rm = TRUE), Four_lev = mean(Four_lev, na.rm = TRUE)) %>% 
  mutate(l.Four_dis = log(Four_dis + 1),
        l.Four_lev = log(Four_lev + 1),
        l.Bool_dis = log(Bool_dis + 1),
        l.Bool_lev = log(Bool_lev + 1))

gflow.WY <- gflow %>% group_by(wateryear) %>% summarise(nYl = sum(!is.na(Yarr_lev)), nYd = sum(!is.na(Yarr_dis)), Yarr_dis = mean(Yarr_dis, na.rm = TRUE), Yarr_lev = mean(Yarr_lev, na.rm = TRUE)) %>% 
  mutate(l.Yarr_dis = log(Yarr_dis + 1),
         l.Yarr_lev = log(Yarr_lev + 1))

hflow.WY <- hflow %>% group_by(wateryear) %>% summarise(nBl = sum(!is.na(Bar_lev)), nCl = sum(!is.na(Col_lev)), nBd = sum(!is.na(Bar_dis)), nCd = sum(!is.na(Col_dis)), Bar_dis = mean(Bar_dis, na.rm = TRUE), Bar_lev = mean(Bar_lev, na.rm = TRUE), Col_dis = mean(Col_dis, na.rm = TRUE), Col_lev = mean(Col_lev, na.rm = TRUE)) %>% 
  mutate(l.Bar_lev = log(Bar_lev + 1),
         l.Bar_dis = log(Bar_dis + 1),
         l.Col_lev = log(Col_lev + 1),
         l.Col_dis = log(Col_dis + 1))

lflow.WY <- lflow %>% group_by(wateryear) %>% summarise(nHl = sum(!is.na(Hay_lev)), nRl = sum(!is.na(Red_lev)), nHd = sum(!is.na(Hay_dis)), nRd = sum(!is.na(Red_dis)), Hay_dis = mean(Hay_dis, na.rm = TRUE), Hay_lev = mean(Hay_lev, na.rm = TRUE), Red_lev = mean(Red_lev, na.rm = TRUE), Red_dis = mean(Red_dis, na.rm = TRUE)) %>% 
  mutate(l.Hay_dis = log(Hay_dis + 1),
         l.Red_lev = log(Red_lev + 1),
         l.Red_dis = log(Red_dis + 1),
         l.Hay_lev = log(Hay_lev + 1))

mflow.WY <- mflow %>% group_by(wateryear) %>% summarise(nOl = sum(!is.na(ox_lev)), nOd = sum(!is.na(ox_dis)), ox_lev = mean(ox_lev, na.rm = TRUE), ox_dis = mean(ox_dis, na.rm = TRUE)) %>% 
  mutate(l.ox_dis = log(ox_dis + 1),
         l.ox_lev = log(ox_lev + 1))

pflow.WY <- pflow %>% group_by(wateryear) %>% summarise(nWl = sum(!is.na(will_lev)), nWd = sum(!is.na(will_dis)), will_lev = mean(will_lev, na.rm = TRUE), will_dis = mean(will_dis, na.rm = TRUE)) %>% 
  mutate(l.will_dis = log(will_dis + 1),
         l.will_lev = log(will_lev + 1))

# riverland - summarise daily data, and summarise monthly data
rflow.WY <- rflow %>% group_by(wateryear) %>% summarise(nFd = sum(!is.na(Flows_dis)), nRl = sum(!is.na(RenPS_lev)), Flows_dis = mean(Flows_dis, na.rm = TRUE), RenPS_lev = mean(RenPS_lev, na.rm = TRUE)) %>% 
  mutate(l.Flows_dis = log(Flows_dis + 1),
         l.RenPS_lev = log(RenPS_lev + 1))
rmflow.WY <- rmflow %>% group_by(wateryear) %>% 
  summarise(nFd = sum(!is.na(Flows_dis)), Flows_dis = mean(Flows_dis, na.rm = TRUE)) %>% 
  mutate(l.Flows_dis = log(Flows_dis + 1))


yflow.WY <- yflow %>% group_by(wateryear) %>% summarise(nTl = sum(!is.na(Tur_lev)), nTd = sum(!is.na(Tur_dis)), nWl = sum(!is.na(Wyan_lev)), nWd = sum(!is.na(Wyan_dis)), Wyan_lev = mean(Wyan_lev, na.rm = TRUE), Wyan_dis = mean(Wyan_dis, na.rm = TRUE), Tur_lev = mean(Tur_lev, na.rm = TRUE), Tur_dis = mean(Tur_dis, na.rm = TRUE)) %>% 
  mutate(l.Tur_lev = log(Tur_lev + 1),
         l.Tur_dis = log(Tur_dis + 1),
         l.Wyan_dis = log(Wyan_dis + 1),
         l.Wyan_lev = log(Wyan_lev + 1)) 
```

```{r Create an inundation vs booligal discharge dataframe}
#Add Booligal 95% complete years to a new dataframe
inun <- tibble(.rows = 69) %>% 
  add_column(Bool_dis = cflow.WY %>% filter(nBd >= 347) %>% .$Bool_dis,
             year = cflow.WY %>% filter(nBd >= 347) %>% .$wateryear)

# load annual inundation data 
cinun <- read_excel("raw/correlate_flow_flood.xlsx",sheet = "cumbung_inun") 

# combine these 
inun <- merge(inun, cinun, by = 'year') %>% 
  mutate(l.Bool_dis = log(Bool_dis + 1),
         l.area.km2 = log(area.km2 + 1))
```


# GLMs for correlating annual discharge data

```{r Cumbung}
# generalised linear model
# filter data within the model so that the actual dataset remains unchanged
M1 <- glm(l.Four_dis ~ l.Bool_dis, data=cflow.WY %>% filter(nFd >= 347, nBd >= 347), na.action = na.omit)

# check assumptions
par(mfrow = c(2, 2))
plot(M1) # top left test for equal variances, top right normality 
par(mfrow = c(1, 1))
ggplot(cflow.WY %>% filter(nFd >= 347, nBd >= 347), aes(l.Bool_dis, l.Four_dis), na.rm = TRUE) + geom_point()

# very small dataset, assumptions all ok

summary(M1)
cor.test(cflow.WY$Bool_dis, cflow.WY$Four_dis)

# predict
cflow.WY$predict.l <- predict(M1, cflow.WY)
# create a variable that is predicted log colignan dis
cflow.WY$predict <- exp(cflow.WY$predict.l) - 1
# back transform 
plot(cflow.WY$Four_dis ~ cflow.WY$predict)


ggemmeans.Dis <- ggemmeans(M1, terms = c("l.Bool_dis[all]"))

plot(ggemmeans.Dis, colors = "hero", ci = T, add.data = TRUE, jitter = 0, show.y.title = T, show.title = FALSE) + labs(x="Booligal Discharge",y="Four mile weir Discharge",title="")


# make a new column that is a combination of the old and new data
# if the raw data aren't an NA, it uses the raw data. otherwise, it uses the predicted data
# also make a new column which is nNd (number of data points in each for the new data)
cflow.WY <- cflow.WY %>% mutate(new = case_when(!is.na(Four_dis) ~ Four_dis, TRUE ~ predict),
                                nNd = case_when(nFd != 0 ~ nFd, TRUE ~ nBd))

# check against just predicted data
plot(cflow.WY$new ~ cflow.WY$predict)
# as expected, almost all a 1:1 ratio as most data are from booligal, but some deviation due to four mile data - good
```

```{r Hattah lakes}
# generalised linear model
M1 <- glm(l.Col_dis ~ l.Bar_dis, data=hflow.WY %>% filter(nCd >= 347, nBd >= 347), na.action = na.omit)

# check assumptions
par(mfrow = c(2, 2))
plot(M1) # top left test for equal variances, top right normality 
par(mfrow = c(1, 1))
ggplot(hflow.WY %>% filter(nCd >= 347, nBd >= 347), aes(l.Bar_dis, l.Col_dis), na.rm = TRUE) +
  geom_point()

# log-transform makes normality and equal variance assumptions ok
# linearity is great with log-transformations
# therefore use log-transformed data

summary(M1)
cor.test(hflow.WY$l.Bar_dis, hflow.WY$l.Col_dis)

# predict
hflow.WY$predict.l <- predict(M1, hflow.WY) 
# create a variable that is predicted log colignan dis
hflow.WY$predict <- exp(hflow.WY$predict.l) - 1
# back transform 
plot(hflow.WY$Col_dis ~ hflow.WY$predict)
# visualise relationship between original and predicted colignan discharge

# plot correlation
ggemmeans.Dis <- ggemmeans(M1, terms = c("l.Bar_dis[all]"))
plot(ggemmeans.Dis, colors = "hero", ci = T, add.data = TRUE, jitter = 0, show.y.title = T, show.title = FALSE) + labs(x="Barham Discharge (log)",y="Colignan Discharge (log)",title="")

# export
#write.csv(hflow.WY, "output/hattah_pred.csv")


# make a new column that is a combination of the old and new data
hflow.WY <- hflow.WY %>% mutate(new = case_when(!is.na(Col_dis) ~ Col_dis, TRUE ~ predict),
                                nNd = case_when(nCd != 0 ~ nCd, TRUE ~ nBd))

# check against just predicted data
plot(hflow.WY$new ~ hflow.WY$predict) # as expected
```

```{r Lowbidgee}
# generalised linear model
M1 <- glm(l.Red_dis ~ l.Hay_dis, data=lflow.WY %>% filter(nHd >= 347, nRd >= 347), na.action = na.omit)

# check assumptions
par(mfrow = c(2, 2))
plot(M1) # top left test for equal variances, top right normality 
par(mfrow = c(1, 1))
ggplot(lflow.WY %>% filter(nHd >= 347, nRd >= 347), aes(l.Hay_dis, l.Red_dis), na.rm = TRUE) +
  geom_point()

# log-transform makes normality and equal variance assumptions good
# linearity fine in all cases
# therefore use log-transformed data

summary(M1)
cor.test(lflow.WY$l.Hay_dis, lflow.WY$l.Red_dis)

# predict
lflow.WY$predict.l <- predict(M1, lflow.WY)
# create a variable that is predicted log redbank dis
lflow.WY$predict <- exp(lflow.WY$predict.l) - 1
# back transform 
plot(lflow.WY$Red_dis ~ lflow.WY$predict)
# visualise relationship between original and predicted redbank discharge

# visualise correlation
ggemmeans.Dis <- ggemmeans(M1, terms = c("l.Hay_dis[all]"))

plot(ggemmeans.Dis, colors = "hero", ci = T, add.data = TRUE, jitter = 0, show.y.title = T, show.title = FALSE) + labs(x="Hay Discharge (log)",y="Redbank Discharge (log)",title="")

# export
#write.csv(lflow.WY, "output/lowbidgee_pred.csv")

# make a new column that is a combination of the old and new data
lflow.WY <- lflow.WY %>% mutate(new = case_when(!is.na(Red_dis) ~ Red_dis, TRUE ~ predict),
                                nNd = case_when(nRd != 0 ~ nRd, TRUE ~ nHd))

# check against just predicted data
plot(lflow.WY$new ~ lflow.WY$predict) # as expected
```



# Correlating Inundation area with Booligal discharge

```{r Cumbung Inundation area vs Booligal discharge}
# generalised linear model
M1 <- glm(l.area.km2 ~ l.Bool_dis, data=inun)
summary(M1)
cor.test(inun$l.Bool_dis, inun$l.area.km2)

 # check assumptions
par(mfrow = c(2, 2))
plot(M1) # top left test for equal variances, top right normality
# better with log transformed data, acceptable
par(mfrow = c(1, 1))
ggplot(inun, aes(l.Bool_dis, l.area.km2)) +
  geom_point() # test for linear relationship
# log-log relationship looks roughly linear

# predict
inun$predict.l <- predict(M1, inun)
# create a variable that is predicted log inundation area
inun$predict <- exp(inun$predict.l)
# back transform 
plot(inun$area.km2 ~ inun$predict)



# visualise relationship between original and predicted inundation
ggemmeans.Inun <- ggemmeans(M1, terms = c("l.Bool_dis[all]")) # this must specify the second term listed in glm

plot(ggemmeans.Inun, facet = TRUE, colors = "hero", ci = T, add.data = TRUE, jitter = 0, show.y.title = T, show.title = FALSE) + labs(x="log Booligal discharge",y="log Cumbung inundation area",title="")


# export
#write.csv(inun, "output/cbool_pred_inun.csv")
```



# Generalised additive models - monthly flood data - Cumbung, Hattah, Lowbidgee

Using GAMs instead of GLMs as GLMs don't fit well and assumptions aren't acceptable.

```{r Cumbung}
# modelling monthly data
m <- gam(l.Bool_lev ~   s(l.Four_lev, k=3), data=cflow.m %>% filter(nB >= 29, nF >= 29) %>%  subset(Bool_levA > 0, Four_levA > 0))
# k = 3 is 3rd order polynomial
# only include months with >95% data and water levels above zero


# check normality and equal variances (linearity is not necessary)
par(mfrow = c(2, 2))
gam.check(m) # residual plots for GAMs
# need to log-transform to improve normality

# predict
cflow.m$predict.l <- predict(m)
cflow.m$predict <- exp(cflow.m$predict.l) - 1

# plot
ggpredict <- ggemmeans(m, terms = c("l.Four_lev [all]"))
print(ggplot(ggpredict,aes(x = x, y = predicted))+
        geom_ribbon(aes(ymin =conf.low, ymax = conf.high),alpha=0.2) +
        geom_line(size=0.5)+
        geom_point(data=cflow.m %>% filter(nB >= 29, nF >= 29) %>%  subset(Bool_levA > 0, Four_levA > 0), aes(y=l.Bool_lev, x=l.Four_lev))+
        labs(colour = get_legend_title(ggpredict),
             x = get_x_title(ggpredict),
             y = get_y_title(ggpredict),
             title = get_title(ggpredict))+
        theme_bw(base_size=16) +
        theme(plot.title = element_blank(),
              panel.grid.major = element_blank(),
              panel.grid.minor=element_blank(),
              plot.background = element_blank(),
              panel.background = element_rect(fill =  NA), 
              panel.border = element_rect(fill=NA),
              axis.text.y = element_text(size = 10),
              axis.text.x = element_text(size = 15),
              axis.title.y = element_text(size = 15, vjust= 1),
              axis.title.x = element_text(size = 15))  + 
        #scale_y_continuous(limits = c(0,1)) + 
        scale_x_continuous(labels = scales::comma)+
        labs(x="Log Booligal water level", y=paste0("Log Cumbung water level")))

```


```{r Hattah}
# prep data
#hflow.m <- hflow.m %>% 
#  subset(Bar_levA > 0, Col_levA > 0) %>% # remove negative water level data
#  filter(nB >= 29, nC >= 29) # only include months with >95% data


# modelling monthly data
m <- gam(Bar_levA ~   s(Col_levA, k=3), data=hflow.m %>% filter(nB >= 29, nC >= 29) %>%  subset(Bar_levA > 0, Col_levA > 0))
# k = 3 is 3rd order polynomial
# only include months with >95% data and water levels above zero


# check normality and equal variances (linearity is not necessary)
par(mfrow = c(2, 2))
gam.check(m) # residual plots for GAMs

# predict
hflow.m$predict <- predict(m)

# plot
ggpredict <- ggemmeans(m, terms = c("Col_levA [all]"))
print(ggplot(ggpredict,aes(x = x, y = predicted))+
        geom_ribbon(aes(ymin =conf.low, ymax = conf.high),alpha=0.2) +
        geom_line(size=0.5)+
        geom_point(data=hflow.m %>% filter(nB >= 29, nC >= 29) %>%  subset(Bar_levA > 0, Col_levA > 0), aes(y=Bar_levA, x=Col_levA))+
        labs(colour = get_legend_title(ggpredict),
             x = get_x_title(ggpredict),
             y = get_y_title(ggpredict),
             title = get_title(ggpredict))+
        theme_bw(base_size=16) +
        theme(plot.title = element_blank(),
              panel.grid.major = element_blank(),
              panel.grid.minor=element_blank(),
              plot.background = element_blank(),
              panel.background = element_rect(fill =  NA), 
              panel.border = element_rect(fill=NA),
              axis.text.y = element_text(size = 10),
              axis.text.x = element_text(size = 15),
              axis.title.y = element_text(size = 15, vjust= 1),
              axis.title.x = element_text(size = 15))  + 
        #scale_y_continuous(limits = c(0,1)) + 
        scale_x_continuous(labels = scales::comma)+
        labs(x="Avg.Bar.M [GL]", y=paste0("Avg.Col.M")))
```


```{r Lowbidgee}
# modelling monthly data
m <- gam(Hay_levA ~   s(Red_levA, k=3), data=lflow.m %>% filter(nH >= 29, nR >= 29) %>%  subset(Hay_levA > 0, Red_levA > 0))
# k = 3 is 3rd order polynomial
# only include months with >95% data and water levels above zero

# check normality and equal variances (linearity is not necessary)
par(mfrow = c(2, 2))
gam.check(m) # residual plots for GAMs

# predict
lflow.m$predict <- predict(m)

# plot
ggpredict <- ggemmeans(m, terms = c("Red_levA [all]"))
print(ggplot(ggpredict,aes(x = x, y = predicted))+
        geom_ribbon(aes(ymin =conf.low, ymax = conf.high),alpha=0.2) +
        geom_line(size=0.5)+
        geom_point(data=lflow.m %>% filter(nH >= 29, nR >= 29) %>%  subset(Hay_levA > 0, Red_levA > 0), aes(y=Hay_levA, x=Red_levA))+
        labs(colour = get_legend_title(ggpredict),
             x = get_x_title(ggpredict),
             y = get_y_title(ggpredict),
             title = get_title(ggpredict))+
        theme_bw(base_size=16) +
        theme(plot.title = element_blank(),
              panel.grid.major = element_blank(),
              panel.grid.minor=element_blank(),
              plot.background = element_blank(),
              panel.background = element_rect(fill =  NA), 
              panel.border = element_rect(fill=NA),
              axis.text.y = element_text(size = 10),
              axis.text.x = element_text(size = 15),
              axis.title.y = element_text(size = 15, vjust= 1),
              axis.title.x = element_text(size = 15))  + 
        #scale_y_continuous(limits = c(0,1)) + 
        scale_x_continuous(labels = scales::comma)+
        labs(x="Avg.Bar.M [GL]", y=paste0("Avg.Col.M")))
```



# Time series analysis - Kendall rank correlations

```{r cumbung}
#TS.Dis <- ts(ctrend$Discharge[1:75], 1946, 2020)
TS.Inun <- ts(ctrend$Inun_area[42:75], 1987, 2020)
#TS.Flood <- ts(ctrend$Flood_days[1:75], 1946, 2020)

#MannKendall(TS.Dis)
MannKendall(TS.Inun)
#MannKendall(TS.Flood)

ggplot(ctrend, aes(x = Year, y = Inun_area)) +
  geom_line()

ggplot(ctrend, aes(x = Year, y = Inun_5yr)) +
  geom_line() +
  stat_smooth(method = lm)

#ggplot(ctrend, aes(x = Year, y = Discharge)) +
  #geom_line()

#ggplot(ctrend, aes(x = Year, y = Flood_days)) +
  #geom_line()
```


```{r booligal}
TS.Dis <- ts(btrend$Discharge[1:71], 1946, 2020)
TS.Flood <- ts(btrend$Flood[1:71], 1946, 2020)

MannKendall(TS.Dis)
MannKendall(TS.Flood)

ggplot(btrend, aes(x = Year, y = Discharge)) +
  geom_line()

ggplot(btrend, aes(x = Year, y = Flood)) +
  geom_line()
```


```{r gwydir}
TS.Dis <- ts(gtrend$Discharge[1:52], 1930, 2020)
TS.Flood <- ts(gtrend$Flood[1:52], 1930, 2020)

MannKendall(TS.Dis)
MannKendall(TS.Flood)

ggplot(gtrend, aes(x = Year, y = Discharge)) +
  geom_line()

ggplot(gtrend, aes(x = Year, y = Flood)) +
  geom_line()
```


```{r hattah}
TS.Dis <- ts(htrend$Discharge[1:110], 1913, 2020)
#TS.Flood <- ts(htrend$Flood[1:75], 1946, 2020)

MannKendall(TS.Dis)
#MannKendall(TS.Flood)

ggplot(htrend, aes(x = Year, y = Discharge)) +
  geom_line()

#ggplot(htrend, aes(x = Year, y = Flood)) +
  #geom_line()
```


```{r lowbidgee}
TS.Dis <- ts(ltrend$Discharge[1:108], 1915, 2020)
TS.Flood <- ts(ltrend$Flood[63:108], 1976, 2020)

MannKendall(TS.Dis)
MannKendall(TS.Flood)

ggplot(ltrend, aes(x = Year, y = Discharge)) +
  geom_line()

ggplot(ltrend, aes(x = Year, y = Flood)) +
  geom_line()
```


```{r macquarie}
TS.Dis <- ts(mtrend$Discharge[1:64], 1955, 2018)
TS.Flood <- ts(mtrend$Flood[1:64], 1955, 2018)

MannKendall(TS.Dis)
MannKendall(TS.Flood)

ggplot(mtrend, aes(x = Year, y = Discharge)) +
  geom_line()

ggplot(mtrend, aes(x = Year, y = Flood)) +
  geom_line()
```


```{r paroo}
TS.Dis <- ts(ptrend$Discharge[1:47], 1976, 2020)
TS.Flood <- ts(ptrend$Flood[1:47], 1976, 2020)

MannKendall(TS.Dis)
MannKendall(TS.Flood)

ggplot(ptrend, aes(x = Year, y = Discharge)) +
  geom_line()

ggplot(ptrend, aes(x = Year, y = Flood)) +
  geom_line()
```


```{r riverland}
TS.Dis <- ts(rtrend$Discharge[1:121], 1902, 2020)
TS.Flood <- ts(rtrend$Flood[75:121], 1975, 2020)

MannKendall(TS.Dis)
MannKendall(TS.Flood)

ggplot(rtrend, aes(x = Year, y = Discharge)) +
  geom_line()

ggplot(rtrend, aes(x = Year, y = Flood)) +
  geom_line()
```


```{r yantabulla}
TS.Dis <- ts(ytrend$Discharge[1:27], 1994, 2020)
TS.Flood <- ts(ytrend$Flood[1:27], 1994, 2020)

MannKendall(TS.Dis)
MannKendall(TS.Flood)

ggplot(ytrend, aes(x = Year, y = Discharge)) +
  geom_line()

ggplot(ytrend, aes(x = Year, y = Flood)) +
  geom_line()
```


# Discharge Time series analysis - generalised linear models

```{r cumbung}
# remove years with too few data
cflow.WY.c <- cflow.WY %>% filter(nNd >= 347)

M1 <- glm(new ~ wateryear, data = cflow.WY.c)

# assumptions
par(mfrow = c(2, 2))
plot(M1) # OK
par(mfrow = c(1, 1)) 
ggplot(cflow.WY.c, aes(wateryear, new), na.rm = TRUE) +
  geom_point() 
# no obvious non-linear trends


# predict
cflow.WY.c$predict.Time <- predict(M1, cflow.WY.c)
# create a variable that is predicted four mile discharge over time


# plot
ggemmeans.dis <- ggemmeans(M1, terms = c("wateryear[all]"))

plot(ggemmeans.dis, colors = "hero", ci = T, add.data = TRUE, jitter = 0, show.y.title = T, show.title = FALSE) + labs(x="Year",y="Predicted Four mile weir discharge (ML/day)",title="")


# test
summary(M1)
cor.test(cflow.WY.c$new, cflow.WY.c$wateryear)


# calculate differences over time
100-(cflow.WY.c$predict.Time[length(cflow.WY.c$predict.Time)]/cflow.WY.c$predict.Time[1]*100)
# diff in last 50 years - since 1970, which is row 22
100-(cflow.WY.c$predict.Time[length(cflow.WY.c$predict.Time)]/cflow.WY.c$predict.Time[22]*100)

# average annual % change
(100-(cflow.WY.c$predict.Time[length(cflow.WY.c$predict.Time)]/cflow.WY.c$predict.Time[1]*100))/length(cflow.WY.c$predict.Time)

# prepare data for plotting
cflow.WY.cp <- cflow.WY.c %>% rename(Year = wateryear,  Discharge = new) %>% 
  complete(Year = seq(1946, 2020, by = 1)) # add gaps for missing years
```



```{r booligal}
# remove years with too few data
cflow.WY.b <- cflow.WY %>% filter(nBd >= 347)

M1 <- glm(l.Bool_dis ~ wateryear, data = cflow.WY.b)

# assumptions
par(mfrow = c(2, 2))
plot(M1) # use log
par(mfrow = c(1, 1)) 
ggplot(cflow.WY.b, aes(wateryear, l.Bool_dis), na.rm = TRUE) +
  geom_point() 
# no obvious non-linear trends


# predict
cflow.WY.b$predict.Time.l <- predict(M1, cflow.WY.b)
# create a variable that is predicted log booligal discharge over time
cflow.WY.b$predict.Time <- exp(cflow.WY.b$predict.Time.l)-1
plot(cflow.WY.b$Bool_dis ~ cflow.WY.b$predict.Time)
plot(cflow.WY.b$predict.Time ~ cflow.WY.b$wateryear)


# plot
ggemmeans.dis <- ggemmeans(M1, terms = c("wateryear[all]"))

plot(ggemmeans.dis, colors = "hero", ci = T, add.data = TRUE, jitter = 0, show.y.title = T, show.title = FALSE) + labs(x="Year",y="log Booligal discharge (ML/day)",title="")


# test
summary(M1)
cor.test(cflow.WY.b$l.Bool_dis, cflow.WY.b$wateryear)


# calculate differences over time
100-(cflow.WY.b$predict.Time[length(cflow.WY.b$predict.Time)]/cflow.WY.b$predict.Time[1]*100)
# diff in last 50 years - since 1970, which is row 22
100-(cflow.WY.b$predict.Time[length(cflow.WY.b$predict.Time)]/cflow.WY.b$predict.Time[22]*100)

# average annual % change
(100-(cflow.WY.b$predict.Time[length(cflow.WY.b$predict.Time)]/cflow.WY.b$predict.Time[1]*100))/length(cflow.WY.b$predict.Time)


# prepare data for plotting
cflow.WY.bp <- cflow.WY.b %>% rename(Year = wateryear,  Discharge = Bool_dis) %>% 
  complete(Year = seq(1946, 2020, by = 1)) # add gaps for missing years
```



```{r gwydir}
# remove years with too few data
gflow.WY.g <- gflow.WY %>% filter(nYd >= 347)

M1 <- glm(l.Yarr_dis ~ wateryear, data = gflow.WY.g)

# assumptions
par(mfrow = c(2, 2))
plot(M1) # use log
par(mfrow = c(1, 1)) 
ggplot(gflow.WY.g, aes(wateryear, l.Yarr_dis), na.rm = TRUE) +
  geom_point() 
# no obvious non-linear trends


# predict
gflow.WY.g$predict.Time.l <- predict(M1, gflow.WY.g)
# create a variable that is predicted log gwydir discharge over time
gflow.WY.g$predict.Time <- exp(gflow.WY.g$predict.Time.l)-1
plot(gflow.WY.g$Yarr_dis ~ gflow.WY.g$predict.Time)
plot(gflow.WY.g$predict.Time ~ gflow.WY.g$wateryear)


# plot
ggemmeans.dis <- ggemmeans(M1, terms = c("wateryear[all]"))

plot(ggemmeans.dis, colors = "hero", ci = T, add.data = TRUE, jitter = 0, show.y.title = T, show.title = FALSE) + labs(x="Year",y="log Gwydir discharge (ML/day)",title="")


# test
summary(M1)
cor.test(gflow.WY.g$l.Yarr_dis, gflow.WY.g$wateryear)


# calculate differences over time
100-(gflow.WY.g$predict.Time[length(gflow.WY.g$predict.Time)]/gflow.WY.g$predict.Time[1]*100)
# diff in last 50 years - since 1971, which is row 3
100-(gflow.WY.g$predict.Time[length(gflow.WY.g$predict.Time)]/gflow.WY.g$predict.Time[3]*100)

# prepare data for plotting
gflow.WY.gp <- gflow.WY.g %>% rename(Year = wateryear,  Discharge = Yarr_dis) %>% 
  complete(Year = seq(1930, 2020, by = 1)) # add gaps for missing years
```


```{r hattah}
# remove years with too few data
hflow.WY.h <- hflow.WY %>% filter(nNd >= 347)

M1 <- glm(new ~ wateryear, data = hflow.WY.h)

# assumptions
par(mfrow = c(2, 2))
plot(M1) # non-transformed data looks ok
par(mfrow = c(1, 1)) 
ggplot(hflow.WY.h, aes(wateryear, new), na.rm = TRUE) +
  geom_point() 
# no obvious non-linear trends


# predict
hflow.WY.h$predict.Time <- predict(M1, hflow.WY.h)
# create a variable that is predicted hattah discharge over time


# plot
ggemmeans.dis <- ggemmeans(M1, terms = c("wateryear[all]"))

plot(ggemmeans.dis, colors = "hero", ci = T, add.data = TRUE, jitter = 0, show.y.title = T, show.title = FALSE) + labs(x="Year",y="predicted Colignan discharge (ML/day)",title="")


# test
summary(M1)
cor.test(hflow.WY.h$new, hflow.WY.h$wateryear)


# calculate differences over time
100-(hflow.WY.h$predict.Time[length(hflow.WY.h$predict.Time)]/hflow.WY.h$predict.Time[1]*100)
# diff in last 50 years - since 1970, which is row 55
100-(hflow.WY.h$predict.Time[length(hflow.WY.h$predict.Time)]/hflow.WY.h$predict.Time[55]*100)

# prepare data for plotting
hflow.WY.hp <- hflow.WY.h %>% rename(Year = wateryear,  Discharge = new) %>% 
  complete(Year = seq(1913, 2020, by = 1)) # add gaps for missing years
```


```{r lowbidgee}
# remove years with too few data
lflow.WY.l <- lflow.WY %>% filter(nNd >= 347) %>% mutate(new.l = log(new + 1))

M1 <- glm(new.l ~ wateryear, data = lflow.WY.l)

# assumptions
par(mfrow = c(2, 2))
plot(M1) # use log
par(mfrow = c(1, 1)) 
ggplot(cflow.WY, aes(wateryear, new.l), na.rm = TRUE) +
  geom_point() 
# no obvious non-linear trends


# predict
lflow.WY.l$predict.Time.l <- predict(M1, lflow.WY.l)
# create a variable that is predicted log booligal discharge over time
lflow.WY.l$predict.Time <- exp(lflow.WY.l$predict.Time.l)-1
plot(lflow.WY.l$new ~ lflow.WY.l$predict.Time)
plot(lflow.WY.l$predict.Time ~ lflow.WY.l$wateryear)


# plot
ggemmeans.dis <- ggemmeans(M1, terms = c("wateryear[all]"))

plot(ggemmeans.dis, colors = "hero", ci = T, add.data = TRUE, jitter = 0, show.y.title = T, show.title = FALSE) + labs(x="Year",y="log predicted Redbank discharge (ML/day)",title="")


# test
summary(M1)
cor.test(lflow.WY.l$new.l, lflow.WY.l$wateryear)


# calculate differences over time
100-(lflow.WY.l$predict.Time[length(lflow.WY.l$predict.Time)]/lflow.WY.l$predict.Time[1]*100)
# diff in last 50 years - since 1970, which is row 55
100-(lflow.WY.l$predict.Time[length(lflow.WY.l$predict.Time)]/lflow.WY.l$predict.Time[55]*100)

# prepare data for plotting
lflow.WY.lp <- lflow.WY.l %>% rename(Year = wateryear,  Discharge = new) %>% 
  complete(Year = seq(1915, 2020, by = 1)) # add gaps for missing years
```


```{r macquarie}
# remove years with too few data
mflow.WY.m <- mflow.WY %>% filter(nOd >= 347)

M1 <- glm(l.ox_dis ~ wateryear, data = mflow.WY.m)

# assumptions
par(mfrow = c(2, 2))
plot(M1) #  log looks somewhat better?
par(mfrow = c(1, 1)) 
ggplot(mflow.WY.m, aes(wateryear, l.ox_dis), na.rm = TRUE) +
  geom_point() 
# without transformation, the trend looks somewhat non-linear, decreases massively from 1955 to 1980 then decreases more slowly
# log is better


# predict
mflow.WY.m$predict.Time.l <- predict(M1, mflow.WY.m)
# create a variable that is predicted log booligal discharge over time
mflow.WY.m$predict.Time <- exp(mflow.WY.m$predict.Time.l)-1
plot(mflow.WY.m$ox_dis ~ mflow.WY.m$predict.Time)
plot(mflow.WY.m$predict.Time ~ mflow.WY.m$wateryear)


# plot
ggemmeans.dis <- ggemmeans(M1, terms = c("wateryear[all]"))

plot(ggemmeans.dis, colors = "hero", ci = T, add.data = TRUE, jitter = 0, show.y.title = T, show.title = FALSE) + labs(x="Year",y="Log Oxley discharge (ML/day)",title="")


# test
summary(M1)
cor.test(mflow.WY.m$l.ox_dis, mflow.WY.m$wateryear)


# calculate differences over time
100-(mflow.WY.m$predict.Time[length(mflow.WY.m$predict.Time)]/mflow.WY.m$predict.Time[1]*100)
# diff in last 50 years - since 1970, which is row 13
100-(lflow.WY.l$predict.Time[length(lflow.WY.l$predict.Time)]/lflow.WY.l$predict.Time[13]*100)

# prepare data for plotting
mflow.WY.mp <- mflow.WY.m %>% rename(Year = wateryear,  Discharge = ox_dis) %>% 
  complete(Year = seq(1955, 2018, by = 1)) # add gaps for missing years
```


```{r paroo}
# remove years with too few data
pflow.WY.p <- pflow.WY %>% filter(nWd >= 347)

M1 <- glm(l.will_dis ~ wateryear, data = pflow.WY.p)

# assumptions
par(mfrow = c(2, 2))
plot(M1) # log is better
par(mfrow = c(1, 1)) 
ggplot(pflow.WY.p, aes(wateryear, will_dis), na.rm = TRUE) +
  geom_point() 
# no obvious non-linear trends


# predict
pflow.WY.p$predict.Time.l <- predict(M1, pflow.WY.p)
# create a variable that is predicted log booligal discharge over time
pflow.WY.p$predict.Time <- exp(pflow.WY.p$predict.Time.l)-1
plot(pflow.WY.p$will_dis ~ pflow.WY.p$predict.Time)
plot(pflow.WY.p$predict.Time ~ pflow.WY.p$wateryear)


# plot
ggemmeans.dis <- ggemmeans(M1, terms = c("wateryear[all]"))

plot(ggemmeans.dis, colors = "hero", ci = T, add.data = TRUE, jitter = 0, show.y.title = T, show.title = FALSE) + labs(x="Year",y="log Willara Xing discharge (ML/day)",title="")


# test
summary(M1)
cor.test(pflow.WY.p$l.will_dis, pflow.WY.p$wateryear, na.rm = TRUE)
# not significant


# calculate differences over time
100-(pflow.WY.p$predict.Time[length(pflow.WY.p$predict.Time)]/pflow.WY.p$predict.Time[1]*100)
# diff in last 50 years - same as above, both are less than 50 years

# prepare data for plotting
pflow.WY.pp <- pflow.WY.p %>% rename(Year = wateryear,  Discharge = will_dis) %>% 
  complete(Year = seq(1976, 2020, by = 1)) # add gaps for missing years
```


```{r riverland}
# remove years with too few data
rflow.WY.r <- rflow.WY %>% filter(nFd >= 347)
rmflow.WY.r <- rmflow.WY %>% filter(nFd >= 6)
# combine monthly and daily data
rflow.WY.all <- left_join(rmflow.WY.r,rflow.WY.r,  by = "wateryear") %>% select(wateryear, Flows_dis.x, Flows_dis.y)
# join the columns
rflow.WY.all <- rflow.WY.all %>% 
  mutate(new = case_when(!is.na(Flows_dis.y) ~ Flows_dis.y, TRUE ~ Flows_dis.x),
         new.l = log(new + 1))

M1 <- glm(new.l ~ wateryear, data = rflow.WY.all)

# assumptions
par(mfrow = c(2, 2))
plot(M1) # use log
par(mfrow = c(1, 1)) 
ggplot(rflow.WY.all, aes(wateryear, new.l), na.rm = TRUE) +
  geom_point() 
# roughly linear


# predict
rflow.WY.all$predict.Time.l <- predict(M1, rflow.WY.all)
# create a variable that is predicted log booligal discharge over time
rflow.WY.all$predict.Time <- exp(rflow.WY.all$predict.Time.l)-1
plot(rflow.WY.all$new ~ rflow.WY.all$predict.Time)
plot(rflow.WY.all$predict.Time ~ rflow.WY.all$wateryear)


# plot
ggemmeans.dis <- ggemmeans(M1, terms = c("wateryear[all]"))

plot(ggemmeans.dis, colors = "hero", ci = T, add.data = TRUE, jitter = 0, show.y.title = T, show.title = FALSE) + labs(x="Year",y="log Flows to SA discharge (ML/day)",title="")


# test
summary(M1)
cor.test(rflow.WY.all$new, rflow.WY.all$wateryear)


# calculate differences over time
100-(rflow.WY.all$predict.Time[length(rflow.WY.all$predict.Time)]/rflow.WY.all$predict.Time[1]*100)
# diff in last 50 years - since 1970, which is row 69
100-(rflow.WY.all$predict.Time[length(rflow.WY.all$predict.Time)]/rflow.WY.all$predict.Time[69]*100)

# prepare data for plotting
rflow.WY.rp <- rflow.WY.all %>% rename(Year = wateryear,  Discharge = new) %>% 
  complete(Year = seq(1976, 2020, by = 1)) # add gaps for missing years
```


```{r yantabulla}
# remove years with too few data
yflow.WY.y <- yflow.WY %>% filter(nTd >= 347)

M1 <- glm(l.Tur_dis ~ wateryear, data = yflow.WY.y)

# assumptions
par(mfrow = c(2, 2))
plot(M1) # use log
par(mfrow = c(1, 1)) 
ggplot(yflow.WY.y, aes(wateryear, l.Tur_dis), na.rm = TRUE) +
  geom_point() 
# no obvious non-linear trends


# predict
yflow.WY.y$predict.Time.l <- predict(M1, yflow.WY.y)
# create a variable that is predicted log turra discharge over time
yflow.WY.y$predict.Time <- exp(yflow.WY.y$predict.Time.l)-1
plot(yflow.WY.y$Tur_dis ~ yflow.WY.y$predict.Time)
plot(yflow.WY.y$predict.Time ~ yflow.WY.y$wateryear)


# plot
ggemmeans.dis <- ggemmeans(M1, terms = c("wateryear[all]"))

plot(ggemmeans.dis, colors = "hero", ci = T, add.data = TRUE, jitter = 0, show.y.title = T, show.title = FALSE) + labs(x="Year",y="log Turra discharge (ML/day)",title="")


# test
summary(M1)
cor.test(yflow.WY.y$l.Tur_dis, yflow.WY.y$wateryear)
# not significant


# calculate differences over time
100-(yflow.WY.y$predict.Time[length(yflow.WY.y$predict.Time)]/yflow.WY.y$predict.Time[1]*100)


# prepare data for plotting
yflow.WY.yp <- yflow.WY.y %>% rename(Year = wateryear,  Discharge = Tur_dis) %>% 
  complete(Year = seq(1994, 2020, by = 1)) # add gaps for missing years
```


```{r cumbung inundation}
M1 <- glm(predict ~ year, data = inun)

# assumptions
par(mfrow = c(2, 2))
plot(M1) # looks fine untransformed 
par(mfrow = c(1, 1)) 
ggplot(inun, aes(year, predict), na.rm = TRUE) +
  geom_point() 
# no obvious non-linear trends


# predict
inun$predict.Time.l <- predict(M1, inun)
# create a variable that is predicted inundation area over time


# plot
ggemmeans.dis <- ggemmeans(M1, terms = c("year[all]"))

plot(ggemmeans.dis, colors = "hero", ci = T, add.data = TRUE, jitter = 0, show.y.title = T, show.title = FALSE) + labs(x="Year",y="Predicted Inundation area (km2)",title="")


# test
summary(M1)
cor.test(inun$predict, inun$year)


# calculate differences over time
100-(inun$predict.Time[length(inun$predict.Time)]/inun$predict.Time[1]*100)
# since 1970 (50 years ago)
100-(inun$predict.Time[length(inun$predict.Time)]/inun$predict.Time[22]*100)

#exp(coef(year))
```

# Extract flood frequency data from water level using bankfull levels

```{r}

```



# Flood time series analysis - generalised linear models


```{r cumbung}
# remove years with too few data
cflow.WY.c <- cflow.WY %>% filter(nBd >= 347)

M1 <- glm(predict ~ wateryear, data = cflow.WY.c)

# assumptions
par(mfrow = c(2, 2))
plot(M1) # OK
par(mfrow = c(1, 1)) 
ggplot(cflow.WY.c, aes(wateryear, predict), na.rm = TRUE) +
  geom_point() 
# no obvious non-linear trends


# predict
cflow.WY.c$predict.Time <- predict(M1, cflow.WY.c)
# create a variable that is predicted four mile discharge over time


# plot
ggemmeans.dis <- ggemmeans(M1, terms = c("wateryear[all]"))

plot(ggemmeans.dis, colors = "hero", ci = T, add.data = TRUE, jitter = 0, show.y.title = T, show.title = FALSE) + labs(x="Year",y="Predicted Four mile weir discharge (ML/day)",title="")


# test
summary(M1)
cor.test(cflow.WY.c$predict, cflow.WY.c$wateryear)


# calculate differences over time
100-(cflow.WY.c$predict.Time[length(cflow.WY.c$predict.Time)]/cflow.WY.c$predict.Time[1]*100)
```



```{r booligal}
# remove years with too few data
cflow.WY.b <- cflow.WY %>% filter(nBd >= 347)

M1 <- glm(l.Bool_dis ~ wateryear, data = cflow.WY.b)

# assumptions
par(mfrow = c(2, 2))
plot(M1) # use log
par(mfrow = c(1, 1)) 
ggplot(cflow.WY.b, aes(wateryear, l.Bool_dis), na.rm = TRUE) +
  geom_point() 
# no obvious non-linear trends


# predict
cflow.WY.b$predict.Time.l <- predict(M1, cflow.WY.b)
# create a variable that is predicted log booligal discharge over time
cflow.WY.b$predict.Time <- exp(cflow.WY.b$predict.Time.l)-1
plot(cflow.WY.b$Bool_dis ~ cflow.WY.b$predict.Time)
plot(cflow.WY.b$predict.Time ~ cflow.WY.b$wateryear)


# plot
ggemmeans.dis <- ggemmeans(M1, terms = c("wateryear[all]"))

plot(ggemmeans.dis, colors = "hero", ci = T, add.data = TRUE, jitter = 0, show.y.title = T, show.title = FALSE) + labs(x="Year",y="log Booligal discharge (ML/day)",title="")


# test
summary(M1)
cor.test(cflow.WY.b$l.Bool_dis, cflow.WY.b$wateryear)


# calculate differences over time
100-(cflow.WY.b$predict.Time[length(cflow.WY.b$predict.Time)]/cflow.WY.b$predict.Time[1]*100)
```



```{r gwydir}
# remove years with too few data
gflow.WY.g <- gflow.WY %>% filter(nYd >= 347)

M1 <- glm(l.Yarr_dis ~ wateryear, data = gflow.WY.g)

# assumptions
par(mfrow = c(2, 2))
plot(M1) # use log
par(mfrow = c(1, 1)) 
ggplot(gflow.WY.g, aes(wateryear, l.Yarr_dis), na.rm = TRUE) +
  geom_point() 
# no obvious non-linear trends


# predict
gflow.WY.g$predict.Time.l <- predict(M1, gflow.WY.g)
# create a variable that is predicted log gwydir discharge over time
gflow.WY.g$predict.Time <- exp(gflow.WY.g$predict.Time.l)-1
plot(gflow.WY.g$Yarr_dis ~ gflow.WY.g$predict.Time)
plot(gflow.WY.g$predict.Time ~ gflow.WY.g$wateryear)


# plot
ggemmeans.dis <- ggemmeans(M1, terms = c("wateryear[all]"))

plot(ggemmeans.dis, colors = "hero", ci = T, add.data = TRUE, jitter = 0, show.y.title = T, show.title = FALSE) + labs(x="Year",y="log Gwydir discharge (ML/day)",title="")


# test
summary(M1)
cor.test(gflow.WY.g$l.Yarr_dis, gflow.WY.g$wateryear)


# calculate differences over time
100-(gflow.WY.g$predict.Time[length(gflow.WY.g$predict.Time)]/gflow.WY.g$predict.Time[1]*100)
```


```{r hattah}
# remove years with too few data
hflow.WY.h <- hflow.WY %>% filter(nBd >= 347)

M1 <- glm(predict ~ wateryear, data = hflow.WY.h)

# assumptions
par(mfrow = c(2, 2))
plot(M1) # non-transformed data looks ok
par(mfrow = c(1, 1)) 
ggplot(hflow.WY.h, aes(wateryear, predict), na.rm = TRUE) +
  geom_point() 
# no obvious non-linear trends


# predict
hflow.WY.h$predict.Time <- predict(M1, hflow.WY.h)
# create a variable that is predicted hattah discharge over time


# plot
ggemmeans.dis <- ggemmeans(M1, terms = c("wateryear[all]"))

plot(ggemmeans.dis, colors = "hero", ci = T, add.data = TRUE, jitter = 0, show.y.title = T, show.title = FALSE) + labs(x="Year",y="predicted Colignan discharge (ML/day)",title="")


# test
summary(M1)
cor.test(hflow.WY.h$predict, hflow.WY.h$wateryear)


# calculate differences over time
100-(hflow.WY.h$predict.Time[length(hflow.WY.h$predict.Time)]/hflow.WY.h$predict.Time[1]*100)
```


```{r lowbidgee}
# remove years with too few data
lflow.WY.l <- lflow.WY %>% filter(nHd >= 347)

M1 <- glm(predict.l ~ wateryear, data = lflow.WY.l)

# assumptions
par(mfrow = c(2, 2))
plot(M1) # use log
par(mfrow = c(1, 1)) 
ggplot(cflow.WY, aes(wateryear, predict), na.rm = TRUE) +
  geom_point() 
# no obvious non-linear trends


# predict
lflow.WY.l$predict.Time.l <- predict(M1, lflow.WY.l)
# create a variable that is predicted log booligal discharge over time
lflow.WY.l$predict.Time <- exp(lflow.WY.l$predict.Time.l)-1
plot(lflow.WY.l$predict ~ lflow.WY.l$predict.Time)
plot(lflow.WY.l$predict.Time ~ lflow.WY.l$wateryear)


# plot
ggemmeans.dis <- ggemmeans(M1, terms = c("wateryear[all]"))

plot(ggemmeans.dis, colors = "hero", ci = T, add.data = TRUE, jitter = 0, show.y.title = T, show.title = FALSE) + labs(x="Year",y="log predicted Redbank discharge (ML/day)",title="")


# test
summary(M1)
cor.test(lflow.WY.l$predict.l, lflow.WY.l$wateryear)


# calculate differences over time
100-(lflow.WY.l$predict.Time[length(lflow.WY.l$predict.Time)]/lflow.WY.l$predict.Time[1]*100)
```


```{r macquarie}
# remove years with too few data
mflow.WY.m <- mflow.WY %>% filter(nOd >= 347)

M1 <- glm(l.ox_dis ~ wateryear, data = mflow.WY.m)

# assumptions
par(mfrow = c(2, 2))
plot(M1) #  log looks somewhat better?
par(mfrow = c(1, 1)) 
ggplot(mflow.WY.m, aes(wateryear, l.ox_dis), na.rm = TRUE) +
  geom_point() 
# without transformation, the trend looks somewhat non-linear, decreases massively from 1955 to 1980 then decreases more slowly
# log is better


# predict
mflow.WY.m$predict.Time.l <- predict(M1, mflow.WY.m)
# create a variable that is predicted log booligal discharge over time
mflow.WY.m$predict.Time <- exp(mflow.WY.m$predict.Time.l)-1
plot(mflow.WY.m$ox_dis ~ mflow.WY.m$predict.Time)
plot(mflow.WY.m$predict.Time ~ mflow.WY.m$wateryear)


# plot
ggemmeans.dis <- ggemmeans(M1, terms = c("wateryear[all]"))

plot(ggemmeans.dis, colors = "hero", ci = T, add.data = TRUE, jitter = 0, show.y.title = T, show.title = FALSE) + labs(x="Year",y="Log Oxley discharge (ML/day)",title="")


# test
summary(M1)
cor.test(mflow.WY.m$l.ox_dis, mflow.WY.m$wateryear)


# calculate differences over time
100-(mflow.WY.m$predict.Time[length(mflow.WY.m$predict.Time)]/mflow.WY.m$predict.Time[1]*100)
```


```{r paroo}
# remove years with too few data
pflow.WY.p <- pflow.WY %>% filter(nWd >= 347)

M1 <- glm(l.will_dis ~ wateryear, data = pflow.WY.p)

# assumptions
par(mfrow = c(2, 2))
plot(M1) # log is better
par(mfrow = c(1, 1)) 
ggplot(pflow.WY.p, aes(wateryear, will_dis), na.rm = TRUE) +
  geom_point() 
# no obvious non-linear trends


# predict
pflow.WY.p$predict.Time.l <- predict(M1, pflow.WY.p)
# create a variable that is predicted log booligal discharge over time
pflow.WY.p$predict.Time <- exp(pflow.WY.p$predict.Time.l)-1
plot(pflow.WY.p$will_dis ~ pflow.WY.p$predict.Time)
plot(pflow.WY.p$predict.Time ~ pflow.WY.p$wateryear)


# plot
ggemmeans.dis <- ggemmeans(M1, terms = c("wateryear[all]"))

plot(ggemmeans.dis, colors = "hero", ci = T, add.data = TRUE, jitter = 0, show.y.title = T, show.title = FALSE) + labs(x="Year",y="log Willara Xing discharge (ML/day)",title="")


# test
summary(M1)
cor.test(pflow.WY.p$l.Bool_dis, pflow.WY.p$wateryear)
# not significant


# calculate differences over time
100-(pflow.WY.p$predict.Time[length(pflow.WY.p$predict.Time)]/pflow.WY.p$predict.Time[1]*100)
```


```{r riverland}
# remove years with too few data
rflow.WY.r <- rflow.WY %>% filter(nFd >= 347)

M1 <- glm(l.Flows_dis ~ wateryear, data = rflow.WY.r)

# assumptions
par(mfrow = c(2, 2))
plot(M1) # use log
par(mfrow = c(1, 1)) 
ggplot(rflow.WY.r, aes(wateryear, l.Flows_dis), na.rm = TRUE) +
  geom_point() 
# no obvious non-linear trends


# predict
rflow.WY.r$predict.Time.l <- predict(M1, rflow.WY.r)
# create a variable that is predicted log booligal discharge over time
rflow.WY.r$predict.Time <- exp(rflow.WY.r$predict.Time.l)-1
plot(rflow.WY.r$Flows_dis ~ rflow.WY.r$predict.Time)
plot(rflow.WY.r$predict.Time ~ rflow.WY.r$wateryear)


# plot
ggemmeans.dis <- ggemmeans(M1, terms = c("wateryear[all]"))

plot(ggemmeans.dis, colors = "hero", ci = T, add.data = TRUE, jitter = 0, show.y.title = T, show.title = FALSE) + labs(x="Year",y="log Flows to SA discharge (ML/day)",title="")


# test
summary(M1)
cor.test(rflow.WY.r$l.Flows_dis, rflow.WY.r$wateryear)


# calculate differences over time
100-(rflow.WY.r$predict.Time[length(rflow.WY.r$predict.Time)]/rflow.WY.r$predict.Time[1]*100)
```


```{r yantabulla}
# remove years with too few data
yflow.WY.y <- yflow.WY %>% filter(nTd >= 347)

M1 <- glm(l.Tur_dis ~ wateryear, data = yflow.WY.y)

# assumptions
par(mfrow = c(2, 2))
plot(M1) # use log
par(mfrow = c(1, 1)) 
ggplot(yflow.WY.y, aes(wateryear, l.Tur_dis), na.rm = TRUE) +
  geom_point() 
# no obvious non-linear trends


# predict
yflow.WY.y$predict.Time.l <- predict(M1, yflow.WY.y)
# create a variable that is predicted log booligal discharge over time
yflow.WY.y$predict.Time <- exp(yflow.WY.y$predict.Time.l)-1
plot(yflow.WY.y$Tur_dis ~ yflow.WY.y$predict.Time)
plot(yflow.WY.y$predict.Time ~ yflow.WY.y$wateryear)


# plot
ggemmeans.dis <- ggemmeans(M1, terms = c("wateryear[all]"))

plot(ggemmeans.dis, colors = "hero", ci = T, add.data = TRUE, jitter = 0, show.y.title = T, show.title = FALSE) + labs(x="Year",y="log Turra discharge (ML/day)",title="")


# test
summary(M1)
cor.test(yflow.WY.y$l.Tur_dis, yflow.WY.y$wateryear)
# not significant


# calculate differences over time
100-(yflow.WY.y$predict.Time[length(yflow.WY.y$predict.Time)]/yflow.WY.y$predict.Time[1]*100)

#exp(coef(wateryear))
```



# Time series analysis - plot

```{r time series plot functions}
plot_dis <- function(dataset, title){
  ggplot(dataset, aes(x = Year, y = Discharge)) +
  geom_line()  + 
  stat_smooth(method = lm) +
  labs(x="",y="",title=title) +
  theme_bw() +
  geom_line(size = 1.2) +
  #geom_col(aes(x = Year, y = missing_dis), width = 31, fill = "#f0f0f0") + # not working yet
  theme(text = element_text(size = 32), # change font size
        axis.text = element_text(size = 28),
        plot.margin = margin(10,20,10,10), # give extra space on RHS
        plot.title = element_text(hjust = 0.5)) # centre title
}
```

```{r make discharge plots}
bd <- plot_dis(cflow.WY.bp, "Booligal Wetlands*") +
  scale_x_continuous(breaks = c(1945,1970,2000,2020))
bd
cd <- plot_dis(cflow.WY.cp, "Great Cumbung Swamp*") + # need to add data!
  scale_x_continuous(breaks = c(1945,1970,2000,2020))  +
  labs(x="",y="Discharge (ML/day)") +
  theme(axis.title = element_text(size = 60))
cd
gd <- plot_dis(gflow.WY.gp, "Gwydir Wetlands*") +
  scale_x_continuous(breaks = c(1930,1960,1990,2020))
gd
hd <- plot_dis(hflow.WY.hp, "Hattah Lakes*") +
  scale_x_continuous(breaks = c(1910,1950,1990,2020)) +
  labs(x="Year",y="") +
  theme(axis.title = element_text(size = 60))
hd
ld <- plot_dis(lflow.WY.lp, "Lowbidgee Wetlands*") +
  scale_x_continuous(breaks = c(1910,1950,1990,2020))
ld
md <- plot_dis(mflow.WY.mp, "Macquarie Marshes*") +
  scale_x_continuous(breaks = c(1960,1980,2000,2018)) +
  labs(x="",y="Discharge (ML/day)") +
  theme(axis.title = element_text(size = 60))
md
pd <- plot_dis(pflow.WY.pp, "Paroo Overflow Swamps") +
  labs(x="Year",y="Discharge (ML/day)") +
  theme(text = element_text(size = 30),
        axis.title = element_text(size = 60))
pd
rd <- plot_dis(rflow.WY.rp, "Murray Riverlands*") +
  labs(x="Year",y="") +
  theme(axis.title = element_text(size = 60)) +
  scale_x_continuous(breaks = c(1900,1930,1960,1990,2020))
rd
yd <- plot_dis(yflow.WY.yp, "Yantabulla Swamps") +
  scale_x_continuous(breaks = c(1994,2000,2010,2020))
yd

```

```{r combine discharge plots}
# export at 2100 x 2700 
# add asterisks for significance
# order in terms of significance or standard order

flow <- ggarrange(cd, bd, ld, md, gd, yd, pd, rd, hd,
          ncol = 3, nrow = 3)
flow


# if only one title on each axis

#annotate_figure(bird, bottom = text_grob("Year                          Year                         Year",  size = 60, just = "bottom"), left = text_grob("Discharge (ML/day)         Discharge (ML/day)         Discharge (ML/day)", size = 60, rot = 90)) # rotate 90 degrees
# just = "bottom" helps text not be cut off
```

