---
title: "trend_analysis"
author: "Dana Lanceman"
date: "10/1/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Prepare data

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = 'C:/Users/danal/OneDrive - UNSW/Desktop/Honours_UNSW/Assessment/cumbung_assessment')
```


```{r load packages}
library(readxl)
library(tidyverse)
library(Kendall)
library(ggeffects)
library(zoo)
library(ggpubr)
library(lubridate)
library(mgcv)
library(tidymv)

options(java.parameters = "-Xmx8000m") 
# the above increases the number of MB of ram that java can heap in order for xlsx package to run properly and save large files into an excel file. This line needs to be run before the xlsx package is loaded in. The number 8000 (MB) can be changed to whatever number you require.
library(xlsx)
```


```{r get daily flow/flood data}
# guess_max lets excel use more rows to decide the data type of the data (missing data makes it think the data type is logical otherwise)
# add 'month', 'wateryear' and 'year' columns
# wateryear is July to June and the year associated with each water year is the later year; ie Jan to June 2021 belong to 2021 and July to Dec 2021 belong to 2022

cflow <- read_excel("raw/correlate_flow_flood.xlsx",sheet = "cumbung", guess_max = 30000) %>% mutate(month = month(Date),
                  wateryear = case_when(month <= 6 ~ year(Date),
                                        TRUE ~ (year(Date) + 1)),
                  year = year(Date),
                  ym = make_date(wateryear, month)) %>% 
  select(Date, Bool_dis, Bool_lev, Four_lev, Four_dis, month, wateryear, year, ym)
# check all data types are numerical (except ym)
str(cflow)
```



```{r fill gaps and log transform data}
cflow <- cflow %>% mutate(Bool_dis = na.approx(Bool_dis, maxgap = 10, na.rm = TRUE),
                          Bool_lev = na.approx(Bool_lev, maxgap = 10, na.rm = TRUE),
                          Four_dis = na.approx(Four_dis, maxgap = 10, na.rm = TRUE),
                          Four_lev = na.approx(Four_lev, maxgap = 10, na.rm = TRUE),
                          l.Four_dis = log(Four_dis + 1),
                          l.Four_lev = log(Four_lev + 1),
                          l.Bool_dis = log(Bool_dis + 1),
                          l.Bool_lev = log(Bool_lev + 1))
```


```{r Make binary flood presence/absence variable from water level data and bankfull levels}
cflow <- cflow %>% mutate(floodc = case_when(Four_lev >= 1.82 ~ 1, TRUE ~ 0))
```



```{r summarise water level data by month and log transform}
cflow.m <- cflow %>% group_by(ym) %>% summarise(nB = sum(!is.na(Bool_lev)), nF = sum(!is.na(Four_lev)), Bool_levA = mean(Bool_lev, na.rm = TRUE), Four_levA = mean(Four_lev, na.rm = TRUE), Bool_levM = max(Bool_lev, na.rm = TRUE), Four_levM = max(Four_lev, na.rm = TRUE), Bool_disA = mean(Bool_dis, na.rm = TRUE), floodc = sum(floodc)) %>% 
  mutate(l.Four_lev = log(Four_levA + 1),
        l.Bool_lev = log(Bool_levA + 1),
        l.Four_levM = log(Four_levM + 1),
        l.Bool_levM = log(Bool_levM + 1),
        l.Bool_dis = log(Bool_disA + 1))

# warnings are because Four Mile Weir dataset is much shorter than Booligal so contains missing values for all dates before 2009
```


```{r summarise data per year and log transform annual data}
cflow.WY <- cflow %>% group_by(wateryear) %>% summarise(nBl = sum(!is.na(Bool_lev)), nFl = sum(!is.na(Four_lev)), nBd = sum(!is.na(Bool_dis)), nFd = sum(!is.na(Four_dis)), Bool_dis = mean(Bool_dis, na.rm = TRUE), Four_dis = mean(Four_dis, na.rm = TRUE), Bool_lev = mean(Bool_lev, na.rm = TRUE), Four_lev = mean(Four_lev, na.rm = TRUE), floodc = sum(floodc)) %>% 
  mutate(l.Four_dis = log(Four_dis + 1),
        l.Four_lev = log(Four_lev + 1),
        l.Bool_dis = log(Bool_dis + 1),
        l.Bool_lev = log(Bool_lev + 1))
```

```{r Create an inundation vs booligal discharge dataframe}
#Add Booligal 95% complete years to a new dataframe
inun <- tibble(.rows = 70) %>% 
  add_column(Bool_dis = cflow.WY %>% filter(nBd >= 347) %>% .$Bool_dis,
             year = cflow.WY %>% filter(nBd >= 347) %>% .$wateryear)

# load annual inundation data 
cinun <- read_excel("raw/correlate_flow_flood.xlsx",sheet = "cumbung_inun") 

# combine these 
inun <- merge(inun, cinun, by = 'year') %>% 
  mutate(l.Bool_dis = log(Bool_dis + 1),
         l.area.km2 = log(area.km2 + 1))
```


# Data correlations

```{r GLM for correlating annual discharge}
# generalised linear model
# filter data within the model so that the actual dataset remains unchanged
M1 <- glm(l.Four_dis ~ l.Bool_dis, data=cflow.WY %>% filter(nFd >= 347, nBd >= 347), na.action = na.omit)

# check assumptions
par(mfrow = c(2, 2))
plot(M1) # top left test for equal variances, top right normality 
par(mfrow = c(1, 1))
ggplot(cflow.WY %>% filter(nFd >= 347, nBd >= 347), aes(l.Bool_dis, l.Four_dis), na.rm = TRUE) + geom_point() 

# very small dataset, assumptions all ok

summary(M1)
cor.test(cflow.WY$Bool_dis, cflow.WY$Four_dis)

# predict
cflow.WY$predict.l <- predict(M1, cflow.WY)
# create a variable that is predicted log colignan dis
cflow.WY$predict <- exp(cflow.WY$predict.l) - 1
# back transform 
plot(cflow.WY$Four_dis ~ cflow.WY$predict)


ggemmeans.Dis <- ggemmeans(M1, terms = c("l.Bool_dis[all]"))

options(scipen = 1)

c.flow <- plot(ggemmeans.Dis, dot.alpha = 1,  colors = "hero", ci = T, add.data = TRUE, jitter = 0, show.y.title = T, show.title = FALSE) + 
  labs(x="Booligal flows (ML/day)",y="Four Mile flows (ML/day)",title="Great Cumbung Swamp") +
    geom_line(size = 1) +
  theme_bw() +
  #geom_col(aes(x = Year, y = missing_dis), width = 31, fill = "#f0f0f0") + # not working yet
  theme(plot.title = element_text(size = 35,hjust = 0.5, face = "bold"),
        text = element_text(size = 38), # change font size
        axis.text = element_text(size = 32),
        plot.margin = margin(10,20,10,10), # give extra space on RHS
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()) + # centre title
  scale_y_continuous(breaks = pretty(cflow.WY$l.Four_dis), 
                     labels = signif(exp(pretty(cflow.WY$l.Four_dis)),1)) +
  scale_x_continuous(breaks = pretty(cflow.WY$l.Bool_dis), 
                     labels = signif(exp(pretty(cflow.WY$l.Bool_dis)),1))
c.flow


# make a new column that is a combination of the old and new data
# if the raw data aren't an NA, it uses the raw data. otherwise, it uses the predicted data
# also make a new column which is nNd (number of data points in each for the new data)
cflow.WY <- cflow.WY %>% mutate(new = case_when(!is.na(Four_dis) ~ Four_dis, TRUE ~ predict),
                                nNd = case_when(nFd != 0 ~ nFd, TRUE ~ nBd))

# check against just predicted data
plot(cflow.WY$new ~ cflow.WY$predict)
# as expected, almost all a 1:1 ratio as most data are from booligal, but some deviation due to four mile data - good
```



```{r Cumbung Inundation area vs Booligal discharge}
# generalised linear model
M1 <- glm(l.area.km2 ~ l.Bool_dis, data=inun)
summary(M1)
cor.test(inun$l.Bool_dis, inun$l.area.km2)

 # check assumptions
par(mfrow = c(2, 2))
plot(M1) # top left test for equal variances, top right normality
# better with log transformed data, acceptable
par(mfrow = c(1, 1))
ggplot(inun, aes(l.Bool_dis, l.area.km2)) +
  geom_point() # test for linear relationship
# log-log relationship looks roughly linear

# predict
inun$predict.l <- predict(M1, inun)
# create a variable that is predicted log inundation area
inun$predict <- exp(inun$predict.l)
# back transform 
plot(inun$area.km2 ~ inun$predict)



# visualise relationship between original and predicted inundation
ggemmeans.Inun <- ggemmeans(M1, terms = c("l.Bool_dis[all]")) # this must specify the second term listed in glm

c.inun <- plot(ggemmeans.Inun, colors = "hero", ci = T, add.data = TRUE, jitter = 0, show.y.title = T, show.title = FALSE) + labs(x="Booligal Flows (log)",y="Cumbung Inundation Area (log)",title="") +
    geom_line(size = 1) +
  theme_bw() +
  #geom_col(aes(x = Year, y = missing_dis), width = 31, fill = "#f0f0f0") + # not working yet
  theme(text = element_text(size = 22), # change font size
        axis.text = element_text(size = 24),
        plot.margin = margin(10,20,10,10), # give extra space on RHS
        plot.title = element_text(hjust = 0.5),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()) # centre title
c.inun

# export
#write.csv(inun, "output/cbool_pred_inun.csv")
```


Using GAMs instead of GLMs for flood data as GLMs don't fit well and assumptions aren't acceptable.

```{r Generalised additive models - monthly flood data}
# modelling monthly data
m <- gam(l.Four_lev ~   s(l.Bool_lev, k=3), data=cflow.m %>% filter(nB >= 29, nF >= 29) %>%  subset(Bool_levA > 0, Four_levA > 0))
# k = 3 is 3rd order polynomial
# only include months with >95% data and water levels above zero

summary(m)

# check normality and equal variances (linearity is not necessary)
par(mfrow = c(2, 2))
gam.check(m) # residual plots for GAMs
# normality seems ok, a bit strange but doesn't improve much with transformation
par(mfrow = c(1, 1))

# predict
cflow.m$predict.l <- predict(m,newdata = cflow.m)
cflow.m$predict <- exp(cflow.m$predict.l) - 1


# plot
ggpredict <- ggemmeans(m, terms = c("l.Bool_lev [all]"))
c.flood <- ggplot(ggpredict,aes(x = x, y = predicted))+
        geom_ribbon(aes(ymin =conf.low, ymax = conf.high),alpha=0.2) +
        geom_line(size=0.5)+
        geom_point(data=cflow.m %>% filter(nB >= 29, nF >= 29) %>%  subset(Bool_levA > 0, Four_levA > 0), aes(x=l.Bool_lev, y=l.Four_lev)) +
        theme_bw() +
        theme(panel.grid.major = element_blank(),
              panel.grid.minor=element_blank(),
              text = element_text(size = 38),
              axis.title = element_text(size = 38),
              plot.title = element_text(hjust = 0.5))  + 
        #scale_y_continuous(limits = c(0,1)) + 
        scale_x_continuous(labels = scales::comma)+
        labs(x="Booligal water level (m)", y="Four Mile water level (m)", title = "") +
  scale_y_continuous(breaks = pretty(cflow.m$l.Four_lev, n = 3), 
                     labels = signif(exp(pretty(cflow.m$l.Four_lev, n = 3)),1)) +
  scale_x_continuous(breaks = pretty(cflow.m$l.Bool_lev, n = 3), 
                     labels = signif(exp(pretty(cflow.m$l.Bool_lev, n = 3)),1))
c.flood

```


# Correlate monthly water levels with monthly number of flood days

```{r Cumbung}
# plot relationship between raw water level data and flood days
ggplot(cflow.m, aes(Four_levA, floodc)) +
  geom_point()
# this is a threshold relationship, where there are no flood days until water level > 0.8m


# make a model, only using water level data > 0.8m
m <- gam(floodc ~ s(l.Four_lev, k=3), data=cflow.m %>% filter(nF >= 29, Four_levA >= 0.8))
summary(m)


# check normality and equal variances (linearity is not necessary)
par(mfrow = c(2, 2))
gam.check(m) # residual plots for GAMs
# ok
par(mfrow = c(1, 1))


# plot the modelled relationship
ggpredict <- ggemmeans(m, terms = c("l.Four_lev [all]"))
c.floodday <- ggplot(ggpredict,aes(x = x, y = predicted))+
        geom_ribbon(aes(ymin =conf.low, ymax = conf.high),alpha=0.2) +
        geom_line(size=0.5)+
        geom_point(data=cflow.m %>% filter(nF >= 29, Four_levA >= 0.8), aes(x=l.Four_lev, y=floodc))+
        theme_bw(base_size=16) +
        theme(panel.grid.major = element_blank(),
              panel.grid.minor=element_blank(),
              text = element_text(size = 38),
              axis.title = element_text(size = 38),
              plot.title = element_text(hjust = 0.5))  + 
        #scale_y_continuous(limits = c(0,1)) + 
        scale_x_continuous(labels = scales::comma)+
        labs(x="Four Mile water level (m)", y="Number of flood days", title = "")+
  scale_x_continuous(breaks = pretty(cflow.m$l.Four_lev), 
                     labels = signif(exp(pretty(cflow.m$l.Four_lev)),2)) +
  scale_y_continuous(limits = c(-5, 35))
c.floodday

# predict number of flood days based on predicted four mile water levels

# create a new dataset with the predicted data reassigned to have the same name as water level data in this model
cflow.mp <- cflow.m %>% 
  select(ym, floodc, predict) %>% 
  mutate(l.Four_lev = log(predict + 1)) 
# create a predicted flood days variable by predicting based on this model
cflow.mp$predict.f <- case_when(cflow.mp$predict >= 0.8 ~ predict(m,newdata = cflow.mp), TRUE ~ 0) 
# reassign any negatives to zero (impossible to have a negative number of days)
cflow.mp$predict.f <- case_when(cflow.mp$predict.f < 0 ~ 0, TRUE ~ cflow.mp$predict.f)


# visualise this
plot(cflow.mp$ym, cflow.mp$predict.f)
# visualise relationship between time and flood days with original data
plot(cflow.mp$ym, cflow.mp$floodc)
```


```{r Sum monthly flood days for these datasets to create annual data}
# at this point, only use predicted flood day data, rather than combining with raw data, as predicted data have many more months with 1 flood day (rather than zero), so would make it seem like floods had declined more than they actually have if used in combination with raw data

cflow.mp$Year <- year(cflow.mp$ym)
cflood <- cflow.mp %>%  group_by(Year) %>% summarise(Flood_days = sum(predict.f), nF = sum(!is.na(predict.f))) %>% subset(nF >= 11)  %>% 
  complete(Year = seq(1946, 2020, by = 1)) # add gaps for missing years
# only including years with at least 11 months (95%) of data

```


# Correlating discharge and water level for some gauges used for bankfull analysis

```{r Try plotting daily data}
# untransformed
plot(cflow$Bool_lev, cflow$Bool_dis) # pretty good

# log
plot(cflow$l.Bool_lev, cflow$l.Bool_dis) 
# looks no better or worse with log-transformations

# non-linear though, so use GAMs
```


```{r Try plotting monthly data}
# untransformed
plot(cflow.m$Bool_levA, cflow.m$Bool_disA) # pretty good

# log
plot(cflow.m$l.Bool_lev, cflow.m$l.Bool_dis) 
```


```{r Try plotting annual data}
# untransformed
plot(cflow.WY$Bool_lev, cflow.WY$Bool_dis) # pretty good

# log
plot(cflow.WY$l.Bool_lev, cflow.WY$l.Bool_dis) 

# untransformed look linear, so could potentially use with a GLM
```


Generalised additive models

```{r with monthly data}
# booligal
m <- gam(l.Bool_lev ~   s(l.Bool_dis, k=3), data=cflow.m %>%  filter(Bool_levA > 0, Bool_disA > 0))
m <- gam(Bool_levA ~   s(Bool_disA, k=3), data=cflow.m %>%  filter(Bool_levA > 0, Bool_disA > 0))
# k = 3 is 3rd order polynomial
# only include months with >95% data and water levels above zero

summary(m)

# check normality and equal variances (linearity is not necessary)
par(mfrow = c(2, 2))
gam.check(m) # residual plots for GAMs
par(mfrow = c(1, 1))

# looks better than daily
# log improves normality somewhat but makes other plots look a bit strange
```


Therefore log-transform

```{r Booligal}

m <- gam(l.Bool_lev ~   s(l.Bool_dis, k=4), data=cflow.m %>%  filter(Bool_levA > 0, Bool_disA > 0, nB >= 29, l.Bool_dis > 2))
# k = 4 is 4 knots (captures curve better, except very high values, which aren't of concern anyway)
# only include months with >95% data and water levels above zero
# start with log flow > 2 as few data points are before this

summary(m)

# check normality and equal variances (linearity is not necessary)
par(mfrow = c(2, 2))
gam.check(m) # residual plots for GAMs
par(mfrow = c(1, 1))


# plot
ggpredict <- ggemmeans(m, terms = c("l.Bool_dis [all]"))
b.corr <- ggplot(ggpredict,aes(x = x, y = predicted))+
        geom_ribbon(aes(ymin =conf.low, ymax = conf.high),alpha=0.2) +
        geom_line(size=0.5)+
        geom_point(data=cflow.m %>% filter(nB >= 29, Bool_levA > 0, Bool_disA > 0), aes(x=l.Bool_dis, y=l.Bool_lev)) +
        theme_bw() +
        theme(panel.grid.major = element_blank(),
              panel.grid.minor=element_blank(),
              text = element_text(size = 38),
              axis.title = element_text(size = 38),
              plot.title = element_text(hjust = 0.5))  + 
        #scale_y_continuous(limits = c(0,1)) + 
        scale_x_continuous(labels = scales::comma)+
        labs(x="Booligal flows (log)", y="Booligal water level (log)", title = "")
b.corr



# predict
cflow.m$predictL.l <- case_when(cflow.m$l.Bool_dis <= 2 ~ 0, TRUE ~ predict(m,newdata = cflow.m))
cflow.m$predictL <- exp(cflow.m$predictL.l) - 1
# reassign any negatives to zero (impossible to have a negative number of days)
cflow.m$predictL <- case_when(cflow.m$predictL < 0 ~ 0, TRUE ~ cflow.m$predictL)

# we are interested in 2500ML/day - see what predicted value this correlates to in the below
view(cflow.m)
# 1.97m
```


# Discharge Time series analysis - generalised linear models

```{r cumbung}
# remove years with too few data
cflow.WY.c <- cflow.WY %>% filter(nNd >= 347)

M1 <- glm(new ~ wateryear, data = cflow.WY.c)

# assumptions
par(mfrow = c(2, 2))
plot(M1) # OK
par(mfrow = c(1, 1)) 
ggplot(cflow.WY.c, aes(wateryear, new), na.rm = TRUE) +
  geom_point() 
# no obvious non-linear trends


# predict
cflow.WY.c$predict.Time <- predict(M1, cflow.WY.c)
# create a variable that is predicted four mile discharge over time


# plot
ggemmeans.dis <- ggemmeans(M1, terms = c("wateryear[all]"))

plot(ggemmeans.dis, colors = "hero", ci = T, add.data = TRUE, jitter = 0, show.y.title = T, show.title = FALSE) + labs(x="Year",y="Predicted Four mile weir discharge (ML/day)",title="")


# test
summary(M1)
cor.test(cflow.WY.c$new, cflow.WY.c$wateryear)


# calculate differences over time
100-(cflow.WY.c$predict.Time[length(cflow.WY.c$predict.Time)]/cflow.WY.c$predict.Time[1]*100)
# diff in last 50 years - since 1970, which is row 22
100-(cflow.WY.c$predict.Time[length(cflow.WY.c$predict.Time)]/cflow.WY.c$predict.Time[22]*100)

# average annual % change
(100-(cflow.WY.c$predict.Time[length(cflow.WY.c$predict.Time)]/cflow.WY.c$predict.Time[1]*100))/length(cflow.WY.c$predict.Time)

# prepare data for plotting
cflow.WY.cp <- cflow.WY.c %>% rename(Year = wateryear,  Discharge = new) %>% 
  complete(Year = seq(1946, 2020, by = 1)) # add gaps for missing years
```


```{r cumbung inundation}
M1 <- glm(predict ~ year, data = inun)

# assumptions
par(mfrow = c(2, 2))
plot(M1) # looks fine untransformed 
par(mfrow = c(1, 1)) 
ggplot(inun, aes(year, predict), na.rm = TRUE) +
  geom_point() 
# no obvious non-linear trends


# predict
inun$predict.Time <- predict(M1, inun)
# create a variable that is predicted inundation area over time


# plot
ggemmeans.dis <- ggemmeans(M1, terms = c("year[all]"))

plot(ggemmeans.dis, colors = "hero", ci = T, add.data = TRUE, jitter = 0, show.y.title = T, show.title = FALSE) + labs(x="Year",y="Predicted Inundation area (km2)",title="")


# test
summary(M1)
cor.test(inun$predict, inun$year)


# calculate differences over time
100-(inun$predict.Time[length(inun$predict.Time)]/inun$predict.Time[1]*100)
# since 1970 (50 years ago)
100-(inun$predict.Time[length(inun$predict.Time)]/inun$predict.Time[22]*100)

#exp(coef(year))

# prepare data for plotting
inun.p <- inun %>% rename(Year = year,  Discharge = predict) %>% 
  complete(Year = seq(1946, 2020, by = 1)) # add gaps for missing years
```

# Flood time series analysis - generalised linear models


```{r cumbung}
M1 <- glm(Flood_days ~ Year, data = cflood)

# assumptions
par(mfrow = c(2, 2))
plot(M1) # OK, log makes it much worse
par(mfrow = c(1, 1)) 
ggplot(cflood, aes(Year, Flood_days), na.rm = TRUE) +
  geom_point() 
# no obvious non-linear trends


# predict
cflood$predict.Time <- predict(M1, cflood)
# create a variable that is predicted four mile discharge over time


# plot
ggemmeans.dis <- ggemmeans(M1, terms = c("Year[all]"))

plot(ggemmeans.dis, colors = "hero", ci = T, add.data = TRUE, jitter = 0, show.y.title = T, show.title = FALSE) + labs(x="Year",y="Predicted Four mile weir # flood days",title="")


# test
summary(M1)
cor.test(cflood$Flood_days, cflood$Year)


# calculate differences over time
100-(cflood$predict.Time[length(cflood$predict.Time)]/cflood$predict.Time[1]*100)
# diff in last 50 years - since 1970, which is row 25
100-(cflood$predict.Time[length(cflood$predict.Time)]/cflood$predict.Time[25]*100)

# prepare data for plotting
cflood <- cflood %>% complete(Year = seq(1946, 2020, by = 1)) # add gaps for missing years
```


# Plots

```{r plot functions}
plot_dis <- function(dataset, title){
  ggplot(dataset, aes(x = Year, y = Discharge)) +
    geom_point() +
  stat_smooth(method = lm, aes(ymin = ifelse(..ymin.. < 0, 0, ..ymin..))) + # make sure confidence intervals are not cut off
  labs(x="",y="",title=title) +
  theme_bw() +
  geom_line(size = 1.2) +
  #geom_col(aes(x = Year, y = missing_dis), width = 31, fill = "#f0f0f0") + # not working yet
  theme(text = element_text(size = 32), # change font size
        axis.text = element_text(size = 28),
        plot.margin = margin(10,20,10,10), # give extra space on RHS
        plot.title = element_text(hjust = 0.5)) # centre title
}

plot_flood <- function(dataset, title){
  ggplot(dataset, aes(x = Year, y = Flood_days)) +
    geom_point()  + 
  stat_smooth(method = lm, aes(ymin = ifelse(..ymin.. < 0, 0, ..ymin..))) +
  labs(x="",y="",title=title) +
  theme_bw() +
  geom_line(size = 1.2) +
  #geom_col(aes(x = Year, y = missing_dis), width = 31, fill = "#f0f0f0") + # not working yet
  theme(text = element_text(size = 32), # change font size
        axis.text = element_text(size = 28),
        plot.margin = margin(10,20,10,10), # give extra space on RHS
        plot.title = element_text(hjust = 0.5)) # centre title
}
```

```{r Correlation plots}
# export at 2100 x 2900
corr.plot <- ggarrange(c.flow, c.flood, c.floodday,
          ncol = 1, nrow = 3)
corr.plot
```


```{r make discharge plot}
cd <- plot_dis(cflow.WY.cp, "Great Cumbung Swamp*") + 
  scale_x_continuous(breaks = c(1945,1970,2000,2020))  +
  labs(x="",y="Flow (ML/day)") +
  theme(text = element_text(size = 26),
        axis.title = element_text(size = 60))  +
  scale_y_continuous(limits = c(0, 800))
cd
```

```{r make flood plot}
cf <- plot_flood(cflood, "Great Cumbung Swamp*") + 
  scale_x_continuous(breaks = c(1945,1970,2000,2020))  +
  labs(x="",y="Number of flood days") +
  theme(text = element_text(size = 30),
        axis.title = element_text(size = 60))+
  scale_y_continuous(limits = c(0, 330))
cf
```

```{r Plots for presentation}
# flows
plot_dis(cflow.WY.cp, "") + 
  scale_x_continuous(breaks = c(1945,1970,2000,2020))  +
  labs(x="Year",y="Flow (ML/day)") +
  theme(text = element_text(size = 26),
        axis.title = element_text(size = 35),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank())  +
  scale_y_continuous(limits = c(0, 800))

# floods
plot_flood(cflood, "") + 
  scale_x_continuous(breaks = c(1945,1970,2000,2020))  +
  labs(x="Year",y="Flood days") +
  theme(text = element_text(size = 30),
        axis.title = element_text(size = 35),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank())+
  scale_y_continuous(limits = c(0, 330))

# inundation
ip <- plot_dis(inun.p, "") +
  scale_x_continuous(breaks = c(1946,1970,2000,2020)) +
  scale_y_continuous(limits = c(0, 70))  +
  labs(x="Year",y="Inundation (km2)") +
  theme(text = element_text(size = 30),
        axis.title = element_text(size = 35),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank())
ip
```
