---
title: "trend_analysis"
author: "Dana Lanceman"
date: "10/1/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Prepare data

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = 'C:/Users/danal/OneDrive - UNSW/Desktop/Honours_UNSW/Assessment/cumbung_assessment')
```


```{r load packages}
library(readxl)
library(tidyverse)
library(Kendall)
library(ggeffects)
library(zoo)
library(ggpubr)
library(lubridate)
library(mgcv)
library(tidymv)

options(java.parameters = "-Xmx8000m") 
# the above increases the number of MB of ram that java can heap in order for xlsx package to run properly and save large files into an excel file. This line needs to be run before the xlsx package is loaded in. The number 8000 (MB) can be changed to whatever number you require.
library(xlsx)
```


```{r get daily flow/flood data}
# guess_max lets excel use more rows to decide the data type of the data (missing data makes it think the data type is logical otherwise)
# add a 'wateryear' and 'month' columns

add.date <- function(data){
  data %>% mutate(wateryear = year(Date),
       month = month(Date),
       ym = make_date(wateryear, month)) 
}

cflow <- read_excel("raw/correlate_flow_flood.xlsx",sheet = "cumbung", guess_max = 30000) %>% mutate(wateryear = year(`Date and time`),
       month = month(`Date and time`),
       ym = make_date(wateryear, month))
# check all data types are numerical
str(cflow)

hflow <- read_excel("raw/correlate_flow_flood.xlsx",sheet = "hattah", guess_max = 30000) %>% add.date()
str(hflow)

lflow <- read_excel("raw/correlate_flow_flood.xlsx",sheet = "lowbidgee", guess_max = 30000) %>% add.date()
str(lflow)

yflow <- read_excel("raw/correlate_flow_flood.xlsx",sheet = "yantabulla", guess_max = 30000) %>% add.date() 
str(yflow)

rflow <- read_excel("raw/correlate_flow_flood.xlsx",sheet = "riverland", guess_max = 30000) %>% add.date()
str(rflow)

pflow <- read_excel("raw/correlate_flow_flood.xlsx",sheet = "paroo", guess_max = 30000) %>% add.date()
str(pflow)

gflow <- read_excel("raw/correlate_flow_flood.xlsx",sheet = "gwydir", guess_max = 30000) %>% add.date()
str(gflow)

mflow <- read_excel("raw/correlate_flow_flood.xlsx",sheet = "macquarie", guess_max = 30000) %>% add.date()
str(mflow)


# also get monthly discharge for riverland, which are available long before daily data
rmflow <- read_excel("raw/correlate_flow_flood.xlsx",sheet = "river_monthly", guess_max = 30000) %>% add.date()
str(rmflow)
```



```{r fill gaps and log transform data}
cflow <- cflow %>% mutate(Bool_dis = na.approx(Bool_dis, maxgap = 10, na.rm = TRUE),
                          Bool_lev = na.approx(Bool_lev, maxgap = 10, na.rm = TRUE),
                          Corr_dis = na.approx(Corr_dis, maxgap = 10, na.rm = TRUE),
                          Corr_lev = na.approx(Corr_lev, maxgap = 10, na.rm = TRUE),
                          Four_dis = na.approx(Four_dis, maxgap = 10, na.rm = TRUE),
                          Four_lev = na.approx(Four_lev, maxgap = 10, na.rm = TRUE),
                          l.Four_dis = log(Four_dis + 1),
                          l.Four_lev = log(Four_lev + 1),
                          l.Corr_dis = log(Corr_dis + 1),
                          l.Corr_lev = log(Corr_lev + 1),
                          l.Bool_dis = log(Bool_dis + 1),
                          l.Bool_lev = log(Bool_lev + 1))

hflow <- hflow %>% mutate(Bar_lev = na.approx(Bar_lev, maxgap = 10, na.rm = TRUE),
                          Bar_dis = na.approx(Bar_dis, maxgap = 10, na.rm = TRUE),
                          Col_lev = na.approx(Col_lev, maxgap = 10, na.rm = TRUE),
                          Col_dis = na.approx(Col_dis, maxgap = 10, na.rm = TRUE),
                          l.Bar_lev = log(Bar_lev + 1),
                          l.Bar_dis = log(Bar_dis + 1),
                          l.Col_lev = log(Col_lev + 1),
                          l.Col_dis = log(Col_dis + 1))

lflow <- lflow %>% mutate(Hay_dis = na.approx(Hay_dis, maxgap = 10, na.rm = TRUE),
                          Red_lev = na.approx(Red_lev, maxgap = 10, na.rm = TRUE),
                          Red_dis = na.approx(Red_dis, maxgap = 10, na.rm = TRUE),
                          Hay_lev = na.approx(Hay_lev, maxgap = 10, na.rm = TRUE),
                          Maude_dis = na.approx(Maude_dis, maxgap = 10, na.rm = TRUE),
                          Maude_lev = na.approx(Maude_lev, maxgap = 10, na.rm = TRUE),
                          l.Hay_dis = log(Hay_dis + 1),
                          l.Red_lev = log(Red_lev + 1),
                          l.Red_dis = log(Red_dis + 1),
                          l.Hay_lev = log(Hay_lev + 1),
                          l.Maude_dis = log(Maude_dis + 1),
                          l.Maude_lev = log(Maude_lev + 1))

gflow <- gflow %>% mutate(Yarr_lev = na.approx(Yarr_lev, maxgap = 10, na.rm = TRUE),
                          Yarr_dis = na.approx(Yarr_dis, maxgap = 10, na.rm = TRUE),
                          l.Yarr_dis = log(Yarr_dis + 1),
                          l.Yarr_lev = log(Yarr_lev + 1))

mflow <- mflow %>% mutate(ox_lev = na.approx(ox_lev, maxgap = 10, na.rm = TRUE),
                          ox_dis = na.approx(ox_dis, maxgap = 10, na.rm = TRUE),
                          l.ox_dis = log(ox_dis + 1),
                          l.ox_lev = log(ox_lev + 1))

pflow <- pflow %>% mutate(will_lev = na.approx(will_lev, maxgap = 10, na.rm = TRUE),
                          will_dis = na.approx(will_dis, maxgap = 10, na.rm = TRUE),
                          l.will_dis = log(will_dis + 1),
                          l.will_lev = log(will_lev + 1))

rflow <- rflow %>% mutate(Flows_dis = na.approx(Flows_dis, maxgap = 10, na.rm = TRUE),
                          RenPS_lev = na.approx(RenPS_lev, maxgap = 10, na.rm = TRUE),
                          l.Flows_dis = log(Flows_dis + 1),
                          l.RenPS_lev = log(RenPS_lev + 1))

yflow <- yflow %>% mutate(Wyan_lev = na.approx(Wyan_lev, maxgap = 10, na.rm = TRUE),
                          Wyan_dis = na.approx(Wyan_dis, maxgap = 10, na.rm = TRUE),
                          Tur_lev = na.approx(Tur_lev, maxgap = 10, na.rm = TRUE),
                          Tur_dis = na.approx(Tur_dis, maxgap = 10, na.rm = TRUE),
                          l.Tur_lev = log(Tur_lev + 1),
                          l.Tur_dis = log(Tur_dis + 1),
                          l.Wyan_dis = log(Wyan_dis + 1),
                          l.Wyan_lev = log(Wyan_lev + 1)) 
```


```{r Make binary flood presence/absence variable from water level data and bankfull levels}
cflow <- cflow %>% mutate(floodc = case_when(Four_lev >= 1.648 ~ 1, TRUE ~ 0),
                          floodb = case_when(Bool_lev >= 2.167 ~ 1, TRUE ~ 0))

gflow <- gflow %>% mutate(flood = case_when(Yarr_lev >= 2.885 ~ 1, TRUE ~ 0))

hflow <- hflow %>% mutate(flood = case_when(Col_lev >= 5.867 ~ 1, TRUE ~ 0))

lflow <- lflow %>% mutate(flood = case_when(Red_lev >= 5.604 ~ 1, TRUE ~ 0))

mflow <- mflow %>% mutate(flood = case_when(ox_lev >= 2.644 ~ 1, TRUE ~ 0))

pflow <- pflow %>% mutate(flood = case_when(will_lev >= 1.989 ~ 1, TRUE ~ 0))

rflow <- rflow %>% mutate(flood = case_when(RenPS_lev >= 16.6 ~ 1, TRUE ~ 0))

# using upper literature estimate as have very low confidence in my bankfull estimate (lack of data)
yflow <- yflow %>% mutate(flood = case_when(Tur_lev >= 2.46 ~ 1, TRUE ~ 0))
```



```{r summarise water level data by month and log transform}
# just cumbung, hattah and lowbidgee

cflow.m <- cflow %>% group_by(ym) %>% summarise(nB = sum(!is.na(Bool_lev)), nF = sum(!is.na(Four_lev)), Bool_levA = mean(Bool_lev, na.rm = TRUE), Four_levA = mean(Four_lev, na.rm = TRUE), Bool_levM = max(Bool_lev, na.rm = TRUE), Four_levM = max(Four_lev, na.rm = TRUE), Bool_disA = mean(Bool_dis, na.rm = TRUE), floodc = sum(floodc), floodb = sum(floodb)) %>% 
  mutate(l.Four_lev = log(Four_levA + 1),
        l.Bool_lev = log(Bool_levA + 1),
        l.Four_levM = log(Four_levM + 1),
        l.Bool_levM = log(Bool_levM + 1),
        l.Bool_dis = log(Bool_disA + 1))

hflow.m <- hflow %>% group_by(ym) %>% summarise(nB = sum(!is.na(Bar_lev)), nC = sum(!is.na(Col_lev)), Bar_levA = mean(Bar_lev, na.rm = TRUE), Bar_levM = max(Bar_lev, na.rm = TRUE), Col_levA = mean(Col_lev, na.rm = TRUE), Col_levM = max(Col_lev, na.rm = TRUE), flood = sum(flood)) %>% 
  mutate(l.Bar_lev = log(Bar_levA + 1),
         l.Col_lev = log(Col_levA + 1),
         l.Bar_levM = log(Bar_levM + 1),
         l.Col_levM = log(Col_levM + 1))


lflow.m <- lflow %>% group_by(ym) %>% summarise(nH = sum(!is.na(Hay_lev)), nR = sum(!is.na(Red_lev)), nM = sum(!is.na(Maude_lev)), Hay_levA = mean(Hay_lev, na.rm = TRUE), Hay_levM = max(Hay_lev, na.rm = TRUE), Red_levA = mean(Red_lev, na.rm = TRUE), Red_levM = max(Red_lev, na.rm = TRUE), Maude_levA = mean(Maude_lev, na.rm = TRUE), Maude_disA = mean(Maude_dis, na.rm = TRUE), flood = sum(flood)) %>% 
  mutate(l.Hay_lev = log(Hay_levA + 1),
         l.Red_lev = log(Red_levA + 1),
         l.Hay_levM = log(Hay_levM + 1),
         l.Red_levM = log(Red_levM + 1),
         l.Maude_dis = log(Maude_disA + 1),
         l.Maude_lev = log(Maude_levA + 1))
# this dataset starts with a bunch of NA's because water level data only starts in 1972

# Yantabulla for Turra water level vs discharge

yflow.m <- yflow %>% group_by(ym) %>% summarise(nTd = sum(!is.na(Tur_dis)), nTl = sum(!is.na(Tur_lev)), Tur_dis = mean(Tur_dis, na.rm = TRUE), Tur_lev = mean(Tur_lev, na.rm = TRUE)) %>% 
  mutate(l.Tur_lev = log(Tur_lev + 1),
         l.Tur_dis = log(Tur_dis + 1))
```


```{r summarise data per year and log transform annual data}
cflow.WY <- cflow %>% group_by(wateryear) %>% summarise(nBl = sum(!is.na(Bool_lev)), nFl = sum(!is.na(Four_lev)), nBd = sum(!is.na(Bool_dis)), nFd = sum(!is.na(Four_dis)), Bool_dis = mean(Bool_dis, na.rm = TRUE), Four_dis = mean(Four_dis, na.rm = TRUE), Bool_lev = mean(Bool_lev, na.rm = TRUE), Four_lev = mean(Four_lev, na.rm = TRUE), floodc = sum(floodc), floodb = sum(floodb)) %>% 
  mutate(l.Four_dis = log(Four_dis + 1),
        l.Four_lev = log(Four_lev + 1),
        l.Bool_dis = log(Bool_dis + 1),
        l.Bool_lev = log(Bool_lev + 1))

gflow.WY <- gflow %>% group_by(wateryear) %>% summarise(nYl = sum(!is.na(Yarr_lev)), nYd = sum(!is.na(Yarr_dis)), Yarr_dis = mean(Yarr_dis, na.rm = TRUE), Yarr_lev = mean(Yarr_lev, na.rm = TRUE), flood = sum(flood)) %>% 
  mutate(l.Yarr_dis = log(Yarr_dis + 1),
         l.Yarr_lev = log(Yarr_lev + 1))

hflow.WY <- hflow %>% group_by(wateryear) %>% summarise(nBl = sum(!is.na(Bar_lev)), nCl = sum(!is.na(Col_lev)), nBd = sum(!is.na(Bar_dis)), nCd = sum(!is.na(Col_dis)), Bar_dis = mean(Bar_dis, na.rm = TRUE), Bar_lev = mean(Bar_lev, na.rm = TRUE), Col_dis = mean(Col_dis, na.rm = TRUE), Col_lev = mean(Col_lev, na.rm = TRUE), flood = sum(flood)) %>% 
  mutate(l.Bar_lev = log(Bar_lev + 1),
         l.Bar_dis = log(Bar_dis + 1),
         l.Col_lev = log(Col_lev + 1),
         l.Col_dis = log(Col_dis + 1))

lflow.WY <- lflow %>% group_by(wateryear) %>% summarise(nHl = sum(!is.na(Hay_lev)), nRl = sum(!is.na(Red_lev)), nHd = sum(!is.na(Hay_dis)), nRd = sum(!is.na(Red_dis)), Hay_dis = mean(Hay_dis, na.rm = TRUE), Hay_lev = mean(Hay_lev, na.rm = TRUE), Red_lev = mean(Red_lev, na.rm = TRUE), Red_dis = mean(Red_dis, na.rm = TRUE), Maude_lev = mean(Maude_lev, na.rm = TRUE), Maude_dis = mean(Maude_dis, na.rm = TRUE), flood = sum(flood)) %>% 
  mutate(l.Hay_dis = log(Hay_dis + 1),
         l.Red_lev = log(Red_lev + 1),
         l.Red_dis = log(Red_dis + 1),
         l.Hay_lev = log(Hay_lev + 1),
         l.Maude_dis = log(Maude_dis + 1),
         l.Maude_lev = log(Maude_lev + 1))

mflow.WY <- mflow %>% group_by(wateryear) %>% summarise(nOl = sum(!is.na(ox_lev)), nOd = sum(!is.na(ox_dis)), ox_lev = mean(ox_lev, na.rm = TRUE), ox_dis = mean(ox_dis, na.rm = TRUE), flood = sum(flood)) %>% 
  mutate(l.ox_dis = log(ox_dis + 1),
         l.ox_lev = log(ox_lev + 1))

pflow.WY <- pflow %>% group_by(wateryear) %>% summarise(nWl = sum(!is.na(will_lev)), nWd = sum(!is.na(will_dis)), will_lev = mean(will_lev, na.rm = TRUE), will_dis = mean(will_dis, na.rm = TRUE), flood = sum(flood)) %>% 
  mutate(l.will_dis = log(will_dis + 1),
         l.will_lev = log(will_lev + 1))

# riverland - summarise daily data, and summarise monthly data
rflow.WY <- rflow %>% group_by(wateryear) %>% summarise(nFd = sum(!is.na(Flows_dis)), nRl = sum(!is.na(RenPS_lev)), Flows_dis = mean(Flows_dis, na.rm = TRUE), RenPS_lev = mean(RenPS_lev, na.rm = TRUE), flood = sum(flood)) %>% 
  mutate(l.Flows_dis = log(Flows_dis + 1),
         l.RenPS_lev = log(RenPS_lev + 1))
rmflow.WY <- rmflow %>% group_by(wateryear) %>% 
  summarise(nFd = sum(!is.na(Flows_dis)), Flows_dis = mean(Flows_dis, na.rm = TRUE)) %>% 
  mutate(l.Flows_dis = log(Flows_dis + 1))


yflow.WY <- yflow %>% group_by(wateryear) %>% summarise(nTl = sum(!is.na(Tur_lev)), nTd = sum(!is.na(Tur_dis)), nWl = sum(!is.na(Wyan_lev)), nWd = sum(!is.na(Wyan_dis)), Wyan_lev = mean(Wyan_lev, na.rm = TRUE), Wyan_dis = mean(Wyan_dis, na.rm = TRUE), Tur_lev = mean(Tur_lev, na.rm = TRUE), Tur_dis = mean(Tur_dis, na.rm = TRUE), flood = sum(flood)) %>% 
  mutate(l.Tur_lev = log(Tur_lev + 1),
         l.Tur_dis = log(Tur_dis + 1),
         l.Wyan_dis = log(Wyan_dis + 1),
         l.Wyan_lev = log(Wyan_lev + 1)) 
```

```{r Create an inundation vs booligal discharge dataframe}
#Add Booligal 95% complete years to a new dataframe
inun <- tibble(.rows = 69) %>% 
  add_column(Bool_dis = cflow.WY %>% filter(nBd >= 347) %>% .$Bool_dis,
             year = cflow.WY %>% filter(nBd >= 347) %>% .$wateryear)

# load annual inundation data 
cinun <- read_excel("raw/correlate_flow_flood.xlsx",sheet = "cumbung_inun") 

# combine these 
inun <- merge(inun, cinun, by = 'year') %>% 
  mutate(l.Bool_dis = log(Bool_dis + 1),
         l.area.km2 = log(area.km2 + 1))
```


# GLMs for correlating annual discharge data

```{r Cumbung}
# generalised linear model
# filter data within the model so that the actual dataset remains unchanged
M1 <- glm(l.Four_dis ~ l.Bool_dis, data=cflow.WY %>% filter(nFd >= 347, nBd >= 347), na.action = na.omit)

# check assumptions
par(mfrow = c(2, 2))
plot(M1) # top left test for equal variances, top right normality 
par(mfrow = c(1, 1))
ggplot(cflow.WY %>% filter(nFd >= 347, nBd >= 347), aes(l.Bool_dis, l.Four_dis), na.rm = TRUE) + geom_point() 

# very small dataset, assumptions all ok

summary(M1)
cor.test(cflow.WY$Bool_dis, cflow.WY$Four_dis)

# predict
cflow.WY$predict.l <- predict(M1, cflow.WY)
# create a variable that is predicted log colignan dis
cflow.WY$predict <- exp(cflow.WY$predict.l) - 1
# back transform 
plot(cflow.WY$Four_dis ~ cflow.WY$predict)


ggemmeans.Dis <- ggemmeans(M1, terms = c("l.Bool_dis[all]"))

options(scipen = 1)

c.flow <- plot(ggemmeans.Dis, dot.alpha = 1,  colors = "hero", ci = T, add.data = TRUE, jitter = 0, show.y.title = T, show.title = FALSE) + 
  labs(x="Booligal flows (ML/day)",y="Four Mile flows (ML/day)",title="Great Cumbung Swamp") +
    geom_line(size = 1) +
  theme_bw() +
  #geom_col(aes(x = Year, y = missing_dis), width = 31, fill = "#f0f0f0") + # not working yet
  theme(plot.title = element_text(size = 35,hjust = 0.5, face = "bold"),
        text = element_text(size = 38), # change font size
        axis.text = element_text(size = 32),
        plot.margin = margin(10,20,10,10), # give extra space on RHS
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()) + # centre title
  scale_y_continuous(breaks = pretty(cflow.WY$l.Four_dis), 
                     labels = signif(exp(pretty(cflow.WY$l.Four_dis)),1)) +
  scale_x_continuous(breaks = pretty(cflow.WY$l.Bool_dis), 
                     labels = signif(exp(pretty(cflow.WY$l.Bool_dis)),1))
c.flow


# make a new column that is a combination of the old and new data
# if the raw data aren't an NA, it uses the raw data. otherwise, it uses the predicted data
# also make a new column which is nNd (number of data points in each for the new data)
cflow.WY <- cflow.WY %>% mutate(new = case_when(!is.na(Four_dis) ~ Four_dis, TRUE ~ predict),
                                nNd = case_when(nFd != 0 ~ nFd, TRUE ~ nBd))

# check against just predicted data
plot(cflow.WY$new ~ cflow.WY$predict)
# as expected, almost all a 1:1 ratio as most data are from booligal, but some deviation due to four mile data - good
```

```{r Hattah lakes}
# generalised linear model
M1 <- glm(l.Col_dis ~ l.Bar_dis, data=hflow.WY %>% filter(nCd >= 347, nBd >= 347), na.action = na.omit)

# check assumptions
par(mfrow = c(2, 2))
plot(M1) # top left test for equal variances, top right normality 
par(mfrow = c(1, 1))
ggplot(hflow.WY %>% filter(nCd >= 347, nBd >= 347), aes(l.Bar_dis, l.Col_dis), na.rm = TRUE) +
  geom_point()

# log-transform makes normality and equal variance assumptions ok
# linearity is great with log-transformations
# therefore use log-transformed data

summary(M1)
cor.test(hflow.WY$l.Bar_dis, hflow.WY$l.Col_dis)

# predict
hflow.WY$predict.l <- predict(M1, hflow.WY) 
# create a variable that is predicted log colignan dis
hflow.WY$predict <- exp(hflow.WY$predict.l) - 1
# back transform 
plot(hflow.WY$Col_dis ~ hflow.WY$predict)
# visualise relationship between original and predicted colignan discharge

# plot correlation
ggemmeans.Dis <- ggemmeans(M1, terms = c("l.Bar_dis[all]"))
h.flow <- plot(ggemmeans.Dis, colors = "hero",  dot.alpha = 1, ci = T, add.data = TRUE, jitter = 0, show.y.title = T, show.title = FALSE) + labs(x="Barham flows (ML/day)",y="Colignan flows (ML/day)",title="Hattah Lakes")   +
    geom_line(size = 1) +
  theme_bw() +
  #geom_col(aes(x = Year, y = missing_dis), width = 31, fill = "#f0f0f0") + # not working yet
  theme(text = element_text(size = 38), # change font size
        axis.text = element_text(size = 32),
        plot.margin = margin(10,20,10,10), # give extra space on RHS
        plot.title = element_text(hjust = 0.5, face = "bold", size = 35),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()) + # centre title
  scale_y_continuous(breaks = pretty(hflow.WY$l.Col_dis, n = 3), 
                     labels = signif(exp(pretty(hflow.WY$l.Col_dis, n = 3)),1)) +
  scale_x_continuous(breaks = pretty(hflow.WY$l.Bar_dis, n = 3), 
                     labels = signif(exp(pretty(hflow.WY$l.Bar_dis, n = 3)),1))
h.flow
# export
#write.csv(hflow.WY, "output/hattah_pred.csv")


# make a new column that is a combination of the old and new data
hflow.WY <- hflow.WY %>% mutate(new = case_when(!is.na(Col_dis) ~ Col_dis, TRUE ~ predict),
                                nNd = case_when(nCd != 0 ~ nCd, TRUE ~ nBd))

# check against just predicted data
plot(hflow.WY$new ~ hflow.WY$predict) # as expected
```

```{r Lowbidgee}
# generalised linear model
M1 <- glm(l.Red_dis ~ l.Hay_dis, data=lflow.WY %>% filter(nHd >= 347, nRd >= 347), na.action = na.omit)

# check assumptions
par(mfrow = c(2, 2))
plot(M1) # top left test for equal variances, top right normality 
par(mfrow = c(1, 1))
ggplot(lflow.WY %>% filter(nHd >= 347, nRd >= 347), aes(l.Hay_dis, l.Red_dis), na.rm = TRUE) +
  geom_point()

# log-transform makes normality and equal variance assumptions good
# linearity fine in all cases
# therefore use log-transformed data

summary(M1)
cor.test(lflow.WY$l.Hay_dis, lflow.WY$l.Red_dis)

# predict
lflow.WY$predict.l <- predict(M1, lflow.WY)
# create a variable that is predicted log redbank dis
lflow.WY$predict <- exp(lflow.WY$predict.l) - 1
# back transform 
plot(lflow.WY$Red_dis ~ lflow.WY$predict)
# visualise relationship between original and predicted redbank discharge

# visualise correlation
ggemmeans.Dis <- ggemmeans(M1, terms = c("l.Hay_dis[all]"))

l.flow <- plot(ggemmeans.Dis, colors = "hero",  dot.alpha = 1, ci = T, add.data = TRUE, jitter = 0, show.y.title = T, show.title = FALSE) + labs(x="Hay flows (ML/day)",y="Redbank flows (ML/day)",title="Lowbidgee Wetlands") +
    geom_line(size = 1) +
  theme_bw() +
  #geom_col(aes(x = Year, y = missing_dis), width = 31, fill = "#f0f0f0") + # not working yet
  theme(text = element_text(size = 38), # change font size
        axis.text = element_text(size = 32),
        plot.margin = margin(10,20,10,10), # give extra space on RHS
        plot.title = element_text(hjust = 0.5, face = "bold", size = 35),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()) + # centre title
  scale_y_continuous(breaks = pretty(lflow.WY$l.Red_dis, n = 3), 
                     labels = signif(exp(pretty(lflow.WY$l.Red_dis, n = 3)),1)) +
  scale_x_continuous(breaks = pretty(lflow.WY$l.Hay_dis, n = 3), 
                     labels = signif(exp(pretty(lflow.WY$l.Hay_dis, n = 3)),1))
l.flow
# export
#write.csv(lflow.WY, "output/lowbidgee_pred.csv")

# make a new column that is a combination of the old and new data
lflow.WY <- lflow.WY %>% mutate(new = case_when(!is.na(Red_dis) ~ Red_dis, TRUE ~ predict),
                                nNd = case_when(nRd != 0 ~ nRd, TRUE ~ nHd))

# check against just predicted data
plot(lflow.WY$new ~ lflow.WY$predict) # as expected
```



# Correlating Inundation area with Booligal discharge

```{r Cumbung Inundation area vs Booligal discharge}
# generalised linear model
M1 <- glm(l.area.km2 ~ l.Bool_dis, data=inun)
summary(M1)
cor.test(inun$l.Bool_dis, inun$l.area.km2)

 # check assumptions
par(mfrow = c(2, 2))
plot(M1) # top left test for equal variances, top right normality
# better with log transformed data, acceptable
par(mfrow = c(1, 1))
ggplot(inun, aes(l.Bool_dis, l.area.km2)) +
  geom_point() # test for linear relationship
# log-log relationship looks roughly linear

# predict
inun$predict.l <- predict(M1, inun)
# create a variable that is predicted log inundation area
inun$predict <- exp(inun$predict.l)
# back transform 
plot(inun$area.km2 ~ inun$predict)



# visualise relationship between original and predicted inundation
ggemmeans.Inun <- ggemmeans(M1, terms = c("l.Bool_dis[all]")) # this must specify the second term listed in glm

c.inun <- plot(ggemmeans.Inun, colors = "hero", ci = T, add.data = TRUE, jitter = 0, show.y.title = T, show.title = FALSE) + labs(x="Booligal Flows (log)",y="Cumbung Inundation Area (log)",title="") +
    geom_line(size = 1) +
  theme_bw() +
  #geom_col(aes(x = Year, y = missing_dis), width = 31, fill = "#f0f0f0") + # not working yet
  theme(text = element_text(size = 22), # change font size
        axis.text = element_text(size = 24),
        plot.margin = margin(10,20,10,10), # give extra space on RHS
        plot.title = element_text(hjust = 0.5),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()) # centre title
c.inun

# export
#write.csv(inun, "output/cbool_pred_inun.csv")
```



# Generalised additive models - monthly flood data - Cumbung, Hattah, Lowbidgee

Using GAMs instead of GLMs as GLMs don't fit well and assumptions aren't acceptable.

```{r Cumbung}
# modelling monthly data
m <- gam(l.Four_lev ~   s(l.Bool_lev, k=3), data=cflow.m %>% filter(nB >= 29, nF >= 29) %>%  subset(Bool_levA > 0, Four_levA > 0))
# k = 3 is 3rd order polynomial
# only include months with >95% data and water levels above zero

summary(m)

# check normality and equal variances (linearity is not necessary)
par(mfrow = c(2, 2))
gam.check(m) # residual plots for GAMs
# normality seems ok, a bit strange but doesn't improve much with transformation
par(mfrow = c(1, 1))

# predict
cflow.m$predict.l <- predict(m,newdata = cflow.m)
cflow.m$predict <- exp(cflow.m$predict.l) - 1


# plot
ggpredict <- ggemmeans(m, terms = c("l.Bool_lev [all]"))
c.flood <- ggplot(ggpredict,aes(x = x, y = predicted))+
        geom_ribbon(aes(ymin =conf.low, ymax = conf.high),alpha=0.2) +
        geom_line(size=0.5)+
        geom_point(data=cflow.m %>% filter(nB >= 29, nF >= 29) %>%  subset(Bool_levA > 0, Four_levA > 0), aes(x=l.Bool_lev, y=l.Four_lev)) +
        theme_bw() +
        theme(panel.grid.major = element_blank(),
              panel.grid.minor=element_blank(),
              text = element_text(size = 38),
              axis.title = element_text(size = 38),
              plot.title = element_text(hjust = 0.5))  + 
        #scale_y_continuous(limits = c(0,1)) + 
        scale_x_continuous(labels = scales::comma)+
        labs(x="Booligal water level (m)", y="Four Mile water level (m)", title = "") +
  scale_y_continuous(breaks = pretty(cflow.m$l.Four_lev, n = 3), 
                     labels = signif(exp(pretty(cflow.m$l.Four_lev, n = 3)),1)) +
  scale_x_continuous(breaks = pretty(cflow.m$l.Bool_lev, n = 3), 
                     labels = signif(exp(pretty(cflow.m$l.Bool_lev, n = 3)),1))
c.flood

```


```{r Hattah}
# modelling monthly data
m <- gam(l.Col_lev ~   s(l.Bar_lev, k=3), data=hflow.m %>% filter(nB >= 29, nC >= 29) %>%  subset(Bar_levA > 0, Col_levA > 0))
# k = 3 is 3rd order polynomial
# only include months with >95% data and water levels above zero

summary(m)

# check normality and equal variances (linearity is not necessary)
par(mfrow = c(2, 2))
gam.check(m) # residual plots for GAMs
# normality somewhat better with log transformed

# predict
hflow.m$predict.l <- predict(m,newdata = hflow.m)
hflow.m$predict <- exp(hflow.m$predict.l) - 1


# plot
ggpredict <- ggemmeans(m, terms = c("l.Bar_lev [all]"))
h.flood <- ggplot(ggpredict,aes(x = x, y = predicted))+
        geom_ribbon(aes(ymin =conf.low, ymax = conf.high),alpha=0.2) +
        geom_line(size=0.5)+
        geom_point(data=hflow.m %>% filter(nB >= 29, nC >= 29) %>%  subset(Bar_levA > 0, Col_levA > 0), aes(x=l.Bar_lev, y=l.Col_lev))+
        theme_bw(base_size=16) +
        theme(panel.grid.major = element_blank(),
              panel.grid.minor=element_blank(),
              text = element_text(size = 38),
              axis.title = element_text(size = 38),
              plot.title = element_text(hjust = 0.5))  + 
        #scale_y_continuous(limits = c(0,1)) + 
        scale_x_continuous(labels = scales::comma)+
        labs(x="Barham water level (m)", y="Colignan water level (m)", title = "") +
  scale_y_continuous(breaks = pretty(hflow.m$l.Col_lev, n = 3), 
                     labels = signif(exp(pretty(hflow.m$l.Col_lev, n = 3)),1)) +
  scale_x_continuous(breaks = pretty(hflow.m$l.Bar_lev, n = 3), 
                     labels = signif(exp(pretty(hflow.m$l.Bar_lev, n = 3)),1))
h.flood
```


```{r Lowbidgee}
# modelling monthly data
m <- gam(l.Red_lev ~   s(l.Hay_lev, k=4), data=lflow.m %>% filter(nH >= 29, nR >= 29) %>%  subset(Hay_levA >= 0, Red_levA >= 0))
# k = 3 is 3rd order polynomial
# only include months with >95% data and water levels above zero

summary(m)

# check normality and equal variances (linearity is not necessary)
par(mfrow = c(2, 2))
gam.check(m) # residual plots for GAMs

# predict
lflow.m$predict.l <- predict(m,newdata = lflow.m)
lflow.m$predict <- exp(lflow.m$predict.l) - 1

# plot
ggpredict <- ggemmeans(m, terms = c("l.Hay_lev [all]"))
l.flood <- ggplot(ggpredict,aes(x = x, y = predicted))+
        geom_ribbon(aes(ymin =conf.low, ymax = conf.high),alpha=0.2) +
        geom_line(size=0.5)+
        geom_point(data=lflow.m %>% filter(nH >= 29, nR >= 29) %>%  subset(Hay_levA >= 0, Red_levA >= 0), aes(x=l.Hay_lev, y=l.Red_lev))+
        theme_bw(base_size=16) +
        theme(panel.grid.major = element_blank(),
              panel.grid.minor=element_blank(),
              text = element_text(size = 38),
              axis.title = element_text(size = 38),
              plot.title = element_text(hjust = 0.5))  + 
        #scale_y_continuous(limits = c(0,1)) + 
        scale_x_continuous(labels = scales::comma)+
        labs(x="Hay water level (m)", y="Redbank water level (m)", title = "") +
  scale_y_continuous(breaks = pretty(lflow.m$l.Red_lev, n = 3), 
                     labels = signif(exp(pretty(lflow.m$l.Red_lev, n = 3)),1)) +
  scale_x_continuous(breaks = pretty(lflow.m$l.Hay_lev, n = 3), 
                     labels = signif(exp(pretty(lflow.m$l.Hay_lev, n = 3)),1))
l.flood
```


# Correlate monthly water levels with monthly number of flood days

```{r Cumbung}
# plot relationship between raw water level data and flood days
ggplot(cflow.m, aes(Four_levA, floodc)) +
  geom_point()
# this is a threshold relationship, where there are no flood days until water level > 0.8m


# make a model, only using water level data > 0.8m
m <- gam(floodc ~ s(l.Four_lev, k=3), data=cflow.m %>% filter(nF >= 29, Four_levA >= 0.8))
summary(m)


# check normality and equal variances (linearity is not necessary)
par(mfrow = c(2, 2))
gam.check(m) # residual plots for GAMs
# ok
par(mfrow = c(1, 1))


# plot the modelled relationship
ggpredict <- ggemmeans(m, terms = c("l.Four_lev [all]"))
c.floodday <- ggplot(ggpredict,aes(x = x, y = predicted))+
        geom_ribbon(aes(ymin =conf.low, ymax = conf.high),alpha=0.2) +
        geom_line(size=0.5)+
        geom_point(data=cflow.m %>% filter(nF >= 29, Four_levA >= 0.8), aes(x=l.Four_lev, y=floodc))+
        theme_bw(base_size=16) +
        theme(panel.grid.major = element_blank(),
              panel.grid.minor=element_blank(),
              text = element_text(size = 38),
              axis.title = element_text(size = 38),
              plot.title = element_text(hjust = 0.5))  + 
        #scale_y_continuous(limits = c(0,1)) + 
        scale_x_continuous(labels = scales::comma)+
        labs(x="Four Mile water level (m)", y="Number of flood days", title = "")+
  scale_x_continuous(breaks = pretty(cflow.m$l.Four_lev), 
                     labels = signif(exp(pretty(cflow.m$l.Four_lev)),2)) +
  scale_y_continuous(limits = c(-5, 35))
c.floodday

# predict number of flood days based on predicted four mile water levels

# create a new dataset with the predicted data reassigned to have the same name as water level data in this model
cflow.mp <- cflow.m %>% 
  select(ym, floodc, predict) %>% 
  mutate(l.Four_lev = log(predict + 1)) 
# create a predicted flood days variable by predicting based on this model
cflow.mp$predict.f <- case_when(cflow.mp$predict >= 0.8 ~ predict(m,newdata = cflow.mp), TRUE ~ 0) 
# reassign any negatives to zero (impossible to have a negative number of days)
cflow.mp$predict.f <- case_when(cflow.mp$predict.f < 0 ~ 0, TRUE ~ cflow.mp$predict.f)


# visualise this
plot(cflow.mp$ym, cflow.mp$predict.f)
# visualise relationship between time and flood days with original data
plot(cflow.mp$ym, cflow.mp$floodc)
```


```{r Hattah}
# plot
ggplot(hflow.m, aes(Col_levA, flood)) +
  geom_point()
# threshold relationship - above about 6.3m water level, floods every day, and when below 3.5m,  never floods
ggplot(hflow.m %>% filter(Col_levA <= 6.3, Col_levA >= 3.5), aes(Col_levA, flood)) +
  geom_point()
# so just model relationship within these thresholds


# modelling monthly data - need k = 4 to flatten initial curve
m <- gam(flood ~   s(Col_levA, k=4), data=hflow.m %>% filter(nC >= 29, Col_levA <= 6.3, Col_levA >= 3.5) %>%  subset(Col_levA > 0))
#m <- gam(flood ~   s(l.Col_lev, k=4), data=hflow.m %>% filter(nC >= 29, Col_levA <= 6.3, Col_levA >= 3.5) %>%  subset(Col_levA > 0))
summary(m)

# check normality and equal variances (linearity is not necessary)
par(mfrow = c(2, 2))
gam.check(m) # residual plots for GAMs
# almost identical transformed and raw, use raw
par(mfrow = c(1, 1))

# plot
ggpredict <- ggemmeans(m, terms = c("Col_levA [all]"))
h.floodday <- ggplot(ggpredict,aes(x = x, y = predicted))+
        geom_ribbon(aes(ymin =conf.low, ymax = conf.high),alpha=0.2) +
        geom_line(size=0.5)+
        geom_point(data=hflow.m %>% filter(nC >= 29, Col_levA <= 6.3, Col_levA >= 3.5) %>%  subset(Col_levA > 0), aes(x=Col_levA, y=flood))+
        theme_bw(base_size=16) +
        theme(panel.grid.major = element_blank(),
              panel.grid.minor=element_blank(),
              text = element_text(size = 38),
              axis.title = element_text(size = 38),
              plot.title = element_text(hjust = 0.5))  + 
        scale_x_continuous(labels = pretty(hflow.m$Col_levA, n = 6), 
                           breaks = pretty(hflow.m$Col_levA, n = 6))+
        labs(x="Colignan water level (m)", y="Number of flood days", title = "")  +
  scale_y_continuous(limits = c(-5, 35))
h.floodday


# create a new dataset with the predicted data reassigned to have the same name as water level data in this model
hflow.mp <- hflow.m %>% 
  select(ym, flood, predict) %>% 
  mutate(Col_levA = predict) 
# create a predicted flood days variable by predicting based on this model
hflow.mp$predict.f <- case_when(hflow.mp$Col_levA <= 3.5 ~ 0, hflow.mp$Col_levA <= 6.3 ~ predict(m,newdata = hflow.mp), TRUE ~ 30.5)
# reassign any negatives to zero (impossible to have a negative number of days)
hflow.mp$predict.f <- case_when(hflow.mp$predict.f < 0 ~ 0, TRUE ~ hflow.mp$predict.f)


# visualise this
plot(hflow.mp$ym, hflow.mp$predict.f)
# visualise relationship between time and flood days with original data
plot(hflow.mp$ym, hflow.mp$flood)
```



```{r Lowbidgee}
# plot
ggplot(lflow.m, aes(Red_levA, flood)) +
  geom_point()
# only floods with water levels >2m, therefore start with this threshold


# modelling monthly data - use k = 4 to help flatten early relationship
m <- gam(flood ~   s(Red_levA, k=4), data=lflow.m %>% filter(nR >= 29, Red_levA >= 2) %>%  subset(Red_levA > 0))
summary(m)

# check normality and equal variances (linearity is not necessary)
par(mfrow = c(2, 2))
gam.check(m) # residual plots for GAMs
# plots look a bit weird regardless of transformation, so use raw
par(mfrow = c(1, 1))

# plot
ggpredict <- ggemmeans(m, terms = c("Red_levA [all]"))
l.floodday <- ggplot(ggpredict,aes(x = x, y = predicted))+
        geom_ribbon(aes(ymin =conf.low, ymax = conf.high),alpha=0.2) +
        geom_line(size=0.5)+
        geom_point(data=lflow.m %>% filter(nR >= 29, Red_levA >= 2) %>%  subset(Red_levA > 0), aes(x=Red_levA, y=flood))+
        theme_bw(base_size=16) +
        theme(panel.grid.major = element_blank(),
              panel.grid.minor=element_blank(),
              text = element_text(size = 38),
              axis.title = element_text(size = 38),
              plot.title = element_text(hjust = 0.5))  + 
        scale_x_continuous(labels = pretty(lflow.m$Red_levA, n = 6), 
                           breaks = pretty(lflow.m$Red_levA, n = 6)) +
        labs(x="Redbank water level (m)", y="Number of flood days", title = "")  +
  scale_y_continuous(limits = c(-5, 35))
l.floodday

# create a new dataset with the predicted data reassigned to have the same name as water level data in this model
lflow.mp <- lflow.m %>% 
  select(ym, flood, predict) %>% 
  mutate(Red_levA = predict) 
# create a predicted flood days variable by predicting based on this model
lflow.mp$predict.f <- case_when(lflow.mp$Red_levA <= 2 ~ 0, TRUE ~ predict(m,newdata = lflow.mp))
# reassign any negatives to zero (impossible to have a negative number of days)
lflow.mp$predict.f <- case_when(lflow.mp$predict.f < 0 ~ 0, TRUE ~ lflow.mp$predict.f)


# visualise this
plot(lflow.mp$ym, lflow.mp$predict.f)
# visualise relationship between time and flood days with original data
plot(lflow.mp$ym, lflow.mp$flood)

```

```{r Sum monthly flood days for these datasets to create annual data}
# at this point, only use predicted flood day data, rather than combining with raw data, as predicted data have many more months with 1 flood day (rather than zero), so would make it seem like floods had declined more than they actually have if used in combination with raw data

cflow.mp$Year <- year(cflow.mp$ym)
cflood <- cflow.mp %>%  group_by(Year) %>% summarise(Flood_days = sum(predict.f), nF = sum(!is.na(predict.f))) %>% subset(nF >= 11)  %>% 
  complete(Year = seq(1946, 2020, by = 1)) # add gaps for missing years
# only including years with at least 11 months (95%) of data

hflow.mp$Year <- year(hflow.mp$ym)
hflood <- hflow.mp %>%  group_by(Year) %>% summarise(Flood_days = sum(predict.f), nF = sum(!is.na(predict.f))) %>% subset(nF >= 11) %>% 
  complete(Year = seq(1913, 2020, by = 1))

lflow.mp$Year <- year(lflow.mp$ym)
lflood <- lflow.mp %>% group_by(Year) %>% summarise(Flood_days = sum(predict.f),nF = sum(!is.na(predict.f))) %>% subset(nF >= 11) %>% 
  complete(Year = seq(1956, 2020, by = 1))
```


# Correlating discharge and water level for some gauges used for bankfull analysis

```{r Try plotting daily data}
# untransformed
plot(lflow$Maude_lev, lflow$Maude_dis) # looks good
plot(cflow$Bool_lev, cflow$Bool_dis) # pretty good
plot(yflow$Tur_lev, yflow$Tur_dis) # looks good

# log
plot(lflow$l.Maude_lev, lflow$l.Maude_dis) 
plot(cflow$l.Bool_lev, cflow$l.Bool_dis) 
plot(yflow$l.Tur_lev, yflow$l.Tur_dis) 
# look no better or worse with log-transformations

# all are non-linear though, so use GAMs

# also, Turra is zero inflated. These should be real zero discharge values, but we should model without them as we're interested in discharges above zero
```


```{r Try plotting monthly data}
# untransformed
plot(lflow.m$Maude_levA, lflow.m$Maude_disA) # looks good
plot(cflow.m$Bool_levA, cflow.m$Bool_disA) # pretty good
plot(yflow.m$Tur_lev, yflow.m$Tur_dis) # doesn't seem to be much of a relationship

# log
plot(lflow.m$l.Maude_lev, lflow.m$l.Maude_dis) 
plot(cflow.m$l.Bool_lev, cflow.m$l.Bool_dis) 
plot(yflow.m$l.Tur_lev, yflow.m$l.Tur_dis) 
```


```{r Try plotting annual data}
# untransformed
plot(lflow.WY$Maude_lev, lflow.WY$Maude_dis) # looks good
plot(cflow.WY$Bool_lev, cflow.WY$Bool_dis) # pretty good
plot(yflow.WY$Tur_lev, yflow.WY$Tur_dis) # doesn't seem to be much of a relationship

# log
plot(lflow.WY$l.Maude_lev, lflow.WY$l.Maude_dis) 
plot(cflow.WY$l.Bool_lev, cflow.WY$l.Bool_dis) 
plot(yflow.WY$l.Tur_lev, yflow.WY$l.Tur_dis) 

# untransformed look linear, so could potentially use them with a GLM
# Turra doesnt seem to have much of a trend, but is better with transformation
```


Generalised additive models

```{r Try daily data}
m <- gam(l.Bool_lev ~   s(l.Bool_dis, k=3), data=cflow %>%  subset(Bool_lev > 0, Bool_dis > 0))
m <- gam(Bool_lev ~   s(Bool_dis, k=3), data=cflow %>%  subset(Bool_lev > 0, Bool_dis > 0))
# k = 3 is 3rd order polynomial
# only include months with >95% data and water levels above zero

summary(m)

# check normality and equal variances (linearity is not necessary)
par(mfrow = c(2, 2))
gam.check(m) # residual plots for GAMs
par(mfrow = c(1, 1))

# normality assumption is not satisfied, residual plots look terrible with untransformed and log-transformed data
```

```{r Try monthly data}
# booligal
m <- gam(l.Bool_lev ~   s(l.Bool_dis, k=3), data=cflow.m %>%  filter(Bool_levA > 0, Bool_disA > 0))
m <- gam(Bool_levA ~   s(Bool_disA, k=3), data=cflow.m %>%  filter(Bool_levA > 0, Bool_disA > 0))
# k = 3 is 3rd order polynomial
# only include months with >95% data and water levels above zero

summary(m)

# check normality and equal variances (linearity is not necessary)
par(mfrow = c(2, 2))
gam.check(m) # residual plots for GAMs
par(mfrow = c(1, 1))

# looks better than daily
# log improves normality somewhat but makes other plots look a bit strange


# maude
m <- gam(l.Maude_lev ~   s(l.Maude_dis, k=3), data=lflow.m %>%  filter(Maude_levA > 0, Maude_disA > 0))
m <- gam(Maude_levA ~   s(Maude_disA, k=3), data=lflow.m %>%  filter(Maude_levA > 0, Maude_disA > 0))
# k = 3 is 3rd order polynomial
# only include months with >95% data and water levels above zero

summary(m)

# check normality and equal variances (linearity is not necessary)
par(mfrow = c(2, 2))
gam.check(m) # residual plots for GAMs
par(mfrow = c(1, 1))

# log is much much better


# turra
m <- gam(l.Tur_lev ~   s(l.Tur_dis, k=3), data=yflow.m %>%  filter(Tur_lev > 0, Tur_dis > 0))
m <- gam(Tur_lev ~   s(Tur_dis, k=3), data=yflow.m %>%  filter(Tur_lev > 0, Tur_dis > 0))
# k = 3 is 3rd order polynomial
# only include months with >95% data and water levels above zero

summary(m)

# check normality and equal variances (linearity is not necessary)
par(mfrow = c(2, 2))
gam.check(m) # residual plots for GAMs
par(mfrow = c(1, 1))

# log is better
```


Therefore use monthly data & log-transform all

```{r Booligal}

m <- gam(l.Bool_lev ~   s(l.Bool_dis, k=4), data=cflow.m %>%  filter(Bool_levA > 0, Bool_disA > 0, nB >= 29, l.Bool_dis > 2))
# k = 4 is 4 knots (captures curve better, except very high values, which aren't of concern anyway)
# only include months with >95% data and water levels above zero
# start with log flow > 2 as few data points are before this

summary(m)

# check normality and equal variances (linearity is not necessary)
par(mfrow = c(2, 2))
gam.check(m) # residual plots for GAMs
par(mfrow = c(1, 1))


# plot
ggpredict <- ggemmeans(m, terms = c("l.Bool_dis [all]"))
b.corr <- ggplot(ggpredict,aes(x = x, y = predicted))+
        geom_ribbon(aes(ymin =conf.low, ymax = conf.high),alpha=0.2) +
        geom_line(size=0.5)+
        geom_point(data=cflow.m %>% filter(nB >= 29, Bool_levA > 0, Bool_disA > 0), aes(x=l.Bool_dis, y=l.Bool_lev)) +
        theme_bw() +
        theme(panel.grid.major = element_blank(),
              panel.grid.minor=element_blank(),
              text = element_text(size = 38),
              axis.title = element_text(size = 38),
              plot.title = element_text(hjust = 0.5))  + 
        #scale_y_continuous(limits = c(0,1)) + 
        scale_x_continuous(labels = scales::comma)+
        labs(x="Booligal flows (log)", y="Booligal water level (log)", title = "")
b.corr



# predict
cflow.m$predictL.l <- case_when(cflow.m$l.Bool_dis <= 2 ~ 0, TRUE ~ predict(m,newdata = cflow.m))
cflow.m$predictL <- exp(cflow.m$predictL.l) - 1
# reassign any negatives to zero (impossible to have a negative number of days)
cflow.m$predictL <- case_when(cflow.m$predictL < 0 ~ 0, TRUE ~ cflow.m$predictL)

# we are interested in 2500ML/day - see what predicted value this correlates to in the below
view(cflow.m)
# 1.97m
```


```{r Maude}

m <- gam(l.Maude_lev ~   s(l.Maude_dis, k=3), data=lflow.m %>%  filter(Maude_levA > 0, Maude_disA > 0, nM >= 29))
# only include months with >95% data and water levels above zero

summary(m)

# check normality and equal variances (linearity is not necessary)
par(mfrow = c(2, 2))
gam.check(m) # residual plots for GAMs
par(mfrow = c(1, 1))


# plot
ggpredict <- ggemmeans(m, terms = c("l.Maude_dis [all]"))
m.corr <- ggplot(ggpredict,aes(x = x, y = predicted))+
        geom_ribbon(aes(ymin =conf.low, ymax = conf.high),alpha=0.2) +
        geom_line(size=0.5)+
        geom_point(data=lflow.m %>% filter(nM >= 29, Maude_levA > 0, Maude_disA > 0), aes(x=l.Maude_dis, y=l.Maude_lev)) +
        theme_bw() +
        theme(panel.grid.major = element_blank(),
              panel.grid.minor=element_blank(),
              text = element_text(size = 38),
              axis.title = element_text(size = 38),
              plot.title = element_text(hjust = 0.5))  + 
        #scale_y_continuous(limits = c(0,1)) + 
        scale_x_continuous(labels = scales::comma)+
        labs(x="Maude flows (log)", y="Maude water level (log)", title = "")
m.corr



# predict
lflow.m$predictL.l <- predict(m,newdata = lflow.m)
lflow.m$predictL <- exp(lflow.m$predictL.l) - 1
# reassign any negatives to zero (impossible to have a negative number of days)
lflow.m$predictL <- case_when(lflow.m$predictL < 0 ~ 0, TRUE ~ lflow.m$predictL)

# we are interested in 20000ML/day - see what predicted value this correlates to in the below
view(lflow.m)
# 6.64m
```


```{r Turra}

m <- gam(l.Tur_lev ~   s(l.Tur_dis, k=4), data=yflow.m %>%  filter(Tur_lev > 0, Tur_dis > 0, nTd >= 29, nTl >= 29))
# k = 4 fits data better
# untransformed looks better for normality, but spreads data worse - use log transformed
# only include months with >95% data and water levels above zero

summary(m)

# check normality and equal variances (linearity is not necessary)
par(mfrow = c(2, 2))
gam.check(m) # residual plots for GAMs
par(mfrow = c(1, 1))


# plot
ggpredict <- ggemmeans(m, terms = c("l.Tur_dis [all]"))
t.corr <- ggplot(ggpredict,aes(x = x, y = predicted))+
        geom_ribbon(aes(ymin =conf.low, ymax = conf.high),alpha=0.2) +
        geom_line(size=0.5)+
        geom_point(data=yflow.m %>% filter(nTd >= 29, nTl >= 29, Tur_lev > 0, Tur_dis > 0), aes(x=l.Tur_dis, y=l.Tur_lev)) +
        theme_bw() +
        theme(panel.grid.major = element_blank(),
              panel.grid.minor=element_blank(),
              text = element_text(size = 38),
              axis.title = element_text(size = 38),
              plot.title = element_text(hjust = 0.5))  + 
        #scale_y_continuous(limits = c(0,1)) + 
        scale_x_continuous(labels = scales::comma)+
        labs(x="Turra flows (log)", y="Turra water level (log)", title = "")
t.corr



# predict
yflow.m$predictL.l <- predict(m,newdata = yflow.m)
yflow.m$predictL <- exp(yflow.m$predictL.l) - 1
# reassign any negatives to zero (impossible to have a negative number of days)
yflow.m$predictL <- case_when(yflow.m$predictL < 0 ~ 0, TRUE ~ yflow.m$predictL)

# we are interested in 1000 and 3700ML/day - see what predicted value this correlates to in the below
view(yflow.m)
# 1.9 and 2.38m - average = 2.14m
```

# Discharge Time series analysis - generalised linear models

```{r cumbung}
# remove years with too few data
cflow.WY.c <- cflow.WY %>% filter(nNd >= 347)

M1 <- glm(new ~ wateryear, data = cflow.WY.c)

# assumptions
par(mfrow = c(2, 2))
plot(M1) # OK
par(mfrow = c(1, 1)) 
ggplot(cflow.WY.c, aes(wateryear, new), na.rm = TRUE) +
  geom_point() 
# no obvious non-linear trends


# predict
cflow.WY.c$predict.Time <- predict(M1, cflow.WY.c)
# create a variable that is predicted four mile discharge over time


# plot
ggemmeans.dis <- ggemmeans(M1, terms = c("wateryear[all]"))

plot(ggemmeans.dis, colors = "hero", ci = T, add.data = TRUE, jitter = 0, show.y.title = T, show.title = FALSE) + labs(x="Year",y="Predicted Four mile weir discharge (ML/day)",title="")


# test
summary(M1)
cor.test(cflow.WY.c$new, cflow.WY.c$wateryear)


# calculate differences over time
100-(cflow.WY.c$predict.Time[length(cflow.WY.c$predict.Time)]/cflow.WY.c$predict.Time[1]*100)
# diff in last 50 years - since 1970, which is row 22
100-(cflow.WY.c$predict.Time[length(cflow.WY.c$predict.Time)]/cflow.WY.c$predict.Time[22]*100)

# average annual % change
(100-(cflow.WY.c$predict.Time[length(cflow.WY.c$predict.Time)]/cflow.WY.c$predict.Time[1]*100))/length(cflow.WY.c$predict.Time)

# prepare data for plotting
cflow.WY.cp <- cflow.WY.c %>% rename(Year = wateryear,  Discharge = new) %>% 
  complete(Year = seq(1946, 2020, by = 1)) # add gaps for missing years
```



```{r booligal}
# remove years with too few data
cflow.WY.b <- cflow.WY %>% filter(nBd >= 347)

M1 <- glm(l.Bool_dis ~ wateryear, data = cflow.WY.b)

# assumptions
par(mfrow = c(2, 2))
plot(M1) # use log
par(mfrow = c(1, 1)) 
ggplot(cflow.WY.b, aes(wateryear, l.Bool_dis), na.rm = TRUE) +
  geom_point() 
# no obvious non-linear trends


# predict
cflow.WY.b$predict.Time.l <- predict(M1, cflow.WY.b)
# create a variable that is predicted log booligal discharge over time
cflow.WY.b$predict.Time <- exp(cflow.WY.b$predict.Time.l)-1
plot(cflow.WY.b$Bool_dis ~ cflow.WY.b$predict.Time)
plot(cflow.WY.b$predict.Time ~ cflow.WY.b$wateryear)


# plot
ggemmeans.dis <- ggemmeans(M1, terms = c("wateryear[all]"))

plot(ggemmeans.dis, colors = "hero", ci = T, add.data = TRUE, jitter = 0, show.y.title = T, show.title = FALSE) + labs(x="Year",y="log Booligal discharge (ML/day)",title="")


# test
summary(M1)
cor.test(cflow.WY.b$l.Bool_dis, cflow.WY.b$wateryear)


# calculate differences over time
100-(cflow.WY.b$predict.Time[length(cflow.WY.b$predict.Time)]/cflow.WY.b$predict.Time[1]*100)
# diff in last 50 years - since 1970, which is row 22
100-(cflow.WY.b$predict.Time[length(cflow.WY.b$predict.Time)]/cflow.WY.b$predict.Time[22]*100)

# average annual % change
(100-(cflow.WY.b$predict.Time[length(cflow.WY.b$predict.Time)]/cflow.WY.b$predict.Time[1]*100))/length(cflow.WY.b$predict.Time)


# prepare data for plotting
cflow.WY.bp <- cflow.WY.b %>% rename(Year = wateryear,  Discharge = Bool_dis) %>% 
  complete(Year = seq(1946, 2020, by = 1)) # add gaps for missing years
```



```{r gwydir}
# remove years with too few data
gflow.WY.g <- gflow.WY %>% filter(nYd >= 347)
gflow.WY.g.short <- gflow.WY.g %>% filter(wateryear >= 1971)
M1 <- glm(l.Yarr_dis ~ wateryear, data = gflow.WY.g)
M1 <- glm(l.Yarr_dis ~ wateryear, data = gflow.WY.g.short)

# assumptions
par(mfrow = c(2, 2))
plot(M1) # use log
par(mfrow = c(1, 1)) 
ggplot(gflow.WY.g, aes(wateryear, l.Yarr_dis), na.rm = TRUE) +
  geom_point() 
# no obvious non-linear trends


# predict
gflow.WY.g$predict.Time.l <- predict(M1, gflow.WY.g)
# create a variable that is predicted log gwydir discharge over time
gflow.WY.g$predict.Time <- exp(gflow.WY.g$predict.Time.l)-1
plot(gflow.WY.g$Yarr_dis ~ gflow.WY.g$predict.Time)
plot(gflow.WY.g$predict.Time ~ gflow.WY.g$wateryear)

# predict
gflow.WY.g.short$predict.Time.l <- predict(M1, gflow.WY.g.short)
# create a variable that is predicted log gwydir discharge over time
gflow.WY.g.short$predict.Time <- exp(gflow.WY.g.short$predict.Time.l)-1
plot(gflow.WY.g.short$Yarr_dis ~ gflow.WY.g.short$predict.Time)
plot(gflow.WY.g.short$predict.Time ~ gflow.WY.g.short$wateryear)


# plot
ggemmeans.dis <- ggemmeans(M1, terms = c("wateryear[all]"))

plot(ggemmeans.dis, colors = "hero", ci = T, add.data = TRUE, jitter = 0, show.y.title = T, show.title = FALSE) + labs(x="Year",y="log Gwydir discharge (ML/day)",title="")


# test
summary(M1)
cor.test(gflow.WY.g$l.Yarr_dis, gflow.WY.g$wateryear)
cor.test(gflow.WY.g.short$l.Yarr_dis, gflow.WY.g.short$wateryear)

# calculate differences over time
100-(gflow.WY.g$predict.Time[length(gflow.WY.g$predict.Time)]/gflow.WY.g$predict.Time[1]*100)
# diff in last 50 years - since 1971, which is row 3
100-(gflow.WY.g$predict.Time[length(gflow.WY.g$predict.Time)]/gflow.WY.g$predict.Time[3]*100)

# calculate differences over time - without early data points
100-(gflow.WY.g.short$predict.Time[length(gflow.WY.g.short$predict.Time)]/gflow.WY.g.short$predict.Time[1]*100)

# prepare data for plotting
gflow.WY.gp <- gflow.WY.g %>% rename(Year = wateryear,  Discharge = Yarr_dis) %>% 
  complete(Year = seq(1930, 2020, by = 1)) # add gaps for missing years
```


```{r hattah}
# remove years with too few data
hflow.WY.h <- hflow.WY %>% filter(nNd >= 347)

M1 <- glm(new ~ wateryear, data = hflow.WY.h)

# assumptions
par(mfrow = c(2, 2))
plot(M1) # non-transformed data looks ok
par(mfrow = c(1, 1)) 
ggplot(hflow.WY.h, aes(wateryear, new), na.rm = TRUE) +
  geom_point() 
# no obvious non-linear trends


# predict
hflow.WY.h$predict.Time <- predict(M1, hflow.WY.h)
# create a variable that is predicted hattah discharge over time


# plot
ggemmeans.dis <- ggemmeans(M1, terms = c("wateryear[all]"))

plot(ggemmeans.dis, colors = "hero", ci = T, add.data = TRUE, jitter = 0, show.y.title = T, show.title = FALSE) + labs(x="Year",y="predicted Colignan discharge (ML/day)",title="")


# test
summary(M1)
cor.test(hflow.WY.h$new, hflow.WY.h$wateryear)


# calculate differences over time
100-(hflow.WY.h$predict.Time[length(hflow.WY.h$predict.Time)]/hflow.WY.h$predict.Time[1]*100)
# diff in last 50 years - since 1970, which is row 55
100-(hflow.WY.h$predict.Time[length(hflow.WY.h$predict.Time)]/hflow.WY.h$predict.Time[55]*100)

# prepare data for plotting
hflow.WY.hp <- hflow.WY.h %>% rename(Year = wateryear,  Discharge = new) %>% 
  complete(Year = seq(1913, 2020, by = 1)) # add gaps for missing years
```


```{r lowbidgee}
# remove years with too few data
lflow.WY.l <- lflow.WY %>% filter(nNd >= 347) %>% mutate(new.l = log(new + 1))

M1 <- glm(new.l ~ wateryear, data = lflow.WY.l)

# assumptions
par(mfrow = c(2, 2))
plot(M1) # use log
par(mfrow = c(1, 1)) 
ggplot(lflow.WY.l, aes(wateryear, new.l), na.rm = TRUE) +
  geom_point() 
# no obvious non-linear trends


# predict
lflow.WY.l$predict.Time.l <- predict(M1, lflow.WY.l)
# create a variable that is predicted log booligal discharge over time
lflow.WY.l$predict.Time <- exp(lflow.WY.l$predict.Time.l)-1
plot(lflow.WY.l$new ~ lflow.WY.l$predict.Time)
plot(lflow.WY.l$predict.Time ~ lflow.WY.l$wateryear)


# plot
ggemmeans.dis <- ggemmeans(M1, terms = c("wateryear[all]"))

plot(ggemmeans.dis, colors = "hero", ci = T, add.data = TRUE, jitter = 0, show.y.title = T, show.title = FALSE) + labs(x="Year",y="log predicted Redbank discharge (ML/day)",title="")


# test
summary(M1)
cor.test(lflow.WY.l$new.l, lflow.WY.l$wateryear)


# calculate differences over time
100-(lflow.WY.l$predict.Time[length(lflow.WY.l$predict.Time)]/lflow.WY.l$predict.Time[1]*100)
# diff in last 50 years - since 1970, which is row 55
100-(lflow.WY.l$predict.Time[length(lflow.WY.l$predict.Time)]/lflow.WY.l$predict.Time[55]*100)

# prepare data for plotting
lflow.WY.lp <- lflow.WY.l %>% rename(Year = wateryear,  Discharge = new) %>% 
  complete(Year = seq(1915, 2020, by = 1)) # add gaps for missing years
```


```{r macquarie}
# remove years with too few data
mflow.WY.m <- mflow.WY %>% filter(nOd >= 347)

M1 <- glm(l.ox_dis ~ wateryear, data = mflow.WY.m)

# assumptions
par(mfrow = c(2, 2))
plot(M1) #  log looks somewhat better?
par(mfrow = c(1, 1)) 
ggplot(mflow.WY.m, aes(wateryear, l.ox_dis), na.rm = TRUE) +
  geom_point() 
# without transformation, the trend looks somewhat non-linear, decreases massively from 1955 to 1980 then decreases more slowly
# log is better


# predict
mflow.WY.m$predict.Time.l <- predict(M1, mflow.WY.m)
# create a variable that is predicted log booligal discharge over time
mflow.WY.m$predict.Time <- exp(mflow.WY.m$predict.Time.l)-1
plot(mflow.WY.m$ox_dis ~ mflow.WY.m$predict.Time)
plot(mflow.WY.m$predict.Time ~ mflow.WY.m$wateryear)


# plot
ggemmeans.dis <- ggemmeans(M1, terms = c("wateryear[all]"))

plot(ggemmeans.dis, colors = "hero", ci = T, add.data = TRUE, jitter = 0, show.y.title = T, show.title = FALSE) + labs(x="Year",y="Log Oxley discharge (ML/day)",title="")


# test
summary(M1)
cor.test(mflow.WY.m$l.ox_dis, mflow.WY.m$wateryear)


# calculate differences over time
100-(mflow.WY.m$predict.Time[length(mflow.WY.m$predict.Time)]/mflow.WY.m$predict.Time[1]*100)
# diff in last 50 years - since 1970, which is row 13
100-(lflow.WY.l$predict.Time[length(lflow.WY.l$predict.Time)]/lflow.WY.l$predict.Time[13]*100)

# prepare data for plotting
mflow.WY.mp <- mflow.WY.m %>% rename(Year = wateryear,  Discharge = ox_dis) %>% 
  complete(Year = seq(1955, 2018, by = 1)) # add gaps for missing years
```


```{r paroo}
# remove years with too few data
pflow.WY.p <- pflow.WY %>% filter(nWd >= 347)

M1 <- glm(l.will_dis ~ wateryear, data = pflow.WY.p)

# assumptions
par(mfrow = c(2, 2))
plot(M1) # log is better
par(mfrow = c(1, 1)) 
ggplot(pflow.WY.p, aes(wateryear, will_dis), na.rm = TRUE) +
  geom_point() 
# no obvious non-linear trends


# predict
pflow.WY.p$predict.Time.l <- predict(M1, pflow.WY.p)
# create a variable that is predicted log booligal discharge over time
pflow.WY.p$predict.Time <- exp(pflow.WY.p$predict.Time.l)-1
plot(pflow.WY.p$will_dis ~ pflow.WY.p$predict.Time)
plot(pflow.WY.p$predict.Time ~ pflow.WY.p$wateryear)


# plot
ggemmeans.dis <- ggemmeans(M1, terms = c("wateryear[all]"))

plot(ggemmeans.dis, colors = "hero", ci = T, add.data = TRUE, jitter = 0, show.y.title = T, show.title = FALSE) + labs(x="Year",y="log Willara Xing discharge (ML/day)",title="")


# test
summary(M1)
cor.test(pflow.WY.p$l.will_dis, pflow.WY.p$wateryear, na.rm = TRUE)
# not significant


# calculate differences over time
100-(pflow.WY.p$predict.Time[length(pflow.WY.p$predict.Time)]/pflow.WY.p$predict.Time[1]*100)
# diff in last 50 years - same as above, both are less than 50 years

# prepare data for plotting
pflow.WY.pp <- pflow.WY.p %>% rename(Year = wateryear,  Discharge = will_dis) %>% 
  complete(Year = seq(1976, 2020, by = 1)) # add gaps for missing years
```


```{r riverland}
# remove years with too few data
rflow.WY.r <- rflow.WY %>% filter(nFd >= 347)
rmflow.WY.r <- rmflow.WY %>% filter(nFd >= 6)
# combine monthly and daily data
rflow.WY.all <- left_join(rmflow.WY.r,rflow.WY.r,  by = "wateryear") %>% select(wateryear, Flows_dis.x, Flows_dis.y)
# join the columns
rflow.WY.all <- rflow.WY.all %>% 
  mutate(new = case_when(!is.na(Flows_dis.y) ~ Flows_dis.y, TRUE ~ Flows_dis.x),
         new.l = log(new + 1))

M1 <- glm(new.l ~ wateryear, data = rflow.WY.all)

# assumptions
par(mfrow = c(2, 2))
plot(M1) # use log
par(mfrow = c(1, 1)) 
ggplot(rflow.WY.all, aes(wateryear, new.l), na.rm = TRUE) +
  geom_point() 
# roughly linear


# predict
rflow.WY.all$predict.Time.l <- predict(M1, rflow.WY.all)
# create a variable that is predicted log booligal discharge over time
rflow.WY.all$predict.Time <- exp(rflow.WY.all$predict.Time.l)-1
plot(rflow.WY.all$new ~ rflow.WY.all$predict.Time)
plot(rflow.WY.all$predict.Time ~ rflow.WY.all$wateryear)


# plot
ggemmeans.dis <- ggemmeans(M1, terms = c("wateryear[all]"))

plot(ggemmeans.dis, colors = "hero", ci = T, add.data = TRUE, jitter = 0, show.y.title = T, show.title = FALSE) + labs(x="Year",y="log Flows to SA discharge (ML/day)",title="")


# test
summary(M1)
cor.test(rflow.WY.all$new, rflow.WY.all$wateryear)


# calculate differences over time
100-(rflow.WY.all$predict.Time[length(rflow.WY.all$predict.Time)]/rflow.WY.all$predict.Time[1]*100)
# diff in last 50 years - since 1970, which is row 69
100-(rflow.WY.all$predict.Time[length(rflow.WY.all$predict.Time)]/rflow.WY.all$predict.Time[69]*100)

# prepare data for plotting
rflow.WY.rp <- rflow.WY.all %>% rename(Year = wateryear,  Discharge = new) %>% 
  complete(Year = seq(1976, 2020, by = 1)) # add gaps for missing years
```


```{r yantabulla}
# remove years with too few data
yflow.WY.y <- yflow.WY %>% filter(nTd >= 347)

M1 <- glm(l.Tur_dis ~ wateryear, data = yflow.WY.y)

# assumptions
par(mfrow = c(2, 2))
plot(M1) # use log
par(mfrow = c(1, 1)) 
ggplot(yflow.WY.y, aes(wateryear, l.Tur_dis), na.rm = TRUE) +
  geom_point() 
# no obvious non-linear trends


# predict
yflow.WY.y$predict.Time.l <- predict(M1, yflow.WY.y)
# create a variable that is predicted log turra discharge over time
yflow.WY.y$predict.Time <- exp(yflow.WY.y$predict.Time.l)-1
plot(yflow.WY.y$Tur_dis ~ yflow.WY.y$predict.Time)
plot(yflow.WY.y$predict.Time ~ yflow.WY.y$wateryear)


# plot
ggemmeans.dis <- ggemmeans(M1, terms = c("wateryear[all]"))

plot(ggemmeans.dis, colors = "hero", ci = T, add.data = TRUE, jitter = 0, show.y.title = T, show.title = FALSE) + labs(x="Year",y="log Turra discharge (ML/day)",title="")


# test
summary(M1)
cor.test(yflow.WY.y$l.Tur_dis, yflow.WY.y$wateryear)
# not significant


# calculate differences over time
100-(yflow.WY.y$predict.Time[length(yflow.WY.y$predict.Time)]/yflow.WY.y$predict.Time[1]*100)


# prepare data for plotting
yflow.WY.yp <- yflow.WY.y %>% rename(Year = wateryear,  Discharge = Tur_dis) %>% 
  complete(Year = seq(1994, 2020, by = 1)) # add gaps for missing years
```


```{r cumbung inundation}
M1 <- glm(predict ~ year, data = inun)

# assumptions
par(mfrow = c(2, 2))
plot(M1) # looks fine untransformed 
par(mfrow = c(1, 1)) 
ggplot(inun, aes(year, predict), na.rm = TRUE) +
  geom_point() 
# no obvious non-linear trends


# predict
inun$predict.Time <- predict(M1, inun)
# create a variable that is predicted inundation area over time


# plot
ggemmeans.dis <- ggemmeans(M1, terms = c("year[all]"))

plot(ggemmeans.dis, colors = "hero", ci = T, add.data = TRUE, jitter = 0, show.y.title = T, show.title = FALSE) + labs(x="Year",y="Predicted Inundation area (km2)",title="")


# test
summary(M1)
cor.test(inun$predict, inun$year)


# calculate differences over time
100-(inun$predict.Time[length(inun$predict.Time)]/inun$predict.Time[1]*100)
# since 1970 (50 years ago)
100-(inun$predict.Time[length(inun$predict.Time)]/inun$predict.Time[22]*100)

#exp(coef(year))

# prepare data for plotting
inun.p <- inun %>% rename(Year = year,  Discharge = predict) %>% 
  complete(Year = seq(1946, 2020, by = 1)) # add gaps for missing years
```

# Flood time series analysis - generalised linear models


```{r cumbung}
M1 <- glm(Flood_days ~ Year, data = cflood)

# assumptions
par(mfrow = c(2, 2))
plot(M1) # OK, log makes it much worse
par(mfrow = c(1, 1)) 
ggplot(cflood, aes(Year, Flood_days), na.rm = TRUE) +
  geom_point() 
# no obvious non-linear trends


# predict
cflood$predict.Time <- predict(M1, cflood)
# create a variable that is predicted four mile discharge over time


# plot
ggemmeans.dis <- ggemmeans(M1, terms = c("Year[all]"))

plot(ggemmeans.dis, colors = "hero", ci = T, add.data = TRUE, jitter = 0, show.y.title = T, show.title = FALSE) + labs(x="Year",y="Predicted Four mile weir # flood days",title="")


# test
summary(M1)
cor.test(cflood$Flood_days, cflood$Year)


# calculate differences over time
100-(cflood$predict.Time[length(cflood$predict.Time)]/cflood$predict.Time[1]*100)
# diff in last 50 years - since 1970, which is row 25
100-(cflood$predict.Time[length(cflood$predict.Time)]/cflood$predict.Time[25]*100)

# prepare data for plotting
cflood <- cflood %>% complete(Year = seq(1946, 2020, by = 1)) # add gaps for missing years
```



```{r booligal}
# remove years with too few data
cflow.WY.b <- cflow.WY %>% filter(nBl >= 347)

M1 <- glm(floodb ~ wateryear, data = cflow.WY.b)

# assumptions
par(mfrow = c(2, 2))
plot(M1) # roughly normal, log transformed is worse
par(mfrow = c(1, 1)) 
ggplot(cflow.WY.b, aes(wateryear, floodb), na.rm = TRUE) +
  geom_point() 
# no obvious non-linear trends


# predict
cflow.WY.b$predict <- predict(M1, cflow.WY.b)


# plot
ggemmeans.dis <- ggemmeans(M1, terms = c("wateryear[all]"))

plot(ggemmeans.dis, colors = "hero", ci = T, add.data = TRUE, jitter = 0, show.y.title = T, show.title = FALSE) + labs(x="Year",y="Number of flood days at Booligal",title="")


# test
summary(M1)
cor.test(cflow.WY.b$floodb, cflow.WY.b$wateryear)


# calculate differences over time
100-(cflow.WY.b$predict[length(cflow.WY.b$predict)]/cflow.WY.b$predict[1]*100)
# diff in last 50 years - since 1970, which is row 22
100-(cflow.WY.b$predict[length(cflow.WY.b$predict)]/cflow.WY.b$predict[22]*100)

# prep for plotting
bflood <- cflow.WY.b %>% rename(Year = wateryear,  Flood_days = floodb) %>% 
  complete(Year = seq(1946, 2020, by = 1)) # add gaps for missing years
```



```{r gwydir}
# remove years with too few data
gflow.WY.g <- gflow.WY %>% filter(nYl >= 347)

M1 <- glm(flood ~ wateryear, data = gflow.WY.g)
# run with either model, with or without 1930
M1 <- glm(flood ~ wateryear, data = gflow.WY.g %>% filter(wateryear >= 1971))
# note, these results are very different depending on whether the 1930s data point is included

# assumptions
par(mfrow = c(2, 2))
plot(M1) # ok, log looks worse
par(mfrow = c(1, 1)) 
ggplot(gflow.WY.g, aes(wateryear, flood), na.rm = TRUE) +
  geom_point() 
# no obvious non-linear trends


# predict
gflow.WY.g$predict <- predict(M1, gflow.WY.g)


# plot
ggemmeans.dis <- ggemmeans(M1, terms = c("wateryear[all]"))

plot(ggemmeans.dis, colors = "hero", ci = T, add.data = TRUE, jitter = 0, show.y.title = T, show.title = FALSE) + labs(x="Year",y="Number of flood days at Gwydir",title="")


# test
summary(M1)
cor.test(gflow.WY.g$flood, gflow.WY.g$wateryear)
gflow.WY.g.short <- gflow.WY.g %>% filter(wateryear >= 1971)
cor.test(gflow.WY.g.short$flood, gflow.WY.g.short$wateryear)


# calculate differences over time
100-(gflow.WY.g$predict[length(gflow.WY.g$predict)]/gflow.WY.g$predict[1]*100)
# diff in last 50 years - since 1971, which is row 2
100-(gflow.WY.g$predict[length(gflow.WY.g$predict)]/gflow.WY.g$predict[2]*100)

# calculate differences over time - shorter dataset model
100-(gflow.WY.g.short$predict[length(gflow.WY.g.short$predict)]/gflow.WY.g.short$predict[1]*100)

# prep for plotting
gflood <- gflow.WY.g %>% rename(Year = wateryear,  Flood_days = flood) %>% 
  complete(Year = seq(1930, 2020, by = 1)) # add gaps for missing years
```


```{r hattah}

M1 <- glm(Flood_days ~ Year, data = hflood)

# assumptions
par(mfrow = c(2, 2))
plot(M1) # non-transformed data looks ok, transformed is worse
par(mfrow = c(1, 1)) 
ggplot(hflood, aes(Year, Flood_days), na.rm = TRUE) +
  geom_point() 
# no obvious non-linear trends


# predict
hflood$predict.Time <- predict(M1, hflood)
# create a variable that is predicted hattah discharge over time


# plot
ggemmeans.dis <- ggemmeans(M1, terms = c("Year[all]"))

plot(ggemmeans.dis, colors = "hero", ci = T, add.data = TRUE, jitter = 0, show.y.title = T, show.title = FALSE) + labs(x="Year",y="predicted Colignan flood days",title="")


# test
summary(M1)
cor.test(hflood$Flood_days, hflood$Year)


# calculate differences over time
100-(hflood$predict.Time[length(hflood$predict.Time)]/hflood$predict.Time[1]*100)
# diff in last 50 years - since 1970, which is row 58
100-(hflood$predict.Time[length(hflood$predict.Time)]/hflood$predict.Time[58]*100)

# prep for plotting
hflood <- hflood %>% complete(Year = seq(1913, 2020, by = 1)) # add gaps for missing years
```



```{r lowbidgee - predicted}
# comment out one model at a time, with or without 1956
M1 <- glm(log(Flood_days+1) ~ Year, data = lflood)
M1 <- glm(log(Flood_days+1) ~ Year, data = lflood %>% filter(Year >= 1973))

# assumptions
par(mfrow = c(2, 2))
plot(M1) # use log
par(mfrow = c(1, 1)) 
ggplot(lflood, aes(Year, log(Flood_days+1)), na.rm = TRUE) +
  geom_point() 
# no obvious non-linear trends


# predict
lflood$predict.Time.l <- predict(M1, lflood)
# create a variable that is predicted log booligal discharge over time
lflood$predict.Time <- exp(lflood$predict.Time.l)-1
plot(lflood$Flood_days ~ lflood$predict.Time)
plot(lflood$predict.Time ~ lflood$Year)


# plot
ggemmeans.dis <- ggemmeans(M1, terms = c("Year[all]"))

plot(ggemmeans.dis, colors = "hero", ci = T, add.data = TRUE, jitter = 0, show.y.title = T, show.title = FALSE) + labs(x="Year",y="Predicted number of flood days at Lowbidgee",title="")


# test
summary(M1)
lflood.short <- lflood %>% filter(Year >= 1973)
cor.test(log(lflood$Flood_days + 1), lflood$Year)
cor.test(log(lflood.short$Flood_days + 1), lflood.short$Year)

# calculate differences over time
100-(lflood$predict.Time[length(lflood$predict.Time)]/lflood$predict.Time[1]*100)
# diff in last 50 years - since 1973, which is row 3
100-(hflood$predict.Time[length(hflood$predict.Time)]/hflood$predict.Time[2]*100)

# calculate differences over time - without 1956
100-(lflood.short$predict.Time[length(lflood.short$predict.Time)]/lflood.short$predict.Time[1]*100)

# prep for plotting
lflood <- lflood %>% complete(Year = seq(1956, 2020, by = 1)) # add gaps for missing years
```



```{r lowbidgee - raw}
# remove years with too few data
lflow.WY.l <- lflow.WY %>% filter(nRl >= 347)

M1 <- glm(flood ~ wateryear, data = lflow.WY.l)

# assumptions
par(mfrow = c(2, 2))
plot(M1) # log worse, neither great, probably due to zero inflation
par(mfrow = c(1, 1)) 
ggplot(lflow.WY.l, aes(wateryear, flood), na.rm = TRUE) +
  geom_point() 
# no obvious non-linear trends


# predict
lflow.WY.l$predict.Time <- predict(M1, lflow.WY.l)


# plot
ggemmeans.dis <- ggemmeans(M1, terms = c("wateryear[all]"))

plot(ggemmeans.dis, colors = "hero", ci = T, add.data = TRUE, jitter = 0, show.y.title = T, show.title = FALSE) + labs(x="Year",y="Number of flood days at Redbank",title="")


# test
summary(M1)
cor.test(lflow.WY.l$flood, lflow.WY.l$wateryear)


# calculate differences over time
100-(lflow.WY.l$predict.Time[length(lflow.WY.l$predict.Time)]/lflow.WY.l$predict.Time[1]*100)

# prep for plotting
lflood_red <- lflow.WY.l %>% rename(Year = wateryear,  Flood_days = flood) %>% 
  complete(Year = seq(1976, 2020, by = 1)) # add gaps for missing years
```


```{r macquarie}
# remove years with too few data
mflow.WY.m <- mflow.WY %>% filter(nOl >= 347)
mflow.WY.m.short <- mflow.WY.m %>% filter(wateryear >= 1974)

# comment out one at a time, with and without few early data
M1 <- glm(flood ~ wateryear, data = mflow.WY.m)
M1 <- glm(flood ~ wateryear, data = mflow.WY.m.short)

# assumptions
par(mfrow = c(2, 2))
plot(M1) #  ok + little zero inflation
par(mfrow = c(1, 1)) 
ggplot(mflow.WY.m, aes(wateryear, flood), na.rm = TRUE) +
  geom_point() 


# predict
mflow.WY.m$predict.Time <- predict(M1, mflow.WY.m)
mflow.WY.m.short$predict.Time <- predict(M1, mflow.WY.m.short)

# plot
ggemmeans.dis <- ggemmeans(M1, terms = c("wateryear[all]"))

plot(ggemmeans.dis, colors = "hero", ci = T, add.data = TRUE, jitter = 0, show.y.title = T, show.title = FALSE) + labs(x="Year",y="Number of flood days at Oxley",title="")


# test
summary(M1)
cor.test(mflow.WY.m$l.ox_dis, mflow.WY.m$wateryear)
cor.test(mflow.WY.m.short$l.ox_dis, mflow.WY.m.short$wateryear)

# calculate differences over time
100-(mflow.WY.m$predict.Time[length(mflow.WY.m$predict.Time)]/mflow.WY.m$predict.Time[1]*100)
# diff in last 50 years - since 1974, which is row 2
100-(mflow.WY.m$predict.Time[length(mflow.WY.m$predict.Time)]/mflow.WY.m$predict.Time[2]*100)

# calculate differences over time - without 1956
100-(mflow.WY.m.short$predict.Time[length(mflow.WY.m.short$predict.Time)]/mflow.WY.m.short$predict.Time[1]*100)

# prep for plotting
mflood <- mflow.WY.m %>% rename(Year = wateryear,  Flood_days = flood) %>% 
  complete(Year = seq(1955, 2020, by = 1)) # add gaps for missing years
```


```{r paroo}
# remove years with too few data
pflow.WY.p <- pflow.WY %>% filter(nWl >= 347)

M1 <- glm(flood ~ wateryear, data = pflow.WY.p)

# assumptions
par(mfrow = c(2, 2))
plot(M1) # ok, zero inflated
par(mfrow = c(1, 1)) 
ggplot(pflow.WY.p, aes(wateryear, flood), na.rm = TRUE) +
  geom_point() 
# no obvious non-linear trends


# predict
pflow.WY.p$predict <- predict(M1, pflow.WY.p)


# plot
ggemmeans.dis <- ggemmeans(M1, terms = c("wateryear[all]"))

plot(ggemmeans.dis, colors = "hero", ci = T, add.data = TRUE, jitter = 0, show.y.title = T, show.title = FALSE) + labs(x="Year",y="flood days at Willara Xing",title="")


# test
summary(M1)
cor.test(pflow.WY.p$flood, pflow.WY.p$wateryear)
# not significant


# calculate differences over time
100-(pflow.WY.p$predict[length(pflow.WY.p$predict)]/pflow.WY.p$predict[1]*100)

# prep for plotting
pflood <- pflow.WY.p %>% rename(Year = wateryear,  Flood_days = flood) %>% 
  complete(Year = seq(1976, 2020, by = 1)) # add gaps for missing years
```


```{r riverland}
# remove years with too few data
rflow.WY.r <- rflow.WY %>% filter(nRl >= 347)

M1 <- glm(flood ~ wateryear, data = rflow.WY.r)

# assumptions
par(mfrow = c(2, 2))
plot(M1) # a couple outliers, but worse with transformation
par(mfrow = c(1, 1)) 
ggplot(rflow.WY.r, aes(wateryear, flood), na.rm = TRUE) +
  geom_point() 
# no obvious non-linear trends


# predict
rflow.WY.r$predict <- predict(M1, rflow.WY.r)


# plot
ggemmeans.dis <- ggemmeans(M1, terms = c("wateryear[all]"))

plot(ggemmeans.dis, colors = "hero", ci = T, add.data = TRUE, jitter = 0, show.y.title = T, show.title = FALSE) + labs(x="Year",y="Renmark Domestic PS flood days",title="")


# test
summary(M1)
cor.test(rflow.WY.r$flood, rflow.WY.r$wateryear)


# calculate differences over time
100-(rflow.WY.r$predict[length(rflow.WY.r$predict)]/rflow.WY.r$predict[1]*100)

# prep for plotting
rflood <- rflow.WY.r %>% rename(Year = wateryear,  Flood_days = flood) %>% 
  complete(Year = seq(1978, 2020, by = 1)) # add gaps for missing years
```


```{r yantabulla}
# remove years with too few data
yflow.WY.y <- yflow.WY %>% filter(nTl >= 347)

M1 <- glm(flood ~ wateryear, data = yflow.WY.y)

# assumptions
par(mfrow = c(2, 2))
plot(M1) # ok
par(mfrow = c(1, 1)) 
ggplot(yflow.WY.y, aes(wateryear, flood), na.rm = TRUE) +
  geom_point() 
# no obvious non-linear trends


# predict
yflow.WY.y$predict <- predict(M1, yflow.WY.y)


# plot
ggemmeans.dis <- ggemmeans(M1, terms = c("wateryear[all]"))

plot(ggemmeans.dis, colors = "hero", ci = T, add.data = TRUE, jitter = 0, show.y.title = T, show.title = FALSE) + labs(x="Year",y="log Turra discharge (ML/day)",title="")


# test
summary(M1)
cor.test(yflow.WY.y$flood, yflow.WY.y$wateryear)
# not significant


# calculate differences over time
100-(yflow.WY.y$predict[length(yflow.WY.y$predict)]/yflow.WY.y$predict[1]*100)

#exp(coef(wateryear))

# prep for plotting
yflood <- yflow.WY.y %>% rename(Year = wateryear,  Flood_days = flood) %>% 
  complete(Year = seq(1994, 2020, by = 1)) # add gaps for missing years
```



# Plots

```{r plot functions}
plot_dis <- function(dataset, title){
  ggplot(dataset, aes(x = Year, y = Discharge)) +
    geom_point() +
  stat_smooth(method = lm, aes(ymin = ifelse(..ymin.. < 0, 0, ..ymin..))) + # make sure confidence intervals are not cut off
  labs(x="",y="",title=title) +
  theme_bw() +
  geom_line(size = 1.2) +
  #geom_col(aes(x = Year, y = missing_dis), width = 31, fill = "#f0f0f0") + # not working yet
  theme(text = element_text(size = 32), # change font size
        axis.text = element_text(size = 28),
        plot.margin = margin(10,20,10,10), # give extra space on RHS
        plot.title = element_text(hjust = 0.5)) # centre title
}

plot_flood <- function(dataset, title){
  ggplot(dataset, aes(x = Year, y = Flood_days)) +
    geom_point()  + 
  stat_smooth(method = lm, aes(ymin = ifelse(..ymin.. < 0, 0, ..ymin..))) +
  labs(x="",y="",title=title) +
  theme_bw() +
  geom_line(size = 1.2) +
  #geom_col(aes(x = Year, y = missing_dis), width = 31, fill = "#f0f0f0") + # not working yet
  theme(text = element_text(size = 32), # change font size
        axis.text = element_text(size = 28),
        plot.margin = margin(10,20,10,10), # give extra space on RHS
        plot.title = element_text(hjust = 0.5)) # centre title
}
```

```{r Correlation plots}
# export at 2100 x 2900
corr.plot <- ggarrange(c.flow, h.flow, l.flow, c.flood, h.flood, l.flood, c.floodday, h.floodday, l.floodday,
          ncol = 3, nrow = 3)
corr.plot
```


```{r make discharge plots}
bd <- plot_dis(cflow.WY.bp, "Booligal Wetlands*") +
  scale_x_continuous(breaks = c(1945,1970,2000,2020)) +
  scale_y_continuous(limits = c(0, 4000))
bd
cd <- plot_dis(cflow.WY.cp, "Great Cumbung Swamp*") + 
  scale_x_continuous(breaks = c(1945,1970,2000,2020))  +
  labs(x="",y="Flow (ML/day)") +
  theme(text = element_text(size = 26),
        axis.title = element_text(size = 60))  +
  scale_y_continuous(limits = c(0, 800))
cd
gd <- plot_dis(gflow.WY.gp, "Gwydir Wetlands*") +
  scale_x_continuous(breaks = c(1930,1960,1990,2020))  +
  scale_y_continuous(limits = c(0, 6000)) 
gd
hd <- plot_dis(hflow.WY.hp, "Hattah Lakes*") +
  scale_x_continuous(breaks = c(1910,1950,1990,2020)) +
  labs(x="",y="Flow (ML/day)") +
  theme(axis.title = element_text(size = 60))  +
  scale_y_continuous(limits = c(0, 90000))
hd
ld <- plot_dis(lflow.WY.lp, "Lowbidgee Wetlands*") +
  scale_x_continuous(breaks = c(1910,1950,1990,2020))  +
  scale_y_continuous(limits = c(0, 16500))
ld
md <- plot_dis(mflow.WY.mp, "Macquarie Marshes*") +
  scale_x_continuous(breaks = c(1960,1980,2000,2018)) +
  labs(x="",y="") +
  theme(axis.title = element_text(size = 60))  +
  scale_y_continuous(limits = c(0, 3600))
md
pd <- plot_dis(pflow.WY.pp, "Paroo Overflow Swamps") +
  labs(x="Year",y="") +
  theme(text = element_text(size = 26),
        axis.title = element_text(size = 60))  +
  scale_y_continuous(limits = c(0, 4500))
pd
rd <- plot_dis(rflow.WY.rp, "Murray Riverlands*") +
  labs(x="Year",y="Flow (ML/day)") +
  theme(axis.title = element_text(size = 60)) +
  scale_x_continuous(breaks = c(1900,1930,1960,1990,2020))  +
  scale_y_continuous(limits = c(0, 160000))
rd
yd <- plot_dis(yflow.WY.yp, "Yantabulla Swamp") + 
  labs(x="Year",y="") +
  theme(axis.title = element_text(size = 60)) +
  scale_x_continuous(breaks = c(1994,2000,2010,2020))  +
  scale_y_continuous(limits = c(0, 3000))
yd

```

```{r make flood plots}
bf <- plot_flood(bflood, "Booligal Wetlands*") +
  scale_x_continuous(breaks = c(1945,1970,2000,2020))+
  scale_y_continuous(limits = c(0, 280))
bf
cf <- plot_flood(cflood, "Great Cumbung Swamp*") + 
  scale_x_continuous(breaks = c(1945,1970,2000,2020))  +
  labs(x="",y="Number of flood days") +
  theme(text = element_text(size = 30),
        axis.title = element_text(size = 60))+
  scale_y_continuous(limits = c(0, 330))
cf
gf <- plot_flood(gflood, "Gwydir Wetlands") +
  scale_x_continuous(breaks = c(1930,1960,1990,2020))+
  scale_y_continuous(limits = c(0, 30))
gf
hf <- plot_flood(hflood, "Hattah Lakes*") +
  scale_x_continuous(breaks = c(1910,1950,1990,2020)) +
  labs(x="",y="Number of flood days") +
  theme(axis.title = element_text(size = 60))+
  scale_y_continuous(limits = c(0, 180))
hf
lf <- plot_flood(lflood, "Lowbidgee Wetlands*") +
  scale_x_continuous(breaks = c(1960,1980,2000,2020))+
  scale_y_continuous(limits = c(0, 210))
lf


mf <- plot_flood(mflood, "Macquarie Marshes*") +
  scale_x_continuous(breaks = c(1960,1980,2000,2018))+
  scale_y_continuous(limits = c(0, 360))
mf
pf <- plot_flood(pflood, "Paroo Overflow Swamps") +
  labs(x="Year",y="") +
  theme(text = element_text(size = 30),
        axis.title = element_text(size = 60))+
  scale_y_continuous(limits = c(0, 60))
pf
rf <- plot_flood(rflood, "Murray Riverlands") +
  labs(x="Year",y="Number of flood days") +
  theme(axis.title = element_text(size = 60)) +
  scale_x_continuous(breaks = c(1900,1930,1960,1990,2020))+
  scale_y_continuous(limits = c(0, 130))
rf
yf <- plot_flood(yflood, "Yantabulla Swamp")  +
  labs(x="Year",y="") +
  theme(axis.title = element_text(size = 60)) +
  scale_x_continuous(breaks = c(1994,2000,2010,2020))+
  scale_y_continuous(limits = c(0, 80))
yf
```

```{r Plots for presentation}
# flows
plot_dis(cflow.WY.cp, "") + 
  scale_x_continuous(breaks = c(1945,1970,2000,2020))  +
  labs(x="Year",y="Flow (ML/day)") +
  theme(text = element_text(size = 26),
        axis.title = element_text(size = 35),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank())  +
  scale_y_continuous(limits = c(0, 800))

# floods
plot_flood(cflood, "") + 
  scale_x_continuous(breaks = c(1945,1970,2000,2020))  +
  labs(x="Year",y="Flood days") +
  theme(text = element_text(size = 30),
        axis.title = element_text(size = 35),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank())+
  scale_y_continuous(limits = c(0, 330))

# inundation
ip <- plot_dis(inun.p, "") +
  scale_x_continuous(breaks = c(1946,1970,2000,2020)) +
  scale_y_continuous(limits = c(0, 70))  +
  labs(x="Year",y="Inundation (km2)") +
  theme(text = element_text(size = 30),
        axis.title = element_text(size = 35),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank())
ip
```


```{r combine discharge plots}
# export at 2100 x 2800 
# add asterisks for significance
# order in terms of significance or standard order

flow <- ggarrange(cd, bd, gd,hd,ld, md,  rd, pd, yd,  
          ncol = 3, nrow = 3, align = "hv")
flow

# if trying to remove space between plots (but cuts off axis numbers)
#flow <- ggarrange(
#  cd, ggplot() + theme_void(),  bd,ggplot() + theme_void(),  gd, ggplot() + #theme_void(),hd, ggplot() + theme_void(),  ld, ggplot() + theme_void(),  md, ggplot() #+ theme_void(),rd, ggplot() + theme_void(),  pd, ggplot() + theme_void(),  yd,
#  nrow = 3, ncol = 6, align = "hv", widths = c(1, -0.01, 1, -0.01, 1, -0.01))
#flow


# if only one title on each axis

#annotate_figure(bird, bottom = text_grob("Year                          Year                         Year",  size = 60, just = "bottom"), left = text_grob("Discharge (ML/day)         Discharge (ML/day)         Discharge (ML/day)", size = 60, rot = 90)) # rotate 90 degrees
# just = "bottom" helps text not be cut off
```

```{r combine flood plots}
flood <- ggarrange(cf, bf, gf, hf,lf, mf, rf,pf,yf,   
          ncol = 3, nrow = 3, align = "hv")
flood
```

