---
title: "Cumbung Waterbird Data Preparation"
author: "Dana Lanceman"
format: html
editor: visual
---

# Cumbung Waterbird Data Preparation

This document is for preparing the most recently updated Aerial Waterbird Survey waterbird data in the Great Cumbung Swamp for data analysis.

## 1. Set up

Load packages.

```{r Packages}
library(tidyverse)
library(here)
library(readxl)
```

Load data file.

```{r Data}
raw <- read_xlsx(here("raw/Cumbung_waterbirds_2008-2024_20251003.xlsx"))
```

## 2. Data manipulation

Subset data for only 2010-2024 (2008 data were from a different program so aren't necessarily comparable).

Also remove the columns we don't need.

Separate into count 1 (ReplicateC = NA or 1) and count 2 (ReplicateC = 2) as we can't use these data together as they're non-independent.

```{r Subset_select}
raw2 <- raw %>% subset(SurvYear %in% c(2010:2024)) %>% 
  select(SurvYear, Name, MDBWetland, Subwetland, CommonName, Scientific, ReplicateC, Count, Nest, Broods, BreedingIn)

count1 <- raw2 %>% subset(ReplicateC != 2)

count2 <- raw2 %>% subset(ReplicateC == 2)
```

### Annual totals

Calculate total annual waterbird counts. Calculate for both count 1 and count 2, but count 2 data are likely incomplete so we can't use them.

```{r Annual_totals}
Annual_total1 <- count1 %>% group_by(SurvYear) %>% 
  summarise(Count = sum(Count), Nest = sum(Nest), Broods = sum(Broods), Breeding = sum(BreedingIn))

mean(Annual_total1$Count)
sd(Annual_total1$Count)/sqrt(15)

Annual_total2 <- count2 %>% group_by(SurvYear) %>% 
  summarise(Count = sum(Count), Nest = sum(Nest), Broods = sum(Broods), Breeding = sum(BreedingIn))
```

Count 2 data are incomplete, so from now on only use Count 1 data.

### Species totals

Also sum the total count of each species for each year.

Also sum the mean and max counts for each species over all years.

```{r Species_totals}
Species_total1 <- count1 %>% group_by(SurvYear, Scientific, CommonName) %>% 
  summarise(Count = sum(Count), Nest = sum(Nest), Broods = sum(Broods), Breeding = sum(BreedingIn))

SpeciesStats <- Species_total1 %>% group_by(Scientific, CommonName) %>% 
  summarise(meanCount = mean(Count), maxCount = max(Count))
```

### Exports

Export the above dataframes so I can import them for other analyses.

```{r Export}
write_csv(count1, file = here("raw/WaterbirdSubset/CumbungWaterbirdCount1.csv"))

write_csv(Annual_total1, file = here("raw/WaterbirdSubset/CumbungWaterbirdTotalCount1.csv"))

write_csv(Species_total1, file = here("raw/WaterbirdSubset/CumbungWaterbirdSpeciesCount1.csv"))

write_csv(SpeciesStats, file = here("raw/WaterbirdSubset/CumbungWaterbirdSpeciesStats.csv"))
```
