---
title: "Waterbirds"
author: "Dana Lanceman"
date: "8/23/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r get packages}
#library(here)
library(tidyverse)
library(readxl)
library(viridis)
library(ggpubr)
library(lubridate)
library(padr)
library(ggeffects)
library(scales)
```

Load waterbird datasets we created in our other quarto document (WaterbirdDataPrep). Also import a dataset derived from Wetlands International data (https://wpp.wetlands.org/explore) that includes 1% population thresholds for each of our waterbirds. I added the species data onto this dataset in excel (because R was not appending properly). Also import flow and inundation data.

```{r get data}
waterbird <- read.csv("raw/WaterbirdSubset/CumbungWaterbirdTotalCount1.csv") %>% 
  mutate(ln.bird = log(Count)) %>% 
  rename(waterYear = SurvYear) # changing this to waterYear so matches flow column name

waterbirdSp <- read.csv("raw/WaterbirdSubset/CumbungWaterbirdSpeciesCount1.csv")

waterbirdSpStats <- read.csv("raw/WaterbirdSubset/CumbungWaterbirdSpeciesStats.csv")

wbirdpop <- read_xlsx("raw/Cumbung_Birds_Pop_Status.xlsx")

flow <- read_csv("raw/fourmile_flows_412194_daily.csv") %>% 
  mutate(Date = as.Date(Date, format = "%d/%m/%Y")) %>% 
  filter(!(Quality_D %in% c(255, 201, 180))) %>% 
  select(Date,Discharge) %>% 
  pad()

# add water year column, for Australian water year, July to June
# July 2009 to June 2010 will be called 2009, for example
# first need to add month and year columns
flow <- flow %>% mutate(year = year(Date),
                        month = month(Date),
                        waterYear = case_when(month %in% c(7:12) ~ year,
                                              month %in% c(1:6) ~ year - 1))

inun <- read.csv("raw/cumbung_inundation.csv") %>% 
  rename(waterYear = Date)
```


```{r summarise flow data for water year, july to june}
flow_WY <- flow %>% group_by(waterYear) %>% 
                    summarise(discharge_WY = sum(Discharge, na.rm = TRUE),
                              n = sum(!is.na(Discharge)))

# only keep years with >90% data, i.e. at least 329 days of data (ideally, would filter to 95%, but this would remove too many data points)

flow_WY <- flow_WY %>% filter(n >= 329)
```


```{r combine discharge and waterbirds datasets}
wb_flow <- right_join(waterbird, flow_WY) %>% 
# right join as flow dataset has one missing year
           mutate(ln.dis = log(discharge_WY)) %>% 
           select(!n)

wb_inun <- right_join(waterbird, inun) %>% 
           mutate(ln.inun = log(Water_area_km2))
```


# Generalised linear model - discharge vs abundance

Assumptions:
- Independance of each data points
- Correct distribution of the residuals (normal distribution of residuals)
- Correct specification of the variance structure (equal variances of residuals)
- Linear relationship between the response and the linear predictor


```{r function for plotting}
birdplot <- function(predict_data, title){
  plot(predict_data, facet = FALSE) + 
  labs(x="Flow (ML/year)",y="Waterbirds",title=title) +
  theme_bw() +
  geom_line(linewidth = 1.2) +
  geom_point(data = ggemmeans.Inun, aes(x = x, y = attr(ggemmeans.Inun, "rawdata")$response), size = 6) +
  theme(axis.title = element_text(size = 64), # change font size
        axis.text = element_text(size = 48),
        axis.text.x = element_text(vjust = -0.5),
        axis.title.x = element_text(vjust = -1),
        axis.text.y = element_text(margin = margin(t = 0, r = 10, b = 0, l = 0)),
        panel.grid.major = element_blank(), # remove grid lines
        panel.grid.minor = element_blank(),
        plot.margin = margin(10,20,30,10), # extra space on RHS
        plot.title = element_text(hjust = 0.5)) # centre title
}

```



```{r discharge}
# generalised linear model
M1 <- glm(ln.bird ~ ln.dis, data=wb_flow)
summary(M1)
cor.test(wb_flow$ln.bird, wb_flow$ln.dis)
# p = 0.08

# check assumptions
par(mfrow = c(2, 2))
plot(M1) # top left equal variances, top right normality 
par(mfrow = c(1, 1))
plot(wb_flow$ln.dis,wb_flow$ln.bird) # roughly linear relationship
# assumptions ok

# predict
ggemmeans.Inun <- ggemmeans(M1, terms = c("ln.dis[all]"))

# plot
options(scipen = 999) # turn off scientific notation
cplot <- birdplot(ggemmeans.Inun, "")  +
  scale_y_continuous(breaks = pretty(wb_flow$ln.bird, n = 4), labels = signif(exp(pretty(wb_flow$ln.bird, n = 4)),1)) +
  scale_x_continuous(breaks = pretty(wb_flow$ln.dis,n = 3), 
                     labels = signif(exp(pretty(wb_flow$ln.dis, n = 3)),1))  
cplot

# save plot
png("output2/birdflow.png", height = 900, width = 1200)
cplot
dev.off()
```




```{r inundation}
# generalised linear model
M1 <- glm(ln.bird ~ ln.inun, data=wb_inun)
summary(M1)
cor.test(wb_inun$ln.bird, wb_inun$ln.inun)
# p = 0.08

# check assumptions
par(mfrow = c(2, 2))
plot(M1) # top left equal variances, top right normality 
par(mfrow = c(1, 1))
plot(wb_inun$ln.inun,wb_inun$ln.bird) # not a non linear relationship
# assumptions ok

# predict
ggemmeans.Inun <- ggemmeans(M1, terms = c("ln.inun[all]"))

# plot
options(scipen = 999) # turn off scientific notation
cplot4 <- birdplot(ggemmeans.Inun, "")  +
  scale_y_continuous(breaks = pretty(wb_inun$ln.bird, n = 4), labels = signif(exp(pretty(wb_inun$ln.bird, n = 4)),1)) +
  scale_x_continuous(breaks = pretty(wb_inun$ln.inun,n = 3), 
                     labels = signif(exp(pretty(wb_inun$ln.inun, n = 3)),1))  + 
  labs(x="Inundation (km2)",y="Waterbirds")
cplot4

# save plot
png("output2/birdinun.png", height = 900, width = 1200)
cplot4
dev.off()
```



# Generalised linear model - time vs abundance

```{r cumbung}
# generalised linear model
M1 <- glm(ln.bird ~ waterYear, data=wb_flow)
summary(M1)
cor.test(wb_flow$ln.bird, wb_flow$waterYear)

# check assumptions
par(mfrow = c(2, 2))
plot(M1) # top left equal variances, top right normality 
par(mfrow = c(1, 1))
plot(wb_flow$waterYear,wb_flow$ln.bird) # linear relationship
# assumptions ok

# predict
ggemmeans.Inun <- ggemmeans(M1, terms = c("waterYear[all]"))

# plot
cplot2 <- birdplot(ggemmeans.Inun, "") +
  scale_x_continuous(breaks = c(2010,2015,2020)) +
  theme(title = element_text(size = 28))  +
  scale_y_continuous(breaks = pretty(wb_flow$ln.bird), 
                     labels = signif(exp(pretty(wb_flow$ln.bird)),1))
cplot2

# save plot
png("output2/birdtime.png", height = 900, width = 1200) 
cplot2
dev.off()

# plot for presentation
cplot3 <- birdplot(ggemmeans.Inun, "") +
  labs(x= "Year", y = "Waterbirds") +
  scale_x_continuous(breaks = c(2010,2015,2020)) +
  scale_y_continuous(breaks = pretty(wb_flow$ln.bird), 
                     labels = signif(exp(pretty(wb_flow$ln.bird)),1))
cplot3

wb_flow$predict.Time <- predict(M1, wb_flow)

# calculate differences over time
100-(wb_flow$predict.Time[length(wb_flow$predict.Time)]/wb_flow$predict.Time[1]*100)
```


# Ramsar values

```{r Waterbird_summary_stats}
# average birds per year
mean(waterbird$Count)
max(waterbird$Count)
```


For significant proportions below, note 1990 data that were collected separately (by Maher) were not included in these analysis and were added separately to results tables created.

```{r Significant proportions}
# subset waterbird population data and waterbird species data for only essential cols
# there is an issue with blank spaces displaying separately so make these all the same, otherwise the datasets won't join properly
wbirdpop2 <- as.data.frame(wbirdpop) %>% select(Scientific, CommonName, `1% Threshold`) %>% 
  mutate(Scientific = str_replace_all(Scientific, "[:space:]", " "),
         CommonName = str_replace_all(CommonName, "[:space:]", " "))

waterbirdSp2 <- as.data.frame(waterbirdSp) %>% select(Scientific, CommonName, SurvYear, Count) %>% 
  mutate(Scientific = str_replace_all(Scientific, "[:space:]", " "),
         CommonName = str_replace_all(CommonName, "[:space:]", " "))

# merge datasets
wbirdSpCountPropSig <- right_join(wbirdpop2, waterbirdSp2, by = c("Scientific", "CommonName"))

# create a column for proportion of each year's count of the significant threshold; if above 1, there is a significant proportion for that species that year.
wbirdSpCountPropSig <- wbirdSpCountPropSig %>% mutate(propThreshold = Count/`1% Threshold`)

# subset only for Significant Proportions (>1) - this dataframe lists all instances that species exceeded thresholds
wbirdSpSigPropOnly <- wbirdSpCountPropSig %>% subset(propThreshold >= 1)

# export as csv
write.csv(wbirdSpSigPropOnly, file = "output2/csv/SigProportions.csv")

# number of unique species that exceeded proportions
unique(wbirdSpSigPropOnly$CommonName)
```

