---
title: "Waterbirds"
author: "Dana Lanceman"
date: "8/23/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = 'C:/Users/danal/OneDrive - UNSW/Desktop/Honours_UNSW/Assessment/cumbung_assessment')


# remove scientific notation
options(scipen = 100)
```

```{r setup}
library(tidyverse)
library(readxl)
library(viridis)
library(ggeffects)
library(ggpubr)

c <- read_excel("raw/waterbird_trends.xlsx", "cumbung") %>% 
  mutate(ln.dis = log(Discharge),
         ln.bird = log(Waterbirds))
b <- read_excel("raw/waterbird_trends.xlsx", "booligal") %>% 
  mutate(ln.dis = log(Discharge),
         ln.bird = log(Waterbirds))
g <- read_excel("raw/waterbird_trends.xlsx", "gwydir") %>% 
  mutate(ln.dis = log(Discharge),
         ln.bird = log(Waterbirds))
m <- read_excel("raw/waterbird_trends.xlsx", "macquarie") %>% 
  mutate(ln.dis = log(Discharge),
         ln.bird = log(Waterbirds))
r <- read_excel("raw/waterbird_trends.xlsx", "riverland") %>% 
  mutate(ln.dis = log(Discharge),
         ln.bird = log(Waterbirds))
y <- read_excel("raw/waterbird_trends.xlsx", "yantabulla") %>% 
  mutate(ln.dis = log(Discharge),
         ln.bird = log(Waterbirds))
p <- read_excel("raw/waterbird_trends.xlsx", "paroo") %>% 
  mutate(ln.dis = log(Discharge),
         ln.bird = log(Waterbirds))
h <- read_excel("raw/waterbird_trends.xlsx", "hattah") %>% 
  mutate(ln.dis = log(Discharge),
         ln.bird = log(Waterbirds))
l <- read_excel("raw/waterbird_trends.xlsx", "lowbidgee") %>% 
  mutate(ln.dis = log(Discharge),
         ln.bird = log(Waterbirds))
```


# Generalised linear models - discharge vs abundance

Assumptions:
- Independance of each data points
- Correct distribution of the residuals (normal distribution of residuals)
- Correct specification of the variance structure (equal variances of residuals)
- Linear relationship between the response and the linear predictor


```{r function for plotting}
birdplot <- function(predict_data, title){
  plot(predict_data, facet = FALSE, colors = "hero", ci = T, dot.alpha = 1, dot.size = 5, add.data = TRUE, jitter = 0, show.y.title = T, show.title = FALSE) + 
  labs(x="",y="",title=title) +
  theme_bw() +
  geom_line(size = 1.2) +
  theme(text = element_text(size = 32), # change font size
        axis.text = element_text(size = 28),
        panel.grid.major = element_blank(), # remove grid lines
        panel.grid.minor = element_blank(),
        plot.margin = margin(10,20,10,10), # extra space on RHS
        plot.title = element_text(hjust = 0.5)) # centre title
}

```



```{r cumbung}
# generalised linear model
M1 <- glm(ln.bird ~ ln.dis, data=c)
summary(M1)
cor.test(c$ln.bird, c$ln.dis)
# p = 0.0179

# check assumptions
par(mfrow = c(2, 2))
plot(M1) # top left equal variances, top right normality 
par(mfrow = c(1, 1))
plot(c$ln.dis,c$ln.bird) # linear relationship
# assumptions ok

# predict
ggemmeans.Inun <- ggemmeans(M1, terms = c("ln.dis[all]"))

# plot
cplot <- birdplot(ggemmeans.Inun, "Great Cumbung Swamp*")  +
  scale_y_continuous(breaks = pretty(c$ln.bird), labels = signif(exp(pretty(c$ln.bird)),1)) +
  scale_x_continuous(breaks = pretty(c$ln.dis,n = 3), 
                     labels = signif(exp(pretty(c$ln.dis, n = 3)),1)) +
  theme(title = element_text(size = 28),
        plot.margin = margin(10,30,10,10)) # extra space on RHS
cplot

# presentation plot
cplot4 <- birdplot(ggemmeans.Inun, "")  +
  labs(x = "Flows (ML/day)", y = "Waterbird abundance") +
  scale_y_continuous(breaks = pretty(c$ln.bird), labels = signif(exp(pretty(c$ln.bird)),1)) +
  scale_x_continuous(breaks = pretty(c$ln.dis,n = 3), 
                     labels = signif(exp(pretty(c$ln.dis, n = 3)),1)) +
  theme(title = element_text(size = 28),
        plot.margin = margin(10,30,10,10)) # extra space on RHS
cplot4
```

```{r booligal}
# generalised linear model
M1 <- glm(ln.bird ~ ln.dis, data=b)
summary(M1)
cor.test(b$ln.bird, b$ln.dis)
# p = 0.0142

# check assumptions
par(mfrow = c(2, 2))
plot(M1) # top left equal variances, top right normality 
par(mfrow = c(1, 1))
plot(b$ln.dis, b$ln.bird) # linear relationship
# assumptions ok, one outlier (2010 = 389000 waterbirds)

# predict
ggemmeans.Inun <- ggemmeans(M1, terms = c("ln.dis[all]"))

bplot <- birdplot(ggemmeans.Inun, "Booligal Wetlands*")  +
  scale_y_continuous(breaks = pretty(b$ln.bird), 
                     labels = signif(exp(pretty(b$ln.bird)),1)) +
  scale_x_continuous(breaks = pretty(b$ln.dis, n = 3), 
                     labels = signif(exp(pretty(b$ln.dis, n = 3)),1))
bplot
```

```{r gwydir}
# generalised linear model
M1 <- glm(ln.bird ~ ln.dis, data=g)
summary(M1)
cor.test(g$ln.bird, g$ln.dis)
# p = 0.1766

# check assumptions
par(mfrow = c(2, 2))
plot(M1) # top left equal variances, top right normality 
par(mfrow = c(1, 1))
plot(g$ln.dis, g$ln.bird) # linear relationship
# assumptions seem ok

# predict
ggemmeans.Inun <- ggemmeans(M1, terms = c("ln.dis[all]"))

gplot <- birdplot(ggemmeans.Inun, "Gwydir Wetlands")+
  scale_y_continuous(breaks = pretty(g$ln.bird, n = 4), 
                     labels = signif(exp(pretty(g$ln.bird, n = 4)),1)) +
  scale_x_continuous(breaks = pretty(g$ln.dis, n = 4), 
                     labels = signif(exp(pretty(g$ln.dis, n = 4)),1))
gplot
```

```{r hattah}
# generalised linear model
M1 <- glm(ln.bird ~ ln.dis, data=h)
summary(M1)
cor.test(h$ln.bird, h$ln.dis)
# p = 0.1633

# check assumptions
par(mfrow = c(2, 2))
plot(M1) # top left equal variances, top right normality 
par(mfrow = c(1, 1))
plot(h$ln.dis,h$ln.bird) # linear relationship
# Assumtions seem ok

# predict
ggemmeans.Inun <- ggemmeans(M1, terms = c("ln.dis[all]"))

hplot <- birdplot(ggemmeans.Inun, "Hattah Lakes") +
  scale_y_continuous(breaks = pretty(h$ln.bird), 
                     labels = signif(exp(pretty(h$ln.bird)),1)) +
  scale_x_continuous(breaks = pretty(h$ln.dis, n = 3), 
                     labels = signif(exp(pretty(h$ln.dis, n = 3)),1))
hplot
```

```{r lowbidgee}
# generalised linear model
M1 <- glm(ln.bird ~ ln.dis, data=l)
summary(M1)
cor.test(l$ln.bird, l$ln.dis)
# p = 0.0233

# check assumptions
par(mfrow = c(2, 2))
plot(M1) # top left equal variances, top right normality 
par(mfrow = c(1, 1))
plot(l$ln.dis,l$ln.bird) # linear relationship
# assumptions seem ok

# predict
ggemmeans.Inun <- ggemmeans(M1, terms = c("ln.dis[all]"))

lplot <- birdplot(ggemmeans.Inun, "Lowbidgee Wetlands*") +
  scale_y_continuous(breaks = pretty(l$ln.bird, n = 4),
                     labels = signif(exp(pretty(l$ln.bird, n = 4)),1)) +
  scale_x_continuous(breaks = pretty(l$ln.dis), 
                     labels = signif(exp(pretty(l$ln.dis)),1))
lplot
```

```{r macquarie}
# generalised linear model
M1 <- glm(ln.bird ~ ln.dis, data=m)
summary(M1)
cor.test(m$ln.bird, m$ln.dis)
# p = 0.000366

# check assumptions
par(mfrow = c(2, 2))
plot(M1) # top left equal variances, top right normality 
par(mfrow = c(1, 1))
plot(m$ln.dis,m$ln.bird) # linear relationship
# assumptions OK

# predict
ggemmeans.Inun <- ggemmeans(M1, terms = c("ln.dis[all]"))

mplot <- birdplot(ggemmeans.Inun, "Macquarie Marshes*") +
  scale_y_continuous(breaks = pretty(m$ln.bird), 
                     labels = signif(exp(pretty(m$ln.bird)),1)) +
  scale_x_continuous(breaks = pretty(m$ln.dis), 
                     labels = signif(exp(pretty(m$ln.dis)),1)) +
  theme(title = element_text(size = 30)) 
mplot
```

```{r paroo}
# generalised linear model
M1 <- glm(ln.bird ~ ln.dis, data=p)
summary(M1)
cor.test(p$ln.bird, p$ln.dis)
# p = 0.66976

# check assumptions
par(mfrow = c(2, 2))
plot(M1) # top left equal variances, top right normality 
par(mfrow = c(1, 1))
plot(p$ln.dis, p$ln.bird) # linear relationship
# assumptions OK

# predict
ggemmeans.Inun <- ggemmeans(M1, terms = c("ln.dis[all]"))

pplot <- birdplot(ggemmeans.Inun, "Paroo Overflow Swamps") +
  scale_y_continuous(breaks = pretty(p$ln.bird), 
                     labels = signif(exp(pretty(p$ln.bird)),1)) +
  scale_x_continuous(breaks = pretty(p$ln.dis, n = 3),
                     labels = signif(exp(pretty(p$ln.dis, n = 3)),1))  +
  theme(title = element_text(size = 28)) 
pplot
```

```{r riverland}
# generalised linear model
M1 <- glm(ln.bird ~ ln.dis, data=r)
summary(M1)
cor.test(r$ln.bird, r$ln.dis)
# p = 0.348

# check assumptions
par(mfrow = c(2, 2))
plot(M1) # top left equal variances, top right normality 
par(mfrow = c(1, 1))
plot(r$ln.dis,r$ln.bird) # linear relationship
# assumptions ok

# predict
ggemmeans.Inun <- ggemmeans(M1, terms = c("ln.dis[all]"))

rplot <- birdplot(ggemmeans.Inun, "Murray Riverlands") +
  scale_y_continuous(breaks = pretty(r$ln.bird), 
                     labels = signif(exp(pretty(r$ln.bird)),1)) +
  scale_x_continuous(breaks = pretty(r$ln.dis, n = 3), 
                     labels = signif(exp(pretty(r$ln.dis, n = 3)),1))
rplot
```

```{r yantabulla}
# generalised linear model
M1 <- glm(ln.bird ~ ln.dis, data=y)
summary(M1)
cor.test(y$ln.bird, y$ln.dis)
# p = 0.168, but only 4 data points

# check assumptions
par(mfrow = c(2, 2))
plot(M1) # top left equal variances, top right normality 
par(mfrow = c(1, 1))
plot(y$ln.dis, y$ln.bird) # linear relationship
# assumptions ok

# predict
ggemmeans.Inun <- ggemmeans(M1, terms = c("ln.dis[all]"))

yplot <- birdplot(ggemmeans.Inun, "Yantabulla Swamp") +
  scale_y_continuous(breaks = pretty(y$ln.bird, n = 3), 
                     labels = signif(exp(pretty(y$ln.bird, n = 3)),1)) +
  scale_x_continuous(breaks = pretty(y$ln.dis, n = 3), 
                     labels = signif(exp(pretty(y$ln.dis, n = 3)),1)) +
  theme(plot.margin = margin(10,30,10,10))
yplot
```

# Generalised linear models - time vs abundance



```{r cumbung}
# generalised linear model
M1 <- glm(ln.bird ~ Year, data=c)
summary(M1)
cor.test(c$ln.bird, c$Year)

# check assumptions
par(mfrow = c(2, 2))
plot(M1) # top left equal variances, top right normality 
par(mfrow = c(1, 1))
plot(c$Year,c$ln.bird) # linear relationship
# assumptions ok

# predict
ggemmeans.Inun <- ggemmeans(M1, terms = c("Year[all]"))

# plot
cplot2 <- birdplot(ggemmeans.Inun, "Great Cumbung Swamp") +
  scale_x_continuous(breaks = c(2010,2015,2020)) +
  theme(title = element_text(size = 28))  +
  scale_y_continuous(breaks = pretty(c$ln.bird), 
                     labels = signif(exp(pretty(c$ln.bird)),1))
cplot2

# plot for presentation
cplot3 <- birdplot(ggemmeans.Inun, "") +
  labs(x= "Year", y = "Waterbird abundance") +
  scale_x_continuous(breaks = c(2010,2015,2020)) +
  theme(title = element_text(size = 28))  +
  scale_y_continuous(breaks = pretty(c$ln.bird), 
                     labels = signif(exp(pretty(c$ln.bird)),1))
cplot3

c$predict.Time <- predict(M1, c)

# calculate differences over time
100-(c$predict.Time[length(c$predict.Time)]/c$predict.Time[1]*100)
```

```{r booligal}
# generalised linear model
M1 <- glm(ln.bird ~ Year, data=b)
summary(M1)
cor.test(b$ln.bird, b$Year)

# check assumptions
par(mfrow = c(2, 2))
plot(M1) # top left equal variances, top right normality 
par(mfrow = c(1, 1))
plot(b$Year, b$ln.bird) # linear relationship
# assumptions ok

# predict
ggemmeans.Inun <- ggemmeans(M1, terms = c("Year[all]"))

bplot2 <- birdplot(ggemmeans.Inun, "Booligal Wetlands")  +
  scale_x_continuous(breaks = c(2010,2015,2020))  +
  scale_y_continuous(breaks = pretty(b$ln.bird, n = 3), 
                     labels = signif(exp(pretty(b$ln.bird, n = 3)),1))
bplot2

b$predict.Time <- predict(M1, b)

# calculate differences over time
100-(b$predict.Time[length(b$predict.Time)]/b$predict.Time[1]*100)
```

```{r gwydir}
# generalised linear model
M1 <- glm(ln.bird ~ Year, data=g)
summary(M1)
cor.test(g$ln.bird, g$Year)

# check assumptions
par(mfrow = c(2, 2))
plot(M1) # top left equal variances, top right normality 
par(mfrow = c(1, 1))
plot(g$Year, g$ln.bird) # linear relationship
# assumptions seem ok

# predict
ggemmeans.Inun <- ggemmeans(M1, terms = c("Year[all]"))

gplot2 <- birdplot(ggemmeans.Inun, "Gwydir Wetlands")  +
  scale_x_continuous(breaks = c(2010,2015,2020)) +
  scale_y_continuous(breaks = pretty(g$ln.bird, n = 3), 
                     labels = signif(exp(pretty(g$ln.bird, n = 3)),1))
gplot2

g$predict.Time <- predict(M1, g)

# calculate differences over time
100-(g$predict.Time[length(g$predict.Time)]/g$predict.Time[1]*100)
```

```{r hattah}
# generalised linear model
M1 <- glm(ln.bird ~ Year, data=h)
summary(M1)
cor.test(h$ln.bird, h$Year)

# check assumptions
par(mfrow = c(2, 2))
plot(M1) # top left equal variances, top right normality 
par(mfrow = c(1, 1))
plot(h$Year,h$ln.bird) # linear relationship
# Assumtions seem ok

# predict
ggemmeans.Inun <- ggemmeans(M1, terms = c("Year[all]"))

hplot2 <- birdplot(ggemmeans.Inun, "Hattah Lakes")  +
  scale_x_continuous(breaks = c(2007,2011,2015,2020)) +
  scale_y_continuous(breaks = pretty(h$ln.bird), 
                     labels = signif(exp(pretty(h$ln.bird)),1))
hplot2

h$predict.Time <- predict(M1, h)

# calculate differences over time
100-(h$predict.Time[length(h$predict.Time)]/h$predict.Time[1]*100)
```

```{r lowbidgee}
# generalised linear model
M1 <- glm(ln.bird ~ Year, data=l)
summary(M1)
cor.test(l$ln.bird, l$Year)

# check assumptions
par(mfrow = c(2, 2))
plot(M1) # top left equal variances, top right normality 
par(mfrow = c(1, 1))
plot(l$Year,l$ln.bird) # linear relationship
# assumptions seem ok

# predict
ggemmeans.Inun <- ggemmeans(M1, terms = c("Year[all]"))

lplot2 <- birdplot(ggemmeans.Inun, "Lowbidgee Wetlands")  +
  scale_x_continuous(breaks = c(2010,2015,2020))  +
  scale_y_continuous(breaks = pretty(l$ln.bird,n = 3), 
                     labels = signif(exp(pretty(l$ln.bird, n = 3)),1))
lplot2

l$predict.Time <- predict(M1, l)

# calculate differences over time
100-(l$predict.Time[length(l$predict.Time)]/l$predict.Time[1]*100)
```

```{r macquarie}
# generalised linear model
M1 <- glm(ln.bird ~ Year, data=m)
summary(M1)
cor.test(m$ln.bird, m$Year)

# check assumptions
par(mfrow = c(2, 2))
plot(M1) # top left equal variances, top right normality 
par(mfrow = c(1, 1))
plot(m$Year,m$ln.bird) # linear relationship
# assumptions OK

# predict
ggemmeans.Inun <- ggemmeans(M1, terms = c("Year[all]"))

mplot2 <- birdplot(ggemmeans.Inun, "Macquarie Marshes")  +
  scale_x_continuous(breaks = c(2010,2015,2020))  +
  scale_y_continuous(breaks = pretty(m$ln.bird, n = 3), 
                     labels = signif(exp(pretty(m$ln.bird, n = 3)),1))
mplot2

m$predict.Time <- predict(M1, m)

# calculate differences over time
100-(m$predict.Time[length(m$predict.Time)]/m$predict.Time[1]*100)
```

```{r paroo}
# generalised linear model
M1 <- glm(ln.bird ~ Year, data=p)
summary(M1)
cor.test(p$ln.bird, p$Year)

# check assumptions
par(mfrow = c(2, 2))
plot(M1) # top left equal variances, top right normality 
par(mfrow = c(1, 1))
plot(p$Year, p$ln.bird) # linear relationship
# assumptions OK

# predict
ggemmeans.Inun <- ggemmeans(M1, terms = c("Year[all]"))

pplot2 <- birdplot(ggemmeans.Inun, "Paroo Overflow Swamps")  +
  scale_x_continuous(breaks = c(2010,2015,2020)) +
  theme(title = element_text(size = 28))  +
  scale_y_continuous(breaks = pretty(p$ln.bird), 
                     labels = signif(exp(pretty(p$ln.bird)),1))
pplot2

p$predict.Time <- predict(M1, p)

# calculate differences over time
100-(p$predict.Time[length(p$predict.Time)]/p$predict.Time[1]*100)
```

```{r riverland}
# generalised linear model
M1 <- glm(ln.bird ~ Year, data=r)
summary(M1)
cor.test(r$ln.bird, r$Year)

# check assumptions
par(mfrow = c(2, 2))
plot(M1) # top left equal variances, top right normality 
par(mfrow = c(1, 1))
plot(r$Year,r$ln.bird) # linear relationship
# assumptions ok

# predict
ggemmeans.Inun <- ggemmeans(M1, terms = c("Year[all]"))

rplot2 <- birdplot(ggemmeans.Inun, "Murray Riverlands") +
  scale_x_continuous(breaks = c(2008,2012,2016,2020))  +
  scale_y_continuous(breaks = pretty(r$ln.bird), 
                     labels = signif(exp(pretty(r$ln.bird)),1))
rplot2

r$predict.Time <- predict(M1, r)

# calculate differences over time
100-(r$predict.Time[length(r$predict.Time)]/r$predict.Time[1]*100)
```

```{r yantabulla}
# generalised linear model
M1 <- glm(ln.bird ~ Year, data=y)
summary(M1)
cor.test(y$ln.bird, y$Year)

# check assumptions
par(mfrow = c(2, 2))
plot(M1) # top left equal variances, top right normality 
par(mfrow = c(1, 1))
plot(y$Year, y$ln.bird) # linear relationship
# assumptions ok

# predict
ggemmeans.Inun <- ggemmeans(M1, terms = c("Year[all]"))

yplot2 <- birdplot(ggemmeans.Inun, "Yantabulla Swamp")  +
  scale_x_continuous(breaks = c(2010,2015,2020))  +
  scale_y_continuous(breaks = pretty(y$ln.bird, n = 3), 
                     labels = signif(exp(pretty(y$ln.bird, n = 3)),1))
yplot2

y$predict.Time <- predict(M1, y)

# calculate differences over time
100-(y$predict.Time[length(y$predict.Time)]/y$predict.Time[1]*100)
```


# Plotting

```{r put discharge plots together}
# export at 2100 x 2800 

bird <- ggarrange(cplot, bplot, gplot, hplot, lplot, mplot, rplot, pplot,yplot,  
          #labels = c("A*", "B*", "C*", "D*", "E", "F", "G", "H", "I"),
          ncol = 3, nrow = 3,
          #hjust = -0.5, # moves plot labels right
          font.label = list(size = 40), align = "hv")

annotate_figure(bird, bottom = text_grob("Average daily flow volume (ML)",  size = 60, just = "bottom"), left = text_grob("Waterbird abundance", size = 60, rot = 90)) # rotate 90 degrees
# just = "bottom" helps text not be cut off
```


```{r put time plots together}
# export at 2100 x 2900 

bird <- ggarrange(cplot2, bplot2, gplot2, hplot2, lplot2, mplot2, rplot2, pplot2, yplot2,
          ncol = 3, nrow = 3,
          #hjust = -0.5, # moves plot labels right
          font.label = list(size = 40), align = "hv")

annotate_figure(bird, bottom = text_grob("Year",  size = 60, just = "bottom"), left = text_grob("Waterbird abundance", size = 60, rot = 90)) # rotate 90 degrees
# just = "bottom" helps text not be cut off
```

